---
title: "Circumpolar Environmental Data - Crop and Scale"
author: "Jan Jansen"
date: "16/07/2021"
output: html_document
---

```{r, warning=FALSE, message=FALSE}
##### load libraries
library(raster)       ## package for raster manipulation
```

__Circumpolar Environmental Data:  Scale environmental variables for analysis in R__  

## Things you need to do to run this script

#### 1. specify the directory where to find the data

- specify your local directory or point to the owncloud directories (THIS DOESN'T WORK YET)

- Download the data from the link below and change "env.dir" in the code snippet below to match your local machine.

Owncloud repository folder is "EnvironmentalData". All data can be found here:
[link](https://owncloud.imas-data-service.cloud.edu.au/index.php/apps/files/?dir=/EnvironmentalData&fileid=186349)

```{r}
## Jan's local machine:
env.dir <- "C:/Users/jjansen/Desktop/science/data_environmental/"

## remote repository (DOESN'T WORK YET):
# env.dir <- "https://data.imas.utas.edu.au/data_transfer/admin/files/EnvironmentalData/"
```

```{r}
##### setting up directories
env.raw <- paste0(env.dir,"raw/")
env.derived <- paste0(env.dir,"derived/")
AAD_dir <- paste0(env.dir,"raw/accessed_through_R")

string.chr <- "Circumpolar_EnvData_"
string.res <- "500m_"
```

## Load data

```{r, eval=FALSE}

```

## Scale data

```{r, eval=FALSE}

```

## Save data

```{r, eval=FALSE}

```
