---
title: "Circumpolar Environmental Data - ROMS and Net Primary Production"
author: "Jan Jansen"
date: "16/07/2021"
output: html_document
---

```{r, warning=FALSE, message=FALSE}
library(ncdf4) # package for netcdf manipulation
library(raster) # package for raster manipulation
library(rgdal) # package for geospatial analysis
library(ggplot2) # package for plotting
```

__Setup NPP-export file for use in ROMS__  
Either download NPP files manually ([link](http://sites.science.oregonstate.edu/ocean.productivity/cafe.model.php)) or from AAD using the data wrangler.  
Script here with manually downloaded files.


__DON'T JUST RUN THIS SCRIPT, EACH FILE TAKES 6-12h TO COMPLETE !!!__  
  
  
### Specify the directory where to find the data

- specify your local directory or point to the owncloud directories (THIS DOESN'T WORK YET)
- Download the data from the link below and change "env.dir" in the code snippet below to match your local machine.
Owncloud repository folder is "EnvironmentalData". All data can be found here:
[link](https://owncloud.imas-data-service.cloud.edu.au/index.php/apps/files/?dir=/EnvironmentalData&fileid=186349)

```{r, tidy=TRUE}
## Jan's local machine:
env.dir <- "C:/Users/jjansen/Desktop/science/data_environmental/"

## remote repository (DOESN'T WORK YET):
# env.dir <- "https://data.imas.utas.edu.au/data_transfer/admin/files/EnvironmentalData/"
```

### Setting up the environment

```{r}
##### setting up directories
env.raw <- paste0(env.dir,"raw/")
env.derived <- paste0(env.dir,"derived/")
AAD_dir <- paste0(env.dir,"raw/accessed_through_R")

##### specifying values
file.bathy <- "Circumpolar_EnvData_bathy"
file.npp <- "Circumpolar_EnvData_NPP_Cafe"
```

```{r}
#### load ROMS netcdf files (original download from GADI using Filezilla)
## 10k models:
grd10k_nc <- nc_open(paste0(env.raw,"waom10extend_grd.nc"))

#### load data from file with initial conditions
#SST <- ncvar_get(ini10k_nc, varid="SST")

#### 
lon_rho <- ncvar_get(grd10k_nc, varid="lon_rho")
lat_rho <- ncvar_get(grd10k_nc, varid="lat_rho")
# lon_psi <- ncvar_get(grd10k_nc, varid="lon_psi")
# lat_psi <- ncvar_get(grd10k_nc, varid="lat_psi")
# lon_u <- ncvar_get(grd10k_nc, varid="lon_u")
# lat_u <- ncvar_get(grd10k_nc, varid="lat_u")
# lon_v <- ncvar_get(grd10k_nc, varid="lon_v")
# lat_v <- ncvar_get(grd10k_nc, varid="lat_v")
# h <- ncvar_get(grd10k_nc, varid="h")

# range(lon_rho)
# range(lat_rho)
```


## two functions:

```{r}
##### prepare data
prepare_data <- function(year){
  full.extent <- c(-180, 180, -90, 90)
  full.proj <- "+proj=longlat"
  
  ## SO extent and individual boxes to calculate npp-chla relationship
  lat_box.f <- c(-90,-50)
  lon_box.f <- c(-180,180)
  lat_box <- c(-90,-70,-60,-50)
  lon_box <- c(-180,-90,0,90,180)
  
  #### load NPP-export files (CAFE):
  n.hdf.raw <- list.files(paste0(env.raw,"NPP_CAFE_monthly/"),pattern=".hdf", full.names=TRUE)
  n.hdf <- n.hdf.raw[grep(year,n.hdf.raw)]
  r <- raster(n.hdf[[1]])
  r[r<0] <- NA_real_
  projection(r) <- full.proj
  extent(r) <- full.extent 
  s <- stack(r)
  for(i in 2:length(n.hdf)){
    r.i <- raster(n.hdf[i])
    r.i[r.i<0] <- NA_real_
    projection(r.i) <- full.proj
    extent(r.i) <- full.extent
    s[[i]] <- r.i
  }
  #plot(s,xlim=c(130,150),ylim=c(-68,-62))
  npp.s <- s
  rm(s,r,r.i)
  
  #### load chla data
  f.hdf.raw <- list.files(paste0(env.raw,"NPP_CAFE_monthly_input/"),pattern=".hdf", full.names=TRUE)
  f.hdf.chl <- f.hdf.raw[grep("chl",f.hdf.raw)]
  f.hdf <- f.hdf.chl[grep(year,f.hdf.chl)]
  r <- raster(f.hdf[1])
  r[r<0] <- NA_real_
  projection(r) <- full.proj
  extent(r) <- full.extent
  s <<- stack(r)
  for(i in 2:length(f.hdf)){
    r.i <- raster(f.hdf[i])
    r.i[r.i<0] <- NA_real_
    projection(r.i) <- full.proj
    extent(r.i) <- full.extent
    s[[i]] <- r.i
  }
  chl.s <- s
  rm(s,r,r.i)
  
  #### crop raster to SO-extent
  npp.filled <<- npp.s.crop <<- crop(npp.s,extent(lon_box.f[1:2],lat_box.f))
  chl.crop <<- crop(chl.s,extent(lon_box.f[1:2],lat_box.f))#,snap="in")
  
  ########## which cells have no NPP-value but a chl-a value
  #### NPP and chla values from raster by lonlat-box
  chl.vals.boxes.l <- list()
  npp.vals.boxes.l <- list()
  
  chl.vals.boxes.l[[1]] <- crop(chl.s,extent(lon_box[1:2],lat_box[3:4]),snap="in")
  chl.vals.boxes.l[[2]] <- crop(chl.s,extent(lon_box[2:3],lat_box[3:4]),snap="in")
  chl.vals.boxes.l[[3]] <- crop(chl.s,extent(lon_box[3:4],lat_box[3:4]),snap="in")
  chl.vals.boxes.l[[4]] <- crop(chl.s,extent(lon_box[4:5],lat_box[3:4]),snap="in")
  
  chl.vals.boxes.l[[5]] <- crop(chl.s,extent(lon_box[1:2],lat_box[2:3]),snap="in")
  chl.vals.boxes.l[[6]] <- crop(chl.s,extent(lon_box[2:3],lat_box[2:3]),snap="in")
  chl.vals.boxes.l[[7]] <- crop(chl.s,extent(lon_box[3:4],lat_box[2:3]),snap="in")
  chl.vals.boxes.l[[8]] <- crop(chl.s,extent(lon_box[4:5],lat_box[2:3]),snap="in")
  
  chl.vals.boxes.l[[9]] <- crop(chl.s,extent(lon_box[1:2],lat_box[1:2]),snap="in")
  chl.vals.boxes.l[[10]] <- crop(chl.s,extent(lon_box[2:3],lat_box[1:2]),snap="in")
  chl.vals.boxes.l[[11]] <- crop(chl.s,extent(lon_box[3:4],lat_box[1:2]),snap="in")
  chl.vals.boxes.l[[12]] <- crop(chl.s,extent(lon_box[4:5],lat_box[1:2]),snap="in")
  
  npp.vals.boxes.l[[1]] <- crop(npp.s,extent(lon_box[1:2],lat_box[3:4]),snap="in")
  npp.vals.boxes.l[[2]] <- crop(npp.s,extent(lon_box[2:3],lat_box[3:4]),snap="in")
  npp.vals.boxes.l[[3]] <- crop(npp.s,extent(lon_box[3:4],lat_box[3:4]),snap="in")
  npp.vals.boxes.l[[4]] <- crop(npp.s,extent(lon_box[4:5],lat_box[3:4]),snap="in")
  
  npp.vals.boxes.l[[5]] <- crop(npp.s,extent(lon_box[1:2],lat_box[2:3]),snap="in")
  npp.vals.boxes.l[[6]] <- crop(npp.s,extent(lon_box[2:3],lat_box[2:3]),snap="in")
  npp.vals.boxes.l[[7]] <- crop(npp.s,extent(lon_box[3:4],lat_box[2:3]),snap="in")
  npp.vals.boxes.l[[8]] <- crop(npp.s,extent(lon_box[4:5],lat_box[2:3]),snap="in")
  
  npp.vals.boxes.l[[9]] <- crop(npp.s,extent(lon_box[1:2],lat_box[1:2]),snap="in")
  npp.vals.boxes.l[[10]] <- crop(npp.s,extent(lon_box[2:3],lat_box[1:2]),snap="in")
  npp.vals.boxes.l[[11]] <- crop(npp.s,extent(lon_box[3:4],lat_box[1:2]),snap="in")
  npp.vals.boxes.l[[12]] <- crop(npp.s,extent(lon_box[4:5],lat_box[1:2]),snap="in")
  
  chl.vals.boxes <<- chl.vals.boxes.l
  npp.vals.boxes <<- npp.vals.boxes.l
}

##### loop over monthly files
monthly_loop <- function(){
  for(j in 1:nlayers(chl.crop)){
    #Sys.time(
      message(paste0("file ",j," of ",nlayers(chl.crop)))
      chl.data <- chl.crop@data@values[,j]
      npp.data <- npp.s.crop@data@values[,j]
      npp.nas <- which(is.na(npp.data))
      chl.nas <- which(is.na(chl.data))
      sel <- which(npp.nas%!in%chl.nas)
      
      ## find which geographical section a missing value comes from
      lonlats <- data.frame(xyFromCell(npp.filled[[j]],npp.nas[sel]))
      if(nrow(lonlats)==0){
        message("no data found to replace, skipping to next iteration")
        next
      }
      lonlats$LonBox <- NA_character_
      lonlats$LatBox <- NA_character_
      lonlats$LatBox[lonlats[,2]>-60] <- "Lat50"
      lonlats$LatBox[lonlats[,2]>-70&lonlats[,2]<=-60] <- "Lat60"
      lonlats$LatBox[lonlats[,2]<=-70] <- "Lat70"
      lonlats$LonBox[lonlats[,1]<=-90] <- "Lon1"
      lonlats$LonBox[lonlats[,1]<=0 & lonlats[,1]>-90] <- "Lon2"
      lonlats$LonBox[lonlats[,1]<=90 & lonlats[,1]>0] <- "Lon3"
      lonlats$LonBox[lonlats[,1]>90] <- "Lon4"
      # lonlats$LonBox <- as.factor(lonlats$LonBox)
      # lonlats$LatBox <- as.factor(lonlats$LatBox)
      lonlats$Box <- factor(apply(lonlats[,3:4],1,paste,collapse="_"))
      box.sel <- list()
      box.sel[[1]] <- which(lonlats$Box=="Lon1_Lat50")
      box.sel[[2]] <- which(lonlats$Box=="Lon2_Lat50")
      box.sel[[3]] <- which(lonlats$Box=="Lon3_Lat50")
      box.sel[[4]] <- which(lonlats$Box=="Lon4_Lat50")
      box.sel[[5]] <- which(lonlats$Box=="Lon1_Lat60")
      box.sel[[6]] <- which(lonlats$Box=="Lon2_Lat60")
      box.sel[[7]] <- which(lonlats$Box=="Lon3_Lat60")
      box.sel[[8]] <- which(lonlats$Box=="Lon4_Lat60")
      box.sel[[9]] <- which(lonlats$Box=="Lon1_Lat70")
      box.sel[[10]] <- which(lonlats$Box=="Lon2_Lat70")
      box.sel[[11]] <- which(lonlats$Box=="Lon3_Lat70")
      box.sel[[12]] <- which(lonlats$Box=="Lon4_Lat70")
      
      ## fill npp value from npp-chla-relation in each area (loop over the 12 areas)
      for(k in 1:12){
        print(paste0("box ",k," of 12"))
        ## choose latlon-box
        sel.b <- npp.nas[sel[box.sel[[k]]]]
        ## chla-value in that box
        #chl.data[sel.b]
        ## chla- & NPP-values in that box
        chl.v.b <- chl.vals.boxes[[k]]@data@values[,j]
        npp.v.b <- npp.vals.boxes[[k]]@data@values[,j]
        ## the box at 0-90 lon and south of 70 lat has almost no values,: combine with the 60-70 lat box
        if(k==11){
          chl.v.b <- c(chl.vals.boxes[[7]]@data@values[,j],chl.vals.boxes[[11]]@data@values[,j])
          npp.v.b <- c(npp.vals.boxes[[7]]@data@values[,j],npp.vals.boxes[[11]]@data@values[,j])
        }
        ## loop through each value in that box and find a mean npp value
        for(i in 1:length(sel.b)){
          ## selector to find chl-a values similar to the value that needs replacing (in that latlon-box)
          sel.v.b <- which(chl.v.b<=chl.data[sel.b][i]+0.01 & chl.v.b>=chl.data[sel.b][i]-0.01)
          ## calculate replacement value from mean of these similar values
          npp.replace <- mean(npp.v.b[sel.v.b], na.rm=TRUE)
          ## if it didn't pick up a value, increase search radius -> change threshold if value is NA
          if(is.na(npp.replace)){
            sel.v.b <- which(chl.v.b<=chl.data[sel.b][i]+0.02 & chl.v.b>=chl.data[sel.b][i]-0.02)
            npp.replace <- mean(npp.v.b[sel.v.b], na.rm=TRUE)
            if(is.na(npp.replace)){
              sel.v.b <- which(chl.v.b<=chl.data[sel.b][i]+0.05 & chl.v.b>=chl.data[sel.b][i]-0.05)
              npp.replace <- mean(npp.v.b[sel.v.b], na.rm=TRUE)
              if(is.na(npp.replace)){      
                print("NA detected")
                message(length(sel.v.b))
                print(k)
                print(i)
              } }
          }else {npp.filled[[j]][sel.b[i]] <<- npp.replace}
        } }
#    )
  }}
```

## Extract NPP values from HDF-files and store as raster files

2002-2020

```{r, eval=FALSE}
# year <- "2002"
# year <- "2003"
# year <- "2004"
# year <- "2005"
# year <- "2006"
# year <- "2007"
# year <- "2008"
# year <- "2009"
# year <- "2010"
# year <- "2011"
# year <- "2012"
# year <- "2013"
# year <- "2014"
# year <- "2015"
# year <- "2016"
# year <- "2017"
# year <- "2018"
# year <- "2019"

# year <- "2020"

prepare_data(year)
monthly_loop()

save(npp.filled,file=paste0(env.derived,file.npp,"_filled_",year,".Rdata"))
save(npp.s.crop,file=paste0(env.derived,file.npp,"_original_",year,".Rdata"))
```

## Extract NPP values from raster files and place them into the netcdf file for use in ROMS

```{r, eval=FALSE}
coords <- cbind(c(lon_rho),c(lat_rho))

#### load data:
# year <- "2002" ## year 2002 only has data for October-December: Need to tweak the code below in that case
# year <- "2003"
# year <- "2004"
# year <- "2005"
# year <- "2006"
# year <- "2007"
# year <- "2008"
# year <- "2009"
# year <- "2010"
# year <- "2011"
# year <- "2012"
# year <- "2013"
# year <- "2014"
# year <- "2015"
# year <- "2016"
# year <- "2017"
# year <- "2018"
# year <- "2019"
# year <- "2020"

load(file=paste0(env.derived,file.npp,"_filled_",year,".Rdata"))

# npp10 <- extract(npp.filled$cafe.2002274, coords)
# npp11 <- extract(npp.filled$cafe.2002305, coords)
# npp12 <- extract(npp.filled$cafe.2002335, coords)

npp1 <- extract(npp.filled[[1]], coords)
npp2 <- extract(npp.filled[[2]], coords)
npp3 <- extract(npp.filled[[3]], coords)
npp10 <- extract(npp.filled[[6]], coords)
npp11 <- extract(npp.filled[[7]], coords)
npp12 <- extract(npp.filled[[8]], coords)

#### run script:
file.copy(paste0(env.raw,"waom10extend_nudge.nc"),
          paste0(env.derived,"waom10extend_nudge_",year,".nc"))
ini10k_nc <- nc_open(paste0(env.derived,"waom10extend_nudge_",year,".nc"), write=TRUE)

npp_winter <- npp1
npp_winter[npp_winter>0] <-0

npp.v <- c(npp1,npp2,npp3,
           npp_winter,npp_winter,npp_winter,npp_winter,npp_winter,npp_winter,
           npp10,npp11,npp12)
npp.v[is.na(npp.v)] <- 0
## change values from mg/m2/day to kg/m2/day (1 kg is 1 million mg)
npp.v <- npp.v/1000000
## now change values from kg/m2/day to kg/m2/s (1 day is 86400s)
npp.v <- npp.v/86400

## define variable dimensions
xdim2 <- ini10k_nc$dim$xi_rho
ydim2 <- ini10k_nc$dim$eta_rho
tdim2 <- ini10k_nc$dim$sst_time
#tdim2 <- ncdim_def( 'Month', 'count', seq(1,3,by=1)) ## only January to March so far

## define variable
#var_q <- ncvar_def('NPP','mg/m2/day',list(xdim2,ydim2,tdim2),longname="NetPrimaryProduction")
var_1 <- ncvar_def('sand_01_sflux','kg/s',list(xdim2,ydim2,tdim2),longname="NetPrimaryProduction",prec="double")
var_2 <- ncvar_def('sand_02_sflux','kg/s',list(xdim2,ydim2,tdim2),longname="NetPrimaryProduction",prec="double")
var_3 <- ncvar_def('sand_03_sflux','kg/s',list(xdim2,ydim2,tdim2),longname="NetPrimaryProduction",prec="double")
var_4 <- ncvar_def('sand_04_sflux','kg/s',list(xdim2,ydim2,tdim2),longname="NetPrimaryProduction",prec="double")
var_5 <- ncvar_def('sand_05_sflux','kg/s',list(xdim2,ydim2,tdim2),longname="NetPrimaryProduction",prec="double")
var_6 <- ncvar_def('sand_06_sflux','kg/s',list(xdim2,ydim2,tdim2),longname="NetPrimaryProduction",prec="double")
var_7 <- ncvar_def('sand_07_sflux','kg/s',list(xdim2,ydim2,tdim2),longname="NetPrimaryProduction",prec="double")
var_8 <- ncvar_def('sand_08_sflux','kg/s',list(xdim2,ydim2,tdim2),longname="NetPrimaryProduction",prec="double")

## add variable to netcdf and then add values to that variable
ncid_old <- ncvar_add(ini10k_nc,var_1)
ncvar_put(ncid_old, "sand_01_sflux", npp.v)
ncid_old <- ncvar_add(ini10k_nc,var_2)
ncvar_put(ncid_old, "sand_02_sflux", npp.v)
ncid_old <- ncvar_add(ini10k_nc,var_3)
ncvar_put(ncid_old, "sand_03_sflux", npp.v)
ncid_old <- ncvar_add(ini10k_nc,var_4)
ncvar_put(ncid_old, "sand_04_sflux", npp.v)
ncid_old <- ncvar_add(ini10k_nc,var_5)
ncvar_put(ncid_old, "sand_05_sflux", npp.v)
ncid_old <- ncvar_add(ini10k_nc,var_6)
ncvar_put(ncid_old, "sand_06_sflux", npp.v)
ncid_old <- ncvar_add(ini10k_nc,var_7)
ncvar_put(ncid_old, "sand_07_sflux", npp.v)
ncid_old <- ncvar_add(ini10k_nc,var_8)
ncvar_put(ncid_old, "sand_08_sflux", npp.v)

## check output
print(ncid_old)
## close file
nc_close(ncid_old)

## check results
ini10k_nc <- nc_open(paste0(env.derived,"waom10extend_nudge_",year,".nc"))
npp <- ncvar_get(ini10k_nc, "sand_01_sflux")
npp[10,17,3]
matrix((npp3/1000000)/86400, ncol=560)[10,17]
```

## generate a climatology NetCDF-file, starting in July and ending in June

```{r, eval=FALSE}
years <- as.character(2002:2019)

## load all npp files for the years
load(file=paste0(env.derived,file.npp,"_filled_2002.Rdata"))
a <- list()
a[[1]] <- npp.filled
names(a) <- "NPP.2002"
for(i in 2:length(years)){
  j <- years[i]
  load(file=paste0(env.dir,"Circumpolar_NPP_Cafe_filled_",j,".Rdata"))
  a[[i]] <- npp.filled
  names(a)[i] <- paste0("NPP.",years[i])
}

## dataframe of npp-values by years, for each month where there is light/data (1,2,3,10,11,12)
npp1.df <- data.frame(matrix(NA, nrow=nrow(coords),ncol=length(years)-1))
names(npp1.df) <- years[-1]
for(i in 2:length(years)){
  npp1.df[,i-1] <- extract(a[[i]][[1]], coords)
}
npp2.df <- data.frame(matrix(NA, nrow=nrow(coords),ncol=length(years)-1))
names(npp2.df) <- years[-1]
for(i in 2:length(years)){
  npp2.df[,i-1] <- extract(a[[i]][[2]], coords)
}
npp3.df <- data.frame(matrix(NA, nrow=nrow(coords),ncol=length(years)-1))
names(npp3.df) <- years[-1]
for(i in 2:length(years)){
  npp3.df[,i-1] <- extract(a[[i]][[3]], coords)
}
npp10.df <- data.frame(matrix(NA, nrow=nrow(coords),ncol=length(years)))
names(npp10.df) <- years
npp10.df[,1] <- extract(a[[1]][[1]], coords)
for(i in 2:length(years)){
  npp10.df[,i] <- extract(a[[i]][[6]], coords)
}
npp11.df <- data.frame(matrix(NA, nrow=nrow(coords),ncol=length(years)))
names(npp11.df) <- years
npp11.df[,1] <- extract(a[[1]][[2]], coords)
for(i in 2:length(years)){
  npp11.df[,i] <- extract(a[[i]][[7]], coords)
}
npp12.df <- data.frame(matrix(NA, nrow=nrow(coords),ncol=length(years)))
names(npp12.df) <- years
npp12.df[,1] <- extract(a[[1]][[3]], coords)
for(i in 2:length(years)){
  npp12.df[,i] <- extract(a[[i]][[8]], coords)
}

## average monthly npp across all years
npp1 <- rowMeans(npp1.df, na.rm=TRUE)
npp2 <- rowMeans(npp2.df, na.rm=TRUE)
npp3 <- rowMeans(npp3.df, na.rm=TRUE)
npp10 <-rowMeans(npp10.df, na.rm=TRUE)
npp11 <-rowMeans(npp11.df, na.rm=TRUE)
npp12 <-rowMeans(npp12.df, na.rm=TRUE)

#### create NetCDF file
file.copy(paste0(env.raw,"waom10extend_nudge.nc"),
          paste0(env.derived,"waom10extend_nudge_climatology20022019.nc"))
ini10k_nc <- nc_open(paste0(env.derived,"waom10extend_nudge_climatology20022019.nc"), write=TRUE)

npp_winter <- npp1
npp_winter[npp_winter>0] <-0
# npp.v <- c(npp_winter,npp_winter,npp_winter,npp10,npp11,npp12,npp1,npp2,npp3,npp_winter,npp_winter,npp_winter)
npp.v <- c(npp1,npp2,npp3,npp_winter,npp_winter,npp_winter,npp_winter,npp_winter,npp_winter,npp10,npp11,npp12)
npp.v[is.na(npp.v)] <- 0
## change values from mg/m2/day to kg/m2/day (1 kg is 1 million mg)
npp.v <- npp.v/1000000
## now change values from kg/m2/day to kg/m2/s (1 day is 86400s)
npp.v <- npp.v/86400

## define variable dimensions
xdim2 <- ini10k_nc$dim$xi_rho
ydim2 <- ini10k_nc$dim$eta_rho
tdim2 <- ini10k_nc$dim$sst_time
#tdim2$vals <- ini10k_nc$dim$sst_time$vals[c(7:12,1:6)]

## define variable
#var_q <- ncvar_def('NPP','mg/m2/day',list(xdim2,ydim2,tdim2),longname="NetPrimaryProduction")
var_1 <- ncvar_def('sand_01_sflux','kg/s',list(xdim2,ydim2,tdim2),longname="NetPrimaryProduction",prec="double")
var_2 <- ncvar_def('sand_02_sflux','kg/s',list(xdim2,ydim2,tdim2),longname="NetPrimaryProduction",prec="double")
var_3 <- ncvar_def('sand_03_sflux','kg/s',list(xdim2,ydim2,tdim2),longname="NetPrimaryProduction",prec="double")
var_4 <- ncvar_def('sand_04_sflux','kg/s',list(xdim2,ydim2,tdim2),longname="NetPrimaryProduction",prec="double")
var_5 <- ncvar_def('sand_05_sflux','kg/s',list(xdim2,ydim2,tdim2),longname="NetPrimaryProduction",prec="double")
var_6 <- ncvar_def('sand_06_sflux','kg/s',list(xdim2,ydim2,tdim2),longname="NetPrimaryProduction",prec="double")
var_7 <- ncvar_def('sand_07_sflux','kg/s',list(xdim2,ydim2,tdim2),longname="NetPrimaryProduction",prec="double")
var_8 <- ncvar_def('sand_08_sflux','kg/s',list(xdim2,ydim2,tdim2),longname="NetPrimaryProduction",prec="double")

## add variable to netcdf and then add values to that variable
ncid_old <- ncvar_add(ini10k_nc,var_1)
ncvar_put(ncid_old, "sand_01_sflux", npp.v)
ncid_old <- ncvar_add(ini10k_nc,var_2)
ncvar_put(ncid_old, "sand_02_sflux", npp.v)
ncid_old <- ncvar_add(ini10k_nc,var_3)
ncvar_put(ncid_old, "sand_03_sflux", npp.v)
ncid_old <- ncvar_add(ini10k_nc,var_4)
ncvar_put(ncid_old, "sand_04_sflux", npp.v)
ncid_old <- ncvar_add(ini10k_nc,var_5)
ncvar_put(ncid_old, "sand_05_sflux", npp.v)
ncid_old <- ncvar_add(ini10k_nc,var_6)
ncvar_put(ncid_old, "sand_06_sflux", npp.v)
ncid_old <- ncvar_add(ini10k_nc,var_7)
ncvar_put(ncid_old, "sand_07_sflux", npp.v)
ncid_old <- ncvar_add(ini10k_nc,var_8)
ncvar_put(ncid_old, "sand_08_sflux", npp.v)

## check output
print(ncid_old)
## close file
nc_close(ncid_old)

## check results
ini10k_nc <- nc_open(paste0(env.dir,"Circumpolar_ROMS/waom10extend_nudge_climatology20022019.nc"))
npp <- ncvar_get(ini10k_nc, "sand_01_sflux")
npp[10,17,6]
matrix((npp12/1000000)/86400, ncol=560)[10,17]
```

## Bash-code needed to copy files onto GADI for the ROMS

```{r, eval=FALSE}
## copy the file fo Gadi using Filezilla, into folder "/g/data/gi0/jj1924"
## then use bash to remove the offset and factor that are somehow in the file
module load nco
# ncatted -O -a add_offset,sand_01_sflux,d,, waom10extend_nudge_2007.nc
# ncatted -O -a add_offset,sand_02_sflux,d,, waom10extend_nudge_2007.nc
# ncatted -O -a add_offset,sand_03_sflux,d,, waom10extend_nudge_2007.nc
# ncatted -O -a add_offset,sand_04_sflux,d,, waom10extend_nudge_2007.nc
# ncatted -O -a add_offset,sand_05_sflux,d,, waom10extend_nudge_2007.nc
# ncatted -O -a add_offset,sand_06_sflux,d,, waom10extend_nudge_2007.nc
# ncatted -O -a add_offset,sand_07_sflux,d,, waom10extend_nudge_2007.nc
# ncatted -O -a add_offset,sand_08_sflux,d,, waom10extend_nudge_2007.nc
# ncatted -O -a scale_factor,sand_01_sflux,d,, waom10extend_nudge_2007.nc
# ncatted -O -a scale_factor,sand_02_sflux,d,, waom10extend_nudge_2007.nc
# ncatted -O -a scale_factor,sand_03_sflux,d,, waom10extend_nudge_2007.nc
# ncatted -O -a scale_factor,sand_04_sflux,d,, waom10extend_nudge_2007.nc
# ncatted -O -a scale_factor,sand_05_sflux,d,, waom10extend_nudge_2007.nc
# ncatted -O -a scale_factor,sand_06_sflux,d,, waom10extend_nudge_2007.nc
# ncatted -O -a scale_factor,sand_07_sflux,d,, waom10extend_nudge_2007.nc
# ncatted -O -a scale_factor,sand_08_sflux,d,, waom10extend_nudge_2007.nc
# ## then make the file readable:
# chmod 755 waom10extend_nudge_2007.nc

ncatted -O -a add_offset,sand_01_sflux,d,, waom10extend_nudge_2008.nc
ncatted -O -a add_offset,sand_02_sflux,d,, waom10extend_nudge_2008.nc
ncatted -O -a add_offset,sand_03_sflux,d,, waom10extend_nudge_2008.nc
ncatted -O -a add_offset,sand_04_sflux,d,, waom10extend_nudge_2008.nc
ncatted -O -a add_offset,sand_05_sflux,d,, waom10extend_nudge_2008.nc
ncatted -O -a add_offset,sand_06_sflux,d,, waom10extend_nudge_2008.nc
ncatted -O -a add_offset,sand_07_sflux,d,, waom10extend_nudge_2008.nc
ncatted -O -a add_offset,sand_08_sflux,d,, waom10extend_nudge_2008.nc
ncatted -O -a scale_factor,sand_01_sflux,d,, waom10extend_nudge_2008.nc
ncatted -O -a scale_factor,sand_02_sflux,d,, waom10extend_nudge_2008.nc
ncatted -O -a scale_factor,sand_03_sflux,d,, waom10extend_nudge_2008.nc
ncatted -O -a scale_factor,sand_04_sflux,d,, waom10extend_nudge_2008.nc
ncatted -O -a scale_factor,sand_05_sflux,d,, waom10extend_nudge_2008.nc
ncatted -O -a scale_factor,sand_06_sflux,d,, waom10extend_nudge_2008.nc
ncatted -O -a scale_factor,sand_07_sflux,d,, waom10extend_nudge_2008.nc
ncatted -O -a scale_factor,sand_08_sflux,d,, waom10extend_nudge_2008.nc
## then make the file readable:
chmod 755 waom10extend_nudge_2008.nc

ncatted -O -a add_offset,sand_01_sflux,d,, waom10extend_nudge_climatology20022019.nc
ncatted -O -a add_offset,sand_02_sflux,d,, waom10extend_nudge_climatology20022019.nc
ncatted -O -a add_offset,sand_03_sflux,d,, waom10extend_nudge_climatology20022019.nc
ncatted -O -a add_offset,sand_04_sflux,d,, waom10extend_nudge_climatology20022019.nc
ncatted -O -a add_offset,sand_05_sflux,d,, waom10extend_nudge_climatology20022019.nc
ncatted -O -a add_offset,sand_06_sflux,d,, waom10extend_nudge_climatology20022019.nc
ncatted -O -a add_offset,sand_07_sflux,d,, waom10extend_nudge_climatology20022019.nc
ncatted -O -a add_offset,sand_08_sflux,d,, waom10extend_nudge_climatology20022019.nc
ncatted -O -a scale_factor,sand_01_sflux,d,, waom10extend_nudge_climatology20022019.nc
ncatted -O -a scale_factor,sand_02_sflux,d,, waom10extend_nudge_climatology20022019.nc
ncatted -O -a scale_factor,sand_03_sflux,d,, waom10extend_nudge_climatology20022019.nc
ncatted -O -a scale_factor,sand_04_sflux,d,, waom10extend_nudge_climatology20022019.nc
ncatted -O -a scale_factor,sand_05_sflux,d,, waom10extend_nudge_climatology20022019.nc
ncatted -O -a scale_factor,sand_06_sflux,d,, waom10extend_nudge_climatology20022019.nc
ncatted -O -a scale_factor,sand_07_sflux,d,, waom10extend_nudge_climatology20022019.nc
ncatted -O -a scale_factor,sand_08_sflux,d,, waom10extend_nudge_climatology20022019.nc
## then make the file readable:
chmod 755 waom10extend_nudge_climatology20022019.nc
```

## generate file to release floats

A table of active cells/coordinates for each monthly file

```{r, eval=FALSE}
######
coords <- data.frame(cbind(c(lon_rho),c(lat_rho)))
names(coords) <- c("lon","lat")
floats_list <- list()
years <- as.character(2002:2018)

## 
load(file=paste0(env.derived,file.npp,"_filled_2002.Rdata"))
a <- extract(npp.filled, coords)
c <- sum(npp.filled, na.rm=TRUE)
floats_list[[1]] <- cbind(coords, a[,2:4])
for(i in 2:length(years)){
  j <- years[i]
  load(file=paste0(env.derived,file.npp,"_filled_",j,".Rdata"))
  print(npp.filled)
  a <- extract(npp.filled, coords)
  floats_list[[i]] <- cbind(coords, a[,c(1:3,6:8)])
  d <- sum(npp.filled, na.rm=TRUE)
  c <- sum(c,d,na.rm=TRUE)
}
e <- extract(c,coords)
length(which(e>0))

## transform to monthly lat-lon points without NAs
m_floats_list <- list()
for(i in 1:length(floats_list)){
  m_floats_list[[i]] <- list()
  for(j in 1:(ncol(floats_list[[i]])-2)){
    m_floats_list[[i]][[j]] <- floats_list[[i]][which(floats_list[[i]][,j+2]>0),1:2]
}}

## produce .txt-file with monthly release-points
# for(i in 1:length(floats_list)){
#   for(j in 1:length(m_floats_list[[i]])){
#     write.table(m_floats_list[[i]][[j]], file=paste0(env.dir,"Circumpolar_NPP_Cafe_FloatsRelease_",substr(names(floats_list[[i]])[j+2],6,12),".txt"),row.names=FALSE)
#   }}

## single table for ROMS:
sel <- which(e>0)
m <- data.frame(matrix(NA,nrow=length(sel),ncol=12))
names(m) <- c("G","C","T","N","Ft0","Fx0","Fy0","Fz0","Fdt","Fdx","Fdy","Fdz")
m[,1:4] <- 1
m[,c(5,8:12)] <- "0.d0"
m[,6:7] <- coords[sel,]
write.table(m, file=paste0(env.derived,file.npp,"_FloatsRelease.txt"),row.names=FALSE, quote=FALSE)

# ## first look which cells are ever ice-free
# load(file=paste0(env.dir,"Circumpolar_NPP_Cafe_filled_2002.Rdata"))
# npp.filled[npp.filled>0] <- 1
# a <- sum(npp.filled, na.rm=TRUE)
# for(i in c("2003","2004","2005","2006","2007","2008","2009","2010","2011","2012","2013","2014","2015","2016","2017","2018")){
#   load(file=paste0(env.dir,"Circumpolar_NPP_Cafe_filled_",i,".Rdata"))
#   npp.filled[npp.filled>0] <- 1
#   b <- sum(npp.filled, na.rm=TRUE)
#   a <- sum(a,b,na.rm=TRUE)
# }
# plot(a)
```

