---
title: "Circumpolar Data - Prepare Analysis"
author: "Jan Jansen"
date: "20/07/2021"
output: html_document
---

```{r, warning=FALSE, message=FALSE}
library(raster)
library(readxl)
library(readr)
library(dplyr)
library(data.table)
library(proj4)
library(stringr)
```

### Setting up the environment

```{r}
## Jan's local machine:
#sci.dir <-      "C:/Users/jjansen/Desktop/science/"
sci.dir <-      "C:/Users/cgros/code/IMAS/"
env.derived <-  paste0(sci.dir,"data_environmental/derived/")
#bio.dir <-      paste0(sci.dir,"data_biological/")

## remote repository (DOESN'T WORK YET):
# env.dir <- "https://data.imas.utas.edu.au/data_transfer/admin/files/EnvironmentalData/"

## common paths (after "sci.dir")
#tools.dir <-    paste0(sci.dir,"SouthernOceanBiodiversityMapping/Useful_Functions_Tools/")
#ARC_Data.dir <- paste0(sci.dir,"SouthernOceanBiodiversityMapping/ARC_Data/")
tools.dir <-    paste0(sci.dir,"Useful_Functions_Tools/")
ARC_Data.dir <- paste0(sci.dir,"ARC_Data/")

## R-drive paths
RS.dir <- "R:/IMAS/Antarctic_Seafloor/Clean_Data_For_Permanent_Storage/"
ann.dir <- paste0(RS.dir,"AnnotationLibrary_AllFinishedSurveys/")

# Image quality path
image.quality.path <- "R:/IMAS/Antarctic_Seafloor/image_quality_analysis/image_quality_score.csv"

```

```{r}
##### load still and diatom sample locations and bathymetry:

## from "Readin_Circumpolar_DownwardImage_Data.Rmd"
load(paste0(ARC_Data.dir,"Circumpolar_DownwardImages_metadata.Rdata"))

## from "ReadIn_Circumpolar_Environmental_Data.Rmd"
r2 <- raster(paste0(env.derived,"Circumpolar_EnvData_500m_shelf_bathy_gebco_depth.grd"))

##### load coastline
#stereo <- as.character(r2@crs)
#load(paste0(env.derived,"Circumpolar_Coastline.Rdata"))

source(paste0(tools.dir,"bubbleplot.R"))
```

### Data Preparation - BASED ON UNCORRECTED INDIVIDUAL SURVEY FILES

```{r}
## load area estimates for each survey (?Why this comment: "no separate area-files for PS96 & PS18")
area_xls_files <- list.files(RS.dir,full.names=TRUE,pattern=".xlsx")
dat_area <- sapply(area_xls_files, readxl::read_excel, simplify=FALSE, skip=1) %>% bind_rows(.id="id")
dat_area_subset <- dat_area[match(unique(dat_area$`image filename`),dat_area$`image filename`),]

## load Biigle image annotations
Biigle_csv_files <- list.files(paste0(ann.dir,"Counts_annotation_files_individual_surveys/"),full.names=TRUE,recursive=TRUE,pattern="254")
dat_counts <- sapply(Biigle_csv_files, readr::read_csv, simplify=FALSE) %>% bind_rows(.id="id")
dat_counts$Label <- gsub(" ","",dat_counts$label_hierarchy) # fixing label names
dat_counts$Label <- gsub(">","__",dat_counts$Label)
dat_counts$Label <- gsub("-","_",dat_counts$Label)

## load %-cover image annotations
cover_csv_files <- list.files(paste0(ann.dir,"Cover_annotation_files_individual_surveys/"),full.names=TRUE,recursive=TRUE,pattern="_03_")
dat_cover <- sapply(cover_csv_files, readr::read_csv, simplify=FALSE) %>% bind_rows(.id="id")
dat_cover$Label <- gsub("1","",dat_cover$Label) # fixing label names
dat_cover$Label <- gsub("2","",dat_cover$Label)
dat_cover$Label <- gsub("3","",dat_cover$Label)
dat_cover$Label <- gsub("4","",dat_cover$Label)
dat_cover$Label <- gsub("5","",dat_cover$Label)
```

```{r}
## metadata for each image, including cellIDs (things like filename, full_path, surveyID, transectID, cellID, lon, lat, x, y, area, CoralNet, Biigle)
image_metadata <- select(dat.list.clean[[1]],Filename.standardised,lon,lat)
for(i in 2:length(dat.list.clean)){
  message(i)
  print(names(dat.list.clean[[i]]))
  if("lon" %in% names(dat.list.clean[[i]])){ dat.temp <- select(dat.list.clean[[i]],Filename.standardised,lon,lat) }
  if("Lon" %in% names(dat.list.clean[[i]])){ dat.temp <- select(dat.list.clean[[i]],Filename.standardised,Lon,Lat) }
  if("Longitude" %in% names(dat.list.clean[[i]])){ dat.temp <- select(dat.list.clean[[i]],Filename.standardised,Longitude,Latitude) }
  if("GPS_lon" %in% names(dat.list.clean[[i]])){ dat.temp <- select(dat.list.clean[[i]],Filename.standardised,GPS_lon,GPS_lat) }  
  names(dat.temp) <- c("Filename.standardised","lon","lat")
  image_metadata <- rbind(image_metadata,dat.temp)
}
image_metadata$Filename.standardised <- gsub(".tif",".jpg",image_metadata$Filename.standardised)
image_metadata$proj_coord_x <- project(image_metadata[,2:3], proj=crs(r2))$x
image_metadata$proj_coord_y <- project(image_metadata[,2:3], proj=crs(r2))$y
image_metadata$cellID <- extract(r2, image_metadata[,4:5], cellnumbers=TRUE)[,1]
ids <- unique(image_metadata$cellID)
#
image_metadata$area <- NA
# find matching filenames for area estimates
idx <- match(image_metadata$Filename.standardised,dat_area_subset$`image filename`)
fill.idx <- which(!is.na(idx))
search.idx <- idx[fill.idx]
image_metadata$area[fill.idx] <- dat_area_subset$`image area in m²`[search.idx] # replace area values where filenames match
image_metadata$area <- as.numeric(image_metadata$area)
#
image_metadata$cover <- "no"
image_metadata$counts <- "no"
# find matching filenames for cover
idx <- match(image_metadata$Filename.standardised,unique(dat_cover$Name))
fill.idx <- which(!is.na(idx))
search.idx <- idx[fill.idx]
image_metadata$cover[fill.idx] <- "yes" # replace area values where filenames match
# find matching filenames for counts
idx <- match(image_metadata$Filename.standardised,unique(dat_counts$filename))
fill.idx <- which(!is.na(idx))
search.idx <- idx[fill.idx]
image_metadata$counts[fill.idx] <- "yes" # replace area values where filenames match
# check:
image_metadata[which(image_metadata$cover=="yes"),]
```


### Counts

#### site(image)-by-species matrix

__count_images__

```{r}
labs.counts <- unique(dat_counts$Label)
nams.counts <- unique(dat_counts$filename)
dat_counts_image_by_species <- data.frame(matrix(NA,nrow=length(nams.counts),ncol=length(labs.counts)))
rownames(dat_counts_image_by_species) <- nams.counts
colnames(dat_counts_image_by_species) <- labs.counts
for(i in 1:length(nams.counts)){
  sel.r <- which(dat_counts$filename==nams.counts[i])  
  for(j in 1:length(labs.counts)){
    sel.c <- which(dat_counts$Label[sel.r]==labs.counts[j])  
    dat_counts_image_by_species[i,j] <- length(sel.c)
  }
}
```

#### site(cell)-by-species matrix

__count_cells__

```{r}
## create new dataframe using image locations
dat_counts_cell_by_species <- data.frame(matrix(NA,nrow=length(ids),ncol=length(labs.counts)))
rownames(dat_counts_cell_by_species) <- ids
colnames(dat_counts_cell_by_species) <- labs.counts
counts_N <- rep(NA, length(ids))
counts_area <- rep(NA, length(ids))
for(i in 1:length(ids)){
  #print(i)
  sel.r <- which(image_metadata$cellID==ids[i]&image_metadata$counts=="yes") # find images that are part of that cell and annotated in biigle
  if(length(sel.r)==0) next # if none of the images are annotated, skip to the next iteration
  sel.names <- image_metadata$Filename.standardised[sel.r] # find the names of these images
  dat.temp <- dat_counts_image_by_species[which(nams.counts%in%sel.names),] # find annotations from these images from the image-dataset
  counts_N[i] <-  nrow(dat.temp)
  counts_area[i] <- sum(image_metadata$area[sel.r])
  dat_counts_cell_by_species[i,] <- colSums(dat.temp)
}
```

### %-cover

#### site(image)-by-species matrix

__cover_images__

```{r}
labs.cov <- unique(dat_cover$Label)
nams.cov <- unique(dat_cover$Name)
dat_cover_image_by_species <- data.frame(matrix(NA,nrow=length(nams.cov),ncol=length(labs.cov)))
rownames(dat_cover_image_by_species) <- nams.cov
colnames(dat_cover_image_by_species) <- labs.cov
for(i in 1:length(nams.cov)){
  sel.r <- which(dat_cover$Name==nams.cov[i])  
  for(j in 1:length(labs.cov)){
    sel.c <- which(dat_cover$Label[sel.r]==labs.cov[j])  
    dat_cover_image_by_species[i,j] <- length(sel.c)
  }
}
```

#### site(cell)-by-species matrix

__cover_cells__

```{r}
## create new dataframe using image locations
dat_cover_cell_by_species <- data.frame(matrix(NA,nrow=length(ids),ncol=length(labs.cov)))
rownames(dat_cover_cell_by_species) <- ids
colnames(dat_cover_cell_by_species) <- labs.cov
cover_N <- rep(NA, length(ids))
cover_area <- rep(NA, length(ids))
for(i in 1:length(ids)){
  #print(i)
  sel.r <- which(image_metadata$cellID==ids[i]&image_metadata$cover=="yes") # find images that are part of that cell and annotated in coralnet
  if(length(sel.r)==0) next # if none of the images are annotated, skip to the next iteration
  sel.names <- image_metadata$Filename.standardised[sel.r] # find the names of these images
  dat.temp <- dat_cover_image_by_species[which(nams.cov%in%sel.names),] # find annotations from these images from the image-dataset
  cover_N[i] <-  nrow(dat.temp)
  cover_area[i] <- sum(image_metadata$area[sel.r])
  dat_cover_cell_by_species[i,] <- colSums(dat.temp)
}
```

### cell metadata

```{r}
cell.coords <- xyFromCell(r2, ids)
cell.lonlat <- project(cell.coords, proj=crs(r2), inverse=TRUE) 

cell_metadata <- data.frame(cbind(ids,cell.lonlat,cell.coords,cover_N, counts_N, cover_area, counts_area))
names(cell_metadata) <- c("cellID", "lon", "lat", "proj_coord_x", "proj_coord_y", "cover_N", "counts_N", "cover_area", "counts_area")

```

### save output

```{r}
cover_cells <- dat_cover_cell_by_species[-which(is.na(rowSums(dat_cover_cell_by_species))),]
count_cells <- dat_counts_cell_by_species[-which(is.na(rowSums(dat_counts_cell_by_species))),]
cover_images <- dat_cover_image_by_species
count_images <- dat_counts_image_by_species

image_metadata
cell_metadata
```

```{r}
save(image_metadata, cell_metadata, cover_cells, cover_images, count_cells, count_images,
     file=paste0(ARC_Data.dir,"Circumpolar_Annotation_Data.Rdata"))
```

### Inspect resulting dataframes

#### locations across a region

```{r, fig.height=10, fig.width=10}
## check if things are correct:
plot(r2, xlim=c(0,300000),ylim=c(-2100000,-1800000))
points(image_metadata[,4:5])
```

#### a single transect

```{r, fig.height=10, fig.width=10}
plot(r2, xlim=c(282500,283500),ylim=c(-2011500,-2010000))
points(image_metadata[,4:5])
points(image_metadata[which(image_metadata$cover=="yes"),4:5],col="red",cex=2)
points(image_metadata[which(image_metadata$counts=="yes"),4:5],col="blue",cex=3)
text(image_metadata[which(image_metadata$cover=="yes"),4:5], labels=image_metadata[which(image_metadata$cover=="yes"),1], adj=-0.2)
```

#### aggregated cover of some living things on a single transect

```{r}
## projected coordinates for plotting
img.coord <- SpatialPoints(coords=image_metadata[,4:5], proj4string=crs(stereo))

#### create subset of data to plot
## all images in that transect
sel <- which(str_detect(image_metadata$Filename.standardised,"tan1901_065"))
## annotated images in that transect
sel2 <- which(str_detect(image_metadata$Filename.standardised,"tan1901_065")&image_metadata$cover=="yes")
# unique(image_metadata$cellID[sel])
# unique(image_metadata$cellID[sel2])

#### just one transect
sel.images <- grep("tan1901_065", rownames(cover_images))
val <- rowSums(cover_images[sel.images,-c(1,3,4,6,7,10)])

```


```{r, fig.height=10, fig.width=10}
plot(r2, xlim=c(282500,283500),ylim=c(-2011500,-2010000))
points(image_metadata[,4:5])
points(image_metadata[which(image_metadata$cover=="yes"),4:5],col="red",cex=0.5)
points(img.coord[sel2], cex=log(val), col="blue")
```






























```{r, eval=FALSE}

image_metadata$Filename.standardised[sel]
t.idx <- match(image_metadata$Filename.standardised[sel],rownames(cover_images))
t.search.idx <- t.idx[which(!is.na(t.idx))]
rownames(cover_images)[t.search.idx]
cover_images[t.search.idx,which(colSums(cover_images[t.search.idx,])>0)]

t.idx2 <- match(image_metadata$Filename.standardised[sel],rownames(count_images))
t.search.idx2 <- t.idx2[which(!is.na(t.idx2))]
rownames(count_images)[t.search.idx2]
count_images[t.search.idx2,which(colSums(count_images[t.search.idx2,])>0)]















## site(cell)-by-metadata matrix (lon,lat,N,depth,...)
cell_locations <- data.frame(cbind("cellID"=cids,N,"depth"=extract(r2,cids)))

##
plot(dat_cover_cell_by_species$Sub_Fine~cell_locations$depth)

summary(fit <- glm((dat_cover_cell_by_species$Sub_Fine/N/108)~cell_locations$depth+I(cell_locations$depth^2), family="binomial"))





















































loc <- PS961_dat[,c(17,1,2)] ## selects filename, long, lat from survey data for ALL images
loc.select <- which(loc$Image_Filename%in%a$Image_Filename) ## selects from "loc" which filenames match the SUBSET filenames in "a"
pos.mat <- loc[loc.select,]



## load environmental data



## focus on sample images only
sample.dat <- read.csv(paste0(s.dir,"SampleImages_DavidBarnes.csv"))

###
img.names <- sample.dat$image
sample.dat$cropped.area.in.m2[-c(10:16,40:49)] <- a.subset$`image area in m²`[which(a.subset$`image filename`%in%img.names)]

#####################
area.dat <- sample.dat[,-3]
names(area.dat) <- c("Survey","image","area")
write.csv(area.dat, file=paste0(s.dir,"Sample_images_area.csv"))

## %-cover - filenames-by-species
PS81.ann <- read.csv(paste0(RS.dir,"PS81/PS81_4_annotation_library_cover/PS81_annotations_03_201912.csv"))
PS96.ann <- read.csv(paste0(RS.dir,"PS96/PS96_4_annotation_library_cover/PS96_annotations_03_201912.csv"))
TAN0802.ann <- read.csv(paste0(RS.dir,"TAN0802/TAN0802_4_annotation_library_cover/TAN0802_annotations_03_202001.csv"))
TAN1802.ann <- read.csv(paste0(RS.dir,"TAN1802/TAN1802_4_annotation_library_cover/TAN1802_annotations_03_202003.csv"))
TAN1901.ann <- read.csv(paste0(RS.dir,"TAN1901/TAN1901_4_annotation_library_cover/TAN1901_annotations_03_202006.csv"))
NBP1402.ann <- read.csv(paste0(RS.dir,"NBP1402/NBP1402_4_annotation_library_cover/NBP1402_annotations_01_CoralNet.csv"))
PS18.ann <- read.csv(paste0(RS.dir,"PS18/PS18_4_annotation_library_cover_old/PS18_annotations_01_CoralNet.csv"))

ann.dat <- rbind(PS81.ann[,-7],PS96.ann[,-7],TAN0802.ann[,-7],TAN1802.ann[,-7],TAN1901.ann[,-7],NBP1402.ann,PS18.ann)

ann.dat.subset <- ann.dat[ann.dat$Name%in%img.names,-c(5:6)]
ann.dat.subset$Label <- factor(ann.dat.subset$Label)

ann.dat.cover <- data.frame(matrix(ncol=length(levels(ann.dat.subset$Label)),nrow=length(img.names)),row.names=img.names)
names(ann.dat.cover) <- levels(ann.dat.subset$Label)

for(i in 1:49){
  ann.dat.cover[i,] <- table(ann.dat.subset[ann.dat.subset$Name==img.names[i],4])
}
ann.dat.cover$total_points <- rowSums(ann.dat.cover)

write.csv(ann.dat.cover, file=paste0(s.dir,"Sample_images_cover.csv"))


## counts - filenames-by-species
PS81.ann <- read.csv(paste0(s.dir,"counts/PS81_full_annotation_report_202001.csv"))
PS96.ann <- read.csv(paste0(s.dir,"counts/PS96_full_annotation_report_202001.csv"))
TAN0802.ann <- read.csv(paste0(s.dir,"counts/TAN0802_full_annotation_report_202001.csv"))
TAN1802.ann <- read.csv(paste0(s.dir,"counts/TAN1802_full_annotation_report_202007.csv"))
TAN1901.ann <- read.csv(paste0(s.dir,"counts/TAN1901_full_annotation_report_202007.csv"))
NBP1402.ann <- read.csv(paste0(s.dir,"counts/2205-nbp1402.csv"))
PS18.ann <- read.csv(paste0(s.dir,"counts/2443-ps18.csv"))

ann.dat <- rbind(PS81.ann,PS96.ann,TAN0802.ann,TAN1802.ann,TAN1901.ann,NBP1402.ann,PS18.ann)
ann.dat.subset <- ann.dat[ann.dat$filename%in%img.names,c(9,3,4,14,15)]
ann.dat.subset$label_name <- factor(ann.dat.subset$label_name)

ann.dat.counts <- data.frame(matrix(ncol=length(levels(ann.dat.subset$label_name)),nrow=length(img.names)),row.names=img.names)
names(ann.dat.counts) <- levels(ann.dat.subset$label_name)

for(i in 1:49){
  ann.dat.counts[i,] <- table(ann.dat.subset[ann.dat.subset$filename==img.names[i],2])
}
ann.dat.counts$total_counts <- rowSums(ann.dat.counts)

write.csv(ann.dat.counts, file=paste0(s.dir,"Sample_images_counts.csv"))







```