---
title: "Read and clean up circumpolar image metadata"
author: "Jan Jansen"
date: "19/07/2021"
output: html_document
---

This script is to create the summary data:  
__"Circumpolar_DownwardImages_metadata.Rdata"__  
and standardised/consistent __metadata csv-files for each individual survey__  

TO DO FOR THE CSV-FILES:  
- transect-dates for CRS-data
- license-information for CRS, all TAN, all JR except 17003
- data-links for CRS, all TAN, all JR except 17003

This summary file contains the following objects:  

*  starting positions for each transect in each survey:
    + unprojected dataframes: __dat.start.list__
    + projected SpatialPoints: __polar.dat.start.list__
  
*  metadata:
    + original dataframes for each survey: __dat.list__
    + total transect lengths: __total.t.length.list__
    + clean data only including useful images: __dat.list.clean__
    + & clean projected SpatialPoints: __polar.dat.list.clean__
    + & clean total transect lengths: __total.t.length.list.clean__
  
*  images selected for annotation only:
    + unprojected dataframes: __dat.subset.list__
    + projected SpatialPoints: __polar.dat.subset.list__
  
*  individual surveys:
    + __polar.dat.XXX__  

*  polar stereographic projection used: __stereo__


```{r, warning=FALSE, message=FALSE}
library(raadtools)
library(readxl)
library(raster)
library(maptools)
library(lubridate)
library(plyr)
```

### Setting up the environment

```{r, warning=FALSE, message=FALSE}
## Jan's local machine:
sci.dir <-      "C:/Users/jjansen/Desktop/science/"
env.derived <-  paste0(sci.dir,"data_environmental/derived/")

## remote repository (DOESN'T WORK YET):
# env.dir <- "https://data.imas.utas.edu.au/data_transfer/admin/files/EnvironmentalData/"

## common paths (after "sci.dir")
tools.dir <-    paste0(sci.dir,"SouthernOceanBiodiversityMapping/Useful_Functions_Tools/")
ARC_Data.dir <- paste0(sci.dir,"SouthernOceanBiodiversityMapping/ARC_Data/")
image.dir <-      paste0(ARC_Data.dir,"prep_image/ReadIn_Circumpolar_DownwardImage_Data_And_Subset_Surveys/")

## polar stereographic projection:
stereo <- "+proj=stere +lat_0=-90 +lat_ts=-71 +lon_0=0 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs"
# stereo <- "+proj=stere +lat_0=-90 +lat_ts=-71 +lon_0=0 +k=1 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0"
```

```{r, warning=FALSE, message=FALSE}
#### load depth
r2 <- raster(paste0(env.derived,"Circumpolar_EnvData_500m_shelf_bathy_gebco_depth.grd"))

# #### load diatom data
# load(file=paste0(image.dir,"Circumpolar_Diatom_locations.Rdata"))

#### load coastline
load(paste0(env.derived,"Circumpolar_Coastline.Rdata"))
```

### load image area information
```{r}
R.dir <- "R:/IMAS/Antarctic_Seafloor/Clean_Data_For_Permanent_Storage/"
area.xlsx <- list.files(R.dir,pattern=".xlsx", full.names=TRUE)
area.raw <- ldply(area.xlsx, read_excel, skip=1)
```


### prepare data

```{r, warning=FALSE, message=FALSE}
#### Entire continental shelf
# par(mar=c(2,2,1,1))
# #png("diatom_overview.png",width=1250,height=1250)
# plot(r2, xlim=c(-3000000,3000000), ylim=c(-3000000,3000000),col=terrain.colors(99), main="Sample sites for sediment cores/grabs")
# plot(coast.proj, add=TRUE)
# scalebar(500000, type="bar", label=c("0","250","500"), below="km")
# points(polar.dat.diatom, pch=16, col="blue")
# legend("topright","Sample-sites",pch=16, col="blue")
#dev.off()

#### load image data
## Ross Sea
load(paste0(image.dir,"TAN0802_dat.Rdata"))
dat.0802 <- dat
total.t.length.v.0802 <- total.t.length.v
total.t.length.v.0802[c(13,16,17,28,32:34,45,46)] <- t.l[c(13,16,17,28,32:34,45,46)] #used distance to start for these transects, not distance along transect

load(paste0(image.dir,"TAN1802_dat.Rdata"))
dat.1802 <- dat
t.l.1802 <- t.l
total.t.length.v.1802 <- total.t.length.v

load(paste0(image.dir,"TAN1901_dat.Rdata"))
dat.1901 <- dat
t.l.1901 <- t.l
total.t.length.v.1901 <- total.t.length.v

load(file=paste0(image.dir,"NBP1502_dat.Rdata"))
dat.NBP1502 <- dat
total.t.length.v.NBP1502 <- total.t.length.v
t.l.NBP1502 <- t.l

## West Antarctic
load(file=paste0(image.dir,"PS81_dat_FINAL.Rdata"))
dat.PS81 <- PS81.dat
total.t.length.v.PS81 <- total.t.length.v

load(file=paste0(image.dir,"PS81_dat_shallow.Rdata"))
dat.PS81_shallow <- dat
total.t.length.v.PS81_shallow <- total.t.length.v

load(file=paste0(image.dir,"PS96_dat_FINAL.Rdata"))
dat.PS96 <- PS96.dat
total.t.length.v.PS96 <- total.t.length.v

load(file=paste0(image.dir,"PS118_dat.Rdata"))
dat.PS118 <- dat
total.t.length.v.PS118 <- total.t.length.v

## CRS
load(file=paste0(image.dir,"CRS_dat.Rdata"))
dat.WAP <- dat
total.t.length.v.WAP <- total.t.length.v

## East Antarctic
load(file=paste0(image.dir,"NBP1402_dat.Rdata"))
dat.NBP1402 <- dat
total.t.length.v.NBP1402 <- total.t.length.v
rm(dat,t.l,total.t.length.v)

## old Weddell Sea surveys
load(file=paste0(image.dir,"PS06_dat.Rdata"))
dat.PS06 <- dat

load(file=paste0(image.dir,"PS14_dat.Rdata"))
dat.PS14 <- dat
total.t.length.v.PS14 <- total.t.length.v

load(file=paste0(image.dir,"PS18_dat.Rdata"))
dat.PS18 <- dat
total.t.length.v.PS18 <- total.t.length.v

load(file=paste0(image.dir,"PS61_dat.Rdata"))
dat.PS61 <- dat
total.t.length.v.PS61 <- total.t.length.v

## Mertz
load(file=paste0(image.dir,"AA2011_dat.Rdata"))
dat.AA2011 <- dat

## SUCS
load(file=paste0(image.dir,"JR262_dat.Rdata"))
dat.JR262 <- dat

load(file=paste0(image.dir,"JR15005_dat.Rdata"))
dat.JR15005 <- dat
total.t.length.v.JR15005 <- total.t.length.v

load(file=paste0(image.dir,"JR17001_dat.Rdata"))
dat.JR17001 <- dat
total.t.length.v.JR17001 <- total.t.length.v

load(file=paste0(image.dir,"JR17003_dat.Rdata"))
dat.JR17003 <- dat
total.t.length.v.JR17003 <- total.t.length.v

## LMG1311
load(file=paste0(image.dir,"LMG1311_dat.Rdata"))
dat.LMG1311 <- dat
total.t.length.v.LMG1311 <- total.t.length.v
```

### Fixing image_select #1

```{r, warning=FALSE, message=FALSE}
##### PS81: something went wrong with the "image_select"-values when redoing the subsetting and replacing shitty images. Only happened to images from image.select>=50
dat.PS81$image.select[dat.PS81$Filename=="PS81_116-3_2013-01-26T12_51_09_5010.jpg"] <- 50
dat.PS81$image.select[dat.PS81$Filename=="PS81_116-3_2013-01-26T14_21_26_5190.jpg"] <- 53
dat.PS81$image.select[dat.PS81$Filename=="PS81_116-3_2013-01-26T13_58_10_5143.jpg"] <- 56
dat.PS81$image.select[dat.PS81$Filename=="PS81_116-3_2013-01-26T12_34_31_4977.jpg"] <- 62
dat.PS81$image.select[dat.PS81$Filename=="PS81_116-3_2013-01-26T13_09_49_5047.jpg"] <- 64
dat.PS81$image.select[dat.PS81$Filename=="PS81_116-3_2013-01-26T13_18_23_5064.jpg"] <- 65
dat.PS81$image.select[dat.PS81$Filename=="PS81_116-3_2013-01-26T12_28_30_4964.jpg"] <- 66

##### TAN1901: something went really wrong with saving the image subsetting.
#####   - Couldn't figure out why the difference between list of annotated images and stored TAN1901_dat.Rdata
#####   - Only one version of TAN1901_dat.Rdata exists and the reproducing the annotated subset using the R-code is not possible
#####   - Seems to be often off by 1 or 2 images, bu most transect are not affected

##
dat.1901$image.select[dat.1901$FileName=="tan1901_095_039.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_095_227.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_095_036.jpg"] <- 2
dat.1901$image.select[dat.1901$FileName=="tan1901_095_229.jpg"] <- 5
##
dat.1901$image.select[dat.1901$FileName=="tan1901_096_054.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_096_196.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_096_140.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_096_245.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_096_264.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_096_084.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_096_206.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_096_159.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_096_066.jpg"] <- 2
dat.1901$image.select[dat.1901$FileName=="tan1901_096_197.jpg"] <- 3
dat.1901$image.select[dat.1901$FileName=="tan1901_096_141.jpg"] <- 4
dat.1901$image.select[dat.1901$FileName=="tan1901_096_242.jpg"] <- 5
dat.1901$image.select[dat.1901$FileName=="tan1901_096_265.jpg"] <- 9
dat.1901$image.select[dat.1901$FileName=="tan1901_096_086.jpg"] <- 10
dat.1901$image.select[dat.1901$FileName=="tan1901_096_208.jpg"] <- 11
##
dat.1901$image.select[dat.1901$FileName=="tan1901_098_228.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_098_226.jpg"] <- 5
##
dat.1901$image.select[dat.1901$FileName=="tan1901_102_090.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_102_056.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_102_116.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_102_221.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_102_088.jpg"] <- 4
dat.1901$image.select[dat.1901$FileName=="tan1901_102_057.jpg"] <- 6
dat.1901$image.select[dat.1901$FileName=="tan1901_102_115.jpg"] <- 8
dat.1901$image.select[dat.1901$FileName=="tan1901_102_222.jpg"] <- 9
##
dat.1901$image.select[dat.1901$FileName=="tan1901_105_093.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_105_122.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_105_092.jpg"] <- 3
dat.1901$image.select[dat.1901$FileName=="tan1901_105_121.jpg"] <- 7
##
dat.1901$image.select[dat.1901$FileName=="tan1901_113_068.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_113_182.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_113_139.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_113_212.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_113_067.jpg"] <- 2
dat.1901$image.select[dat.1901$FileName=="tan1901_113_185.jpg"] <- 3
dat.1901$image.select[dat.1901$FileName=="tan1901_113_140.jpg"] <- 4
dat.1901$image.select[dat.1901$FileName=="tan1901_113_213.jpg"] <- 5
##
dat.1901$image.select[dat.1901$FileName=="tan1901_149_051.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_149_189.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_149_134.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_149_224.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_149_104.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_149_162.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_149_241.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_149_053.jpg"] <- 2
dat.1901$image.select[dat.1901$FileName=="tan1901_149_185.jpg"] <- 3
dat.1901$image.select[dat.1901$FileName=="tan1901_149_133.jpg"] <- 4
dat.1901$image.select[dat.1901$FileName=="tan1901_149_223.jpg"] <- 5
dat.1901$image.select[dat.1901$FileName=="tan1901_149_103.jpg"] <- 6
dat.1901$image.select[dat.1901$FileName=="tan1901_149_160.jpg"] <- 8
dat.1901$image.select[dat.1901$FileName=="tan1901_149_243.jpg"] <- 9
##
dat.1901$image.select[dat.1901$FileName=="tan1901_157_257.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_157_020.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_157_157.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_157_100.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_157_129.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_157_237.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_157_174.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_157_114.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_157_082.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_157_254.jpg"] <- 1
dat.1901$image.select[dat.1901$FileName=="tan1901_157_017.jpg"] <- 2
dat.1901$image.select[dat.1901$FileName=="tan1901_157_155.jpg"] <- 3
dat.1901$image.select[dat.1901$FileName=="tan1901_157_096.jpg"] <- 4
dat.1901$image.select[dat.1901$FileName=="tan1901_157_128.jpg"] <- 8
dat.1901$image.select[dat.1901$FileName=="tan1901_157_238.jpg"] <- 9
dat.1901$image.select[dat.1901$FileName=="tan1901_157_171.jpg"] <- 11
dat.1901$image.select[dat.1901$FileName=="tan1901_157_113.jpg"] <- 12
dat.1901$image.select[dat.1901$FileName=="tan1901_157_083.jpg"] <- 14
##
dat.1901$image.select[dat.1901$FileName=="tan1901_165_107.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_165_222.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_165_033.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_165_151.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_165_108.jpg"] <- 8
dat.1901$image.select[dat.1901$FileName=="tan1901_165_223.jpg"] <- 9
dat.1901$image.select[dat.1901$FileName=="tan1901_165_031.jpg"] <- 10
dat.1901$image.select[dat.1901$FileName=="tan1901_165_150.jpg"] <- 11
##
dat.1901$image.select[dat.1901$FileName=="tan1901_166_245.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_166_018.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_166_148.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_166_063.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_166_244.jpg"] <- 1
dat.1901$image.select[dat.1901$FileName=="tan1901_166_025.jpg"] <- 2
dat.1901$image.select[dat.1901$FileName=="tan1901_166_151.jpg"] <- 3
dat.1901$image.select[dat.1901$FileName=="tan1901_166_065.jpg"] <- 6
##
dat.1901$image.select[dat.1901$FileName=="tan1901_168_207.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_168_127.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_168_170.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_168_053.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_168_206.jpg"] <- 1
dat.1901$image.select[dat.1901$FileName=="tan1901_168_128.jpg"] <- 3
dat.1901$image.select[dat.1901$FileName=="tan1901_168_168.jpg"] <- 5
dat.1901$image.select[dat.1901$FileName=="tan1901_168_054.jpg"] <- 6
```

### add standardised filenames to data

```{r, warning=FALSE, message=FALSE}
## PS06,14,18,61,81,96:
dat <- dat.PS06
dat.PS06$Filename.standardised <- paste0(substr(dat$Filename,1,8),"_",sprintf("%04d",dat$image.select),"__",dat$Filename)
dat <- dat.PS14
dat.PS14$Filename.standardised <- paste0(substr(dat$Filename,1,8),"_",sprintf("%04d",dat$image.select),"__",dat$Filename)
dat <- dat.PS18
dat.PS18$Filename.standardised <- paste0(substr(dat$Filename,1,8),"_",sprintf("%04d",dat$image.select),"__",dat$Filename)
dat <- dat.PS61
dat.PS61$Filename.standardised <- paste0(substr(dat$Filename,1,8),"_",sprintf("%04d",dat$image.select),"__",dat$Filename)
dat <- dat.PS81
dat.PS81$Filename.standardised <- paste0(substr(dat$Filename,1,8),"_",sprintf("%04d",dat$image.select),"__",dat$Filename)
dat <- dat.PS81_shallow
dat.PS81_shallow$Filename.standardised <- paste0(substr(dat$Filename,1,8),"_",sprintf("%04d",dat$image.select),"__",dat$Filename)
dat <- dat.PS96
dat.PS96$Filename.standardised <- paste0(substr(dat$Filename,1,8),"_",sprintf("%04d",dat$image.select),"__",dat$Filename,".jpg")
## PS118:
dat <- dat.PS118
dat.PS118$Filename.standardised <- paste0("PS118_",dat$transectID.forfile,"_",sprintf("%04d",dat$image.select),"__",dat$Filename,".jpg")
## TAN0802 & 1901:
dat <- dat.0802
dat.0802$Filename.standardised <- paste0(substr(dat$FileName,1,12),sprintf("%04d",dat$image.select),"__",dat$FileName)
dat <- dat.1901
dat.1901$Filename.standardised <- paste0(substr(dat$FileName,1,12),sprintf("%04d",dat$image.select),"__",dat$FileName)
## TAN1802:
dat <- dat.1802
dat.1802$Filename.standardised <- paste0("TAN1802_",dat$transectID,"_",sprintf("%04d",dat$image.select),"__",dat$FileName)
## NBP1402
dat <- dat.NBP1402
dat.NBP1402$Filename.standardised <- paste0("NBP1402_",dat$transectID,"_",sprintf("%04d",dat$image.select),"__",dat$FileName)
## NBP1402
dat <- dat.NBP1502
dat.NBP1502$Filename.standardised <- paste0("NBP1502_",dat$transectID,"_",sprintf("%04d",dat$image.select),"__",dat$FileName)
## CRS:
dat <- dat.WAP
dat.WAP$Filename.standardised <- paste0("CRS_",dat$transectID,"_",sprintf("%04d",dat$image.select),"__",dat$Filename)
dat.WAP$Filename.standardised <- gsub(".tif",".jpg",dat.WAP$Filename.standardised)
## AA2011
dat <- dat.AA2011
dat.AA2011$Filename.standardised <- paste0("AA2011_",dat$transectID_folder,"_",sprintf("%04d",dat$image.select),"__",dat$Filename)
## SUCS
dat <- dat.JR262
dat.JR262$Filename.standardised <- paste0("JR262_",dat$transectID,"_",sprintf("%04d",dat$image.select),"__",dat$filename)
dat <- dat.JR15005
dat.JR15005$Filename.standardised <- paste0("JR15005_",dat$transectID,"_",sprintf("%04d",dat$image.select),"__",dat$filename)
dat <- dat.JR17001
dat.JR17001$Filename.standardised <- paste0("JR17001_",dat$transectID,"_",sprintf("%04d",dat$image.select),"__",dat$filename)
dat.JR17001$Filename.standardised <- gsub(" ","",dat.JR17001$Filename.standardised)
dat <- dat.JR17003
dat.JR17003$Filename.standardised <- paste0("JR17003_",dat$transectID,"_",sprintf("%04d",dat$image.select),"__",dat$filename)
## LMG1311
dat <- dat.LMG1311
dat.LMG1311$Filename.standardised <- paste0("LMG1311_",dat$transectID,"_",sprintf("%04d",dat$image.select),"__",dat$filename)

rm(dat)
```

### fixing image_select #2
set.seed() didn't run as expected across all transects. Some images might have been moved to the "bad quality" folder and then the subset rerun again. In short, some transect metadata show different images selected than we actually annotated.  
This chunk is to identify those transects and images and correct the image-names with the correct image number from the subset.

```{r}
'%!in%' <- function(x,y)!('%in%'(x,y))
r.drive <- "R:/IMAS/Antarctic_Seafloor/Clean_Data_For_Permanent_Storage/"

## which survey to do: CRS
dat <- dat.WAP

## first, read in the names of the images that we annotated:
img.names.wap <- list.files(paste0(r.drive,"CRS/CRS_3_colourcorrected_images_for_annotation"),pattern=".jpg")
## and here are the names in the metadata
dat.WAP$Filename.standardised
## compare the two
img.names.wap[which(img.names.wap%!in%dat.WAP$Filename.standardised)]
## transects with issues:
#c(1207,1217,1276,1282,1289,1300)

#########################################
##### 1207:
ann.sel <- grep("CRS_1207",img.names.wap)
img.names.wap[ann.sel]
##### all images with a number greater than 10 are wrong
## select the images:
sel <- which(dat.WAP$transectID==1207&dat.WAP$image.select>10)
## set image_select to 999 for these images
dat$image.select[sel] <- 999
## change it in the filename as well
dat$Filename.standardised[sel] <- paste0(substr(dat.WAP$Filename.standardised[sel],1,9),
                                             sprintf("%04d",dat.WAP$image.select[sel]),
                                             substr(dat.WAP$Filename.standardised[sel],14,50))
## now set the correct image_select for the annotated images
## first check image_select numbers from annotated images
substr(img.names.wap[ann.sel],10,13)
## image names for missing/wrong numbers
repl.sel <- substr(img.names.wap[ann.sel],16,50)[11:15]
## under which idx can we find these images in dat
repl.sel2 <- match(repl.sel,substr(dat.WAP$Filename.standardised[sel],16,50))
## correct the filenames and image_select values for the annotated images
dat$Filename.standardised[sel][repl.sel2] <- paste0(substr(dat.WAP$Filename.standardised[sel][repl.sel2],1,9),
                                             substr(img.names.wap[ann.sel],10,13)[11:15],
                                             substr(dat.WAP$Filename.standardised[sel][repl.sel2],14,50))
## test if everything correct:
img.names.wap[which(img.names.wap%!in%dat$Filename.standardised)]

################################################
##### 1217:
ann.sel <- grep("CRS_1217",img.names.wap)
img.names.wap[ann.sel]
##### images with numbers 10 to ... are wrong
## select the images:
sel <- which(dat.WAP$transectID==1217&dat.WAP$image.select>=10)
## set image_select to 999 for these images
dat$image.select[sel] <- 999
## change it in the filename as well
dat$Filename.standardised[sel] <- paste0(substr(dat.WAP$Filename.standardised[sel],1,9),
                                             sprintf("%04d",dat.WAP$image.select[sel]),
                                             substr(dat.WAP$Filename.standardised[sel],14,50))
## now set the correct image_select for the annotated images
## first check image_select numbers from annotated images
substr(img.names.wap[ann.sel],10,13)
## image names for missing/wrong numbers
repl.sel <- substr(img.names.wap[ann.sel],16,50)[10:12]
## under which idx can we find these images in dat
repl.sel2 <- match(repl.sel,substr(dat.WAP$Filename.standardised[sel],16,50))
## correct the filenames and image_select values for the annotated images
dat$Filename.standardised[sel][repl.sel2] <- paste0(substr(dat.WAP$Filename.standardised[sel][repl.sel2],1,9),
                                             substr(img.names.wap[ann.sel],10,13)[10:12],
                                             substr(dat.WAP$Filename.standardised[sel][repl.sel2],14,50))
## test if everything correct:
img.names.wap[which(img.names.wap%!in%dat$Filename.standardised)]

################################################
##### 1276:
ann.sel <- grep("CRS_1276",img.names.wap)
img.names.wap[ann.sel]
##### images with numbers 6 to ... are wrong
## select the images:
sel <- which(dat.WAP$transectID==1276&dat.WAP$image.select>=6)
## set image_select to 999 for these images
dat$image.select[sel] <- 999
## change it in the filename as well
dat$Filename.standardised[sel] <- paste0(substr(dat.WAP$Filename.standardised[sel],1,9),
                                             sprintf("%04d",dat.WAP$image.select[sel]),
                                             substr(dat.WAP$Filename.standardised[sel],14,50))
## now set the correct image_select for the annotated images
## first check image_select numbers from annotated images
substr(img.names.wap[ann.sel],10,13)
## image names for missing/wrong numbers
repl.sel <- substr(img.names.wap[ann.sel],16,50)[6:9]
## under which idx can we find these images in dat
repl.sel2 <- match(repl.sel,substr(dat.WAP$Filename.standardised[sel],16,50))
## correct the filenames and image_select values for the annotated images
dat$Filename.standardised[sel][repl.sel2] <- paste0(substr(dat.WAP$Filename.standardised[sel][repl.sel2],1,9),
                                             substr(img.names.wap[ann.sel],10,13)[6:9],
                                             substr(dat.WAP$Filename.standardised[sel][repl.sel2],14,50))
## test if everything correct:
img.names.wap[which(img.names.wap%!in%dat$Filename.standardised)]

################################################
##### 1282:
ann.sel <- grep("CRS_1282",img.names.wap)
img.names.wap[ann.sel]
##### images with numbers 3 to ... are wrong
## select the images:
sel <- which(dat.WAP$transectID==1282&dat.WAP$image.select>=3)
## set image_select to 999 for these images
dat$image.select[sel] <- 999
## change it in the filename as well
dat$Filename.standardised[sel] <- paste0(substr(dat.WAP$Filename.standardised[sel],1,9),
                                             sprintf("%04d",dat.WAP$image.select[sel]),
                                             substr(dat.WAP$Filename.standardised[sel],14,50))
## now set the correct image_select for the annotated images
## first check image_select numbers from annotated images
substr(img.names.wap[ann.sel],10,13)
## image names for missing/wrong numbers
repl.sel <- substr(img.names.wap[ann.sel],16,50)[3:7]
## under which idx can we find these images in dat
repl.sel2 <- match(repl.sel,substr(dat.WAP$Filename.standardised[sel],16,50))
## correct the filenames and image_select values for the annotated images
dat$Filename.standardised[sel][repl.sel2] <- paste0(substr(dat.WAP$Filename.standardised[sel][repl.sel2],1,9),
                                             substr(img.names.wap[ann.sel],10,13)[3:7],
                                             substr(dat.WAP$Filename.standardised[sel][repl.sel2],14,50))
## test if everything correct:
img.names.wap[which(img.names.wap%!in%dat$Filename.standardised)]

################################################
##### 1289:
ann.sel <- grep("CRS_1289",img.names.wap)
img.names.wap[ann.sel]
##### images with numbers 2 to ... are wrong
## select the images:
sel <- which(dat.WAP$transectID==1289&dat.WAP$image.select>=2)
## set image_select to 999 for these images
dat$image.select[sel] <- 999
## change it in the filename as well
dat$Filename.standardised[sel] <- paste0(substr(dat.WAP$Filename.standardised[sel],1,9),
                                             sprintf("%04d",dat.WAP$image.select[sel]),
                                             substr(dat.WAP$Filename.standardised[sel],14,50))
## now set the correct image_select for the annotated images
## first check image_select numbers from annotated images
substr(img.names.wap[ann.sel],10,13)
## image names for missing/wrong numbers
repl.sel <- substr(img.names.wap[ann.sel],16,50)[2:6]
## under which idx can we find these images in dat
repl.sel2 <- match(repl.sel,substr(dat.WAP$Filename.standardised[sel],16,50))
## correct the filenames and image_select values for the annotated images
dat$Filename.standardised[sel][repl.sel2] <- paste0(substr(dat.WAP$Filename.standardised[sel][repl.sel2],1,9),
                                             substr(img.names.wap[ann.sel],10,13)[2:6],
                                             substr(dat.WAP$Filename.standardised[sel][repl.sel2],14,50))
## test if everything correct:
img.names.wap[which(img.names.wap%!in%dat$Filename.standardised)]

################################################
##### 1300:
ann.sel <- grep("CRS_1300",img.names.wap)
img.names.wap[ann.sel]
##### images with numbers 8 to ... are wrong
## select the images:
sel <- which(dat.WAP$transectID==1300&dat.WAP$image.select>=8)
## set image_select to 999 for these images
dat$image.select[sel] <- 999
## change it in the filename as well
dat$Filename.standardised[sel] <- paste0(substr(dat.WAP$Filename.standardised[sel],1,9),
                                             sprintf("%04d",dat.WAP$image.select[sel]),
                                             substr(dat.WAP$Filename.standardised[sel],14,50))
## now set the correct image_select for the annotated images
## first check image_select numbers from annotated images
substr(img.names.wap[ann.sel],10,13)
## image names for missing/wrong numbers
repl.sel <- substr(img.names.wap[ann.sel],16,50)[8:18]
## under which idx can we find these images in dat
repl.sel2 <- match(repl.sel,substr(dat.WAP$Filename.standardised[sel],16,50))
## correct the filenames and image_select values for the annotated images
dat$Filename.standardised[sel][repl.sel2] <- paste0(substr(dat.WAP$Filename.standardised[sel][repl.sel2],1,9),
                                             substr(img.names.wap[ann.sel],10,13)[8:18],
                                             substr(dat.WAP$Filename.standardised[sel][repl.sel2],14,50))
## test if everything correct:
img.names.wap[which(img.names.wap%!in%dat$Filename.standardised)]

###############################################3
dat.WAP <- dat
```

```{r}
## check other surveys:
img.names <- list.files(paste0(r.drive,"TAN0802/TAN0802_3_colourcorrected_images_for_annotation"),pattern=".jpg")
img.names[which(img.names%!in%dat.0802$Filename.standardised)]

img.names <- list.files(paste0(r.drive,"TAN1802/TAN1802_3_colourcorrected_images_for_annotation"),pattern=".jpg")
img.names[which(img.names%!in%dat.1802$Filename.standardised)]

img.names <- list.files(paste0(r.drive,"TAN1901/TAN1901_3_colourcorrected_images_for_annotation"),pattern=".jpg")
img.names[which(img.names%!in%dat.1901$Filename.standardised)]

img.names <- list.files(paste0(r.drive,"AA2011/AA2011_3_colourcorrected_images_for_annotation"),pattern=".JPG")
img.names[which(img.names%!in%dat.AA2011$Filename.standardised)]

img.names <- list.files(paste0(r.drive,"JR15005/JR15005_3_cropped_and_colourcorrected_images_for_annotation"),pattern=".png")
img.names[which(img.names%!in%dat.JR15005$Filename.standardised)]

img.names <- list.files(paste0(r.drive,"JR17001/JR17001_3_cropped_and_colourcorrected_images_for_annotation"),pattern=".png")
img.names[which(img.names%!in%dat.JR17001$Filename.standardised)]

img.names <- list.files(paste0(r.drive,"JR17003/JR17003_3_cropped_and_colourcorrected_images_for_annotation"),pattern=".png")
img.names[which(img.names%!in%dat.JR17003$Filename.standardised)]

img.names <- list.files(paste0(r.drive,"JR262/JR262_3_cropped_and_colourcorrected_images_for_annotation"),pattern=".png")
img.names[which(img.names%!in%dat.JR262$Filename.standardised)]

img.names <- list.files(paste0(r.drive,"LMG1311/LMG1311_3_colourcorrected_images_for_annotation"),pattern=".JPG")
img.names[which(img.names%!in%dat.LMG1311$Filename.standardised)]

#########################################
## which survey to do: LMG1311
dat <- dat.LMG1311

## first, read in the names of the images that we annotated:
img.names.lmg <- list.files(paste0(r.drive,"LMG1311/LMG1311_3_colourcorrected_images_for_annotation"),pattern=".JPG")
## and here are the names in the metadata
dat$Filename.standardised
## compare the two
img.names.lmg[which(img.names.lmg%!in%dat$Filename.standardised)]
## transects with issues: 3
#####
ann.sel <- grep("LMG1311_3",img.names.lmg)
img.names.lmg[ann.sel]
##### all images with a number greater than 34 are wrong
## select the images:
sel <- which(dat.LMG1311$transectID==3&dat.LMG1311$image.select>34)
## set image_select to 999 for these images
dat$image.select[sel] <- 999
## change it in the filename as well
dat$Filename.standardised[sel] <- paste0(substr(dat.LMG1311$Filename.standardised[sel],1,10),
                                             sprintf("%04d",dat.LMG1311$image.select[sel]),
                                             substr(dat.LMG1311$Filename.standardised[sel],15,50))
## now set the correct image_select for the annotated images
## first check image_select numbers from annotated images
substr(img.names.lmg[ann.sel],11,14)
## image names for missing/wrong numbers
repl.sel <- substr(img.names.lmg[ann.sel],15,50)[35:43]
## under which idx can we find these images in dat
repl.sel2 <- match(repl.sel,substr(dat.LMG1311$Filename.standardised[sel],15,50))
## correct the filenames and image_select values for the annotated images
dat$Filename.standardised[sel][repl.sel2] <- paste0(substr(dat.LMG1311$Filename.standardised[sel][repl.sel2],1,10),
                                             substr(img.names.lmg[ann.sel],11,14)[35:43],
                                             substr(dat.LMG1311$Filename.standardised[sel][repl.sel2],15,50))
## test if everything correct:
img.names.lmg[which(img.names.lmg%!in%dat$Filename.standardised)]
dat.LMG1311 <- dat
######################################

img.names <- list.files(paste0(r.drive,"NBP1402/NBP1402_3_colourcorrected_images_for_annotation"),pattern=".jpg")
img.names[which(img.names%!in%dat.NBP1402$Filename.standardised)]

img.names <- list.files(paste0(r.drive,"NBP1502/NBP1502_3_colourcorrected_images_for_annotation"),pattern=".jpg")
img.names[which(img.names%!in%dat.NBP1502$Filename.standardised)]

img.names <- list.files(paste0(r.drive,"PS06/PS06_3_colourcorrected_images_for_annotation"),pattern=".jpg")
img.names[which(img.names%!in%dat.PS06$Filename.standardised)]

img.names <- list.files(paste0(r.drive,"PS14/PS14_3_colourcorrected_images_for_annotation"),pattern=".jpg")
img.names[which(img.names%!in%dat.PS14$Filename.standardised)]

img.names <- list.files(paste0(r.drive,"PS18/PS18_3_colourcorrected_images_for_annotation"),pattern=".jpg")
img.names[which(img.names%!in%dat.PS18$Filename.standardised)]

img.names <- list.files(paste0(r.drive,"PS61/PS61_3_colourcorrected_images_for_annotation"),pattern=".jpg")
img.names[which(img.names%!in%dat.PS61$Filename.standardised)]

img.names <- list.files(paste0(r.drive,"PS81/PS81_3_cropped_and_colourcorrected_images_for_annotation"),pattern=".jpg")
img.names[which(img.names%!in%dat.PS81$Filename.standardised)]

img.names <- list.files(paste0(r.drive,"PS81_shallow/PS81_shallow_3_cropped_and_colourcorrected_images_for_annotation"),pattern=".jpg")
img.names[which(img.names%!in%dat.PS81_shallow$Filename.standardised)]

img.names <- list.files(paste0(r.drive,"PS96/PS96_3_cropped_images_for_annotation_NoColourCorrectionNeeded"),pattern=".jpg")
img.names[which(img.names%!in%dat.PS96$Filename.standardised)]

img.names <- list.files(paste0(r.drive,"PS118/PS118_3_cropped_and_colourcorrected_images_for_annotation"),pattern=".jpg")
img.names[which(img.names%!in%dat.PS118$Filename.standardised)]
```

### reproject biological data

```{r, warning=FALSE, message=FALSE}
##
spatial.dat <- data.frame(cbind(dat.0802$GPS_lon, dat.0802$GPS_lat))
names(spatial.dat) <- c("Longitude","Latitude")
coordinates(spatial.dat) <- c("Longitude","Latitude")
proj4string(spatial.dat) <- CRS("+proj=longlat +datum=WGS84")
polar.dat.0802 <- spTransform(spatial.dat, CRS(stereo))
##
spatial.dat <- data.frame(cbind(dat.1802$GPS_lon, dat.1802$GPS_lat))
names(spatial.dat) <- c("Longitude","Latitude")
coordinates(spatial.dat) <- c("Longitude","Latitude")
proj4string(spatial.dat) <- CRS("+proj=longlat +datum=WGS84")
polar.dat.1802 <- spTransform(spatial.dat, CRS(stereo))
##
spatial.dat <- data.frame(cbind(dat.1901$GPS_lon, dat.1901$GPS_lat))
names(spatial.dat) <- c("Longitude","Latitude")
coordinates(spatial.dat) <- c("Longitude","Latitude")
proj4string(spatial.dat) <- CRS("+proj=longlat +datum=WGS84")
polar.dat.1901 <- spTransform(spatial.dat, CRS(stereo))
##
spatial.dat <- data.frame(cbind(dat.PS81$Longitude, dat.PS81$Latitude))
names(spatial.dat) <- c("Longitude","Latitude")
coordinates(spatial.dat) <- c("Longitude","Latitude")
proj4string(spatial.dat) <- CRS("+proj=longlat +datum=WGS84")
polar.dat.PS81 <- spTransform(spatial.dat, CRS(stereo))
##
spatial.dat <- data.frame(cbind(dat.PS81_shallow$Longitude, dat.PS81_shallow$Latitude))
names(spatial.dat) <- c("Longitude","Latitude")
coordinates(spatial.dat) <- c("Longitude","Latitude")
proj4string(spatial.dat) <- CRS("+proj=longlat +datum=WGS84")
polar.dat.PS81_shallow <- spTransform(spatial.dat, CRS(stereo))
##
spatial.dat <- data.frame(cbind(dat.PS96$Longitude, dat.PS96$Latitude))
names(spatial.dat) <- c("Longitude","Latitude")
coordinates(spatial.dat) <- c("Longitude","Latitude")
proj4string(spatial.dat) <- CRS("+proj=longlat +datum=WGS84")
polar.dat.PS96 <- spTransform(spatial.dat, CRS(stereo))
##
spatial.dat <- data.frame(cbind(dat.PS118$Longitude, dat.PS118$Latitude))
names(spatial.dat) <- c("Longitude","Latitude")
coordinates(spatial.dat) <- c("Longitude","Latitude")
proj4string(spatial.dat) <- CRS("+proj=longlat +datum=WGS84")
polar.dat.PS118 <- spTransform(spatial.dat, CRS(stereo))
##
spatial.dat <- data.frame(cbind(dat.WAP$lon, dat.WAP$lat))
names(spatial.dat) <- c("Longitude","Latitude")
coordinates(spatial.dat) <- c("Longitude","Latitude")
proj4string(spatial.dat) <- CRS("+proj=longlat +datum=WGS84")
polar.dat.WAP <- spTransform(spatial.dat, CRS(stereo))
##
spatial.dat <- data.frame(cbind(dat.NBP1402$GPS_lon, dat.NBP1402$GPS_lat))
names(spatial.dat) <- c("Longitude","Latitude")
coordinates(spatial.dat) <- c("Longitude","Latitude")
proj4string(spatial.dat) <- CRS("+proj=longlat +datum=WGS84")
polar.dat.NBP1402 <- spTransform(spatial.dat, CRS(stereo))
##
spatial.dat <- data.frame(cbind(dat.NBP1502$GPS_lon, dat.NBP1502$GPS_lat))
names(spatial.dat) <- c("Longitude","Latitude")
coordinates(spatial.dat) <- c("Longitude","Latitude")
proj4string(spatial.dat) <- CRS("+proj=longlat +datum=WGS84")
polar.dat.NBP1502 <- spTransform(spatial.dat, CRS(stereo))
##
spatial.dat <- data.frame(cbind(dat.PS06$lon, dat.PS06$lat))
names(spatial.dat) <- c("Longitude","Latitude")
coordinates(spatial.dat) <- c("Longitude","Latitude")
proj4string(spatial.dat) <- CRS("+proj=longlat +datum=WGS84")
polar.dat.PS06 <- spTransform(spatial.dat, CRS(stereo))
##
spatial.dat <- data.frame(cbind(dat.PS14$lon, dat.PS14$lat))
names(spatial.dat) <- c("Longitude","Latitude")
coordinates(spatial.dat) <- c("Longitude","Latitude")
proj4string(spatial.dat) <- CRS("+proj=longlat +datum=WGS84")
polar.dat.PS14 <- spTransform(spatial.dat, CRS(stereo))
##
spatial.dat <- data.frame(cbind(dat.PS18$lon, dat.PS18$lat))
names(spatial.dat) <- c("Longitude","Latitude")
coordinates(spatial.dat) <- c("Longitude","Latitude")
proj4string(spatial.dat) <- CRS("+proj=longlat +datum=WGS84")
polar.dat.PS18 <- spTransform(spatial.dat, CRS(stereo))
##
spatial.dat <- data.frame(cbind(dat.PS61$lon, dat.PS61$lat))
names(spatial.dat) <- c("Longitude","Latitude")
coordinates(spatial.dat) <- c("Longitude","Latitude")
proj4string(spatial.dat) <- CRS("+proj=longlat +datum=WGS84")
polar.dat.PS61 <- spTransform(spatial.dat, CRS(stereo))
##
spatial.dat <- data.frame(cbind(dat.AA2011$lon, dat.AA2011$lat))
names(spatial.dat) <- c("Longitude","Latitude")
coordinates(spatial.dat) <- c("Longitude","Latitude")
proj4string(spatial.dat) <- CRS("+proj=longlat +datum=WGS84")
polar.dat.AA2011 <- spTransform(spatial.dat, CRS(stereo))
##
spatial.dat <- data.frame(cbind(dat.JR262$lon, dat.JR262$lat))
names(spatial.dat) <- c("Longitude","Latitude")
coordinates(spatial.dat) <- c("Longitude","Latitude")
proj4string(spatial.dat) <- CRS("+proj=longlat +datum=WGS84")
polar.dat.JR262 <- spTransform(spatial.dat, CRS(stereo))
##
spatial.dat <- data.frame(cbind(dat.JR15005$lon, dat.JR15005$lat))
names(spatial.dat) <- c("Longitude","Latitude")
coordinates(spatial.dat) <- c("Longitude","Latitude")
proj4string(spatial.dat) <- CRS("+proj=longlat +datum=WGS84")
polar.dat.JR15005 <- spTransform(spatial.dat, CRS(stereo))
##
spatial.dat <- data.frame(cbind(dat.JR17001$lon, dat.JR17001$lat))
names(spatial.dat) <- c("Longitude","Latitude")
coordinates(spatial.dat) <- c("Longitude","Latitude")
proj4string(spatial.dat) <- CRS("+proj=longlat +datum=WGS84")
polar.dat.JR17001 <- spTransform(spatial.dat, CRS(stereo))
##
spatial.dat <- data.frame(cbind(dat.JR17003$lon, dat.JR17003$lat))
names(spatial.dat) <- c("Longitude","Latitude")
coordinates(spatial.dat) <- c("Longitude","Latitude")
proj4string(spatial.dat) <- CRS("+proj=longlat +datum=WGS84")
polar.dat.JR17003 <- spTransform(spatial.dat, CRS(stereo))
##
spatial.dat <- data.frame(cbind(dat.LMG1311$lon, dat.LMG1311$lat))
names(spatial.dat) <- c("Longitude","Latitude")
coordinates(spatial.dat) <- c("Longitude","Latitude")
proj4string(spatial.dat) <- CRS("+proj=longlat +datum=WGS84")
polar.dat.LMG1311 <- spTransform(spatial.dat, CRS(stereo))

####
# ## need start and end positions for PS14 and PS18
# spatial.dat <- data.frame(cbind(dat.PS14$Longitude_Start, dat.PS14$Latitude_Start))
# names(spatial.dat) <- c("Longitude","Latitude")
# coordinates(spatial.dat) <- c("Longitude","Latitude")
# proj4string(spatial.dat) <- CRS("+proj=longlat +datum=WGS84")
# polar.dat.PS14.start <- spTransform(spatial.dat, CRS(stereo))
# ##
# spatial.dat <- data.frame(cbind(dat.PS14$Longitude_End, dat.PS14$Latitude_End))
# names(spatial.dat) <- c("Longitude","Latitude")
# coordinates(spatial.dat) <- c("Longitude","Latitude")
# proj4string(spatial.dat) <- CRS("+proj=longlat +datum=WGS84")
# polar.dat.PS14.end <- spTransform(spatial.dat, CRS(stereo))
# ##
# spatial.dat <- data.frame(cbind(dat.PS18$Longitude_Start, dat.PS18$Latitude_Start))
# names(spatial.dat) <- c("Longitude","Latitude")
# spatial.dat <- spatial.dat[-which(is.na(rowSums(spatial.dat))),]
# coordinates(spatial.dat) <- c("Longitude","Latitude")
# proj4string(spatial.dat) <- CRS("+proj=longlat +datum=WGS84")
# polar.dat.PS18.start <- spTransform(spatial.dat, CRS(stereo))
# ##
# spatial.dat <- data.frame(cbind(dat.PS18$Longitude_End, dat.PS18$Latitude_End))
# names(spatial.dat) <- c("Longitude","Latitude")
# spatial.dat <- spatial.dat[-which(is.na(rowSums(spatial.dat))),]
# coordinates(spatial.dat) <- c("Longitude","Latitude")
# proj4string(spatial.dat) <- CRS("+proj=longlat +datum=WGS84")
# polar.dat.PS18.end <- spTransform(spatial.dat, CRS(stereo))
```

### create single files/lists for all surveys together

```{r, warning=FALSE, message=FALSE}
## start-points and ID for all transects:
dat.start.list <- list()
dat.start.list$WAP <- dat.WAP[!duplicated(dat.WAP$transectID),]
dat.start.list$PS06 <- dat.PS06[!duplicated(dat.PS06$transectID),]
dat.start.list$PS14 <- dat.PS14[!duplicated(dat.PS14$transectID),]
dat.start.list$PS18 <- dat.PS18[!duplicated(dat.PS18$transectID),]
dat.start.list$PS61 <- dat.PS61[!duplicated(dat.PS61$transectID),]
dat.start.list$PS81 <- dat.PS81[!duplicated(dat.PS81$transectID),]
dat.start.list$PS81_shallow <- dat.PS81_shallow[!duplicated(dat.PS81_shallow$transectID),]
dat.start.list$PS96 <- dat.PS96[!duplicated(dat.PS96$transectID),]
dat.start.list$PS118 <- dat.PS118[!duplicated(dat.PS118$transectID),]
dat.start.list$TAN0802 <- dat.0802[!duplicated(dat.0802$transectID),]
dat.start.list$TAN1802 <- dat.1802[!duplicated(dat.1802$transectID),]
dat.start.list$TAN1901 <- dat.1901[!duplicated(dat.1901$transectID),]
dat.start.list$NBP1402 <- dat.NBP1402[!duplicated(dat.NBP1402$transectID),]
dat.start.list$NBP1502 <- dat.NBP1502[!duplicated(dat.NBP1502$transectID),]
dat.start.list$AA2011 <- dat.AA2011[!duplicated(dat.AA2011$transectID),]
dat.start.list$JR262 <- dat.JR262[!duplicated(dat.JR262$transectID),]
dat.start.list$JR15005 <- dat.JR15005[!duplicated(dat.JR15005$transectID),]
dat.start.list$JR17001 <- dat.JR17001[!duplicated(dat.JR17001$transectID),]
dat.start.list$JR17003 <- dat.JR17003[!duplicated(dat.JR17003$transectID),]
dat.start.list$LMG1311 <- dat.LMG1311[!duplicated(dat.LMG1311$transectID),]

polar.dat.start.list <- list()
polar.dat.start.list$WAP <- polar.dat.WAP[!duplicated(dat.WAP$transectID),]
polar.dat.start.list$PS06 <- polar.dat.PS06[!duplicated(dat.PS06$transectID),]
polar.dat.start.list$PS14 <- polar.dat.PS14[!duplicated(dat.PS14$transectID),]
polar.dat.start.list$PS18 <- polar.dat.PS18[!duplicated(dat.PS18$transectID),]
polar.dat.start.list$PS61 <- polar.dat.PS61[!duplicated(dat.PS61$transectID),]
polar.dat.start.list$PS81 <- polar.dat.PS81[!duplicated(dat.PS81$transectID),]
polar.dat.start.list$PS81_shallow <- polar.dat.PS81_shallow[!duplicated(dat.PS81_shallow$transectID),]
polar.dat.start.list$PS96 <- polar.dat.PS96[!duplicated(dat.PS96$transectID),]
polar.dat.start.list$PS118 <- polar.dat.PS118[!duplicated(dat.PS118$transectID),]
polar.dat.start.list$TAN0802 <- polar.dat.0802[!duplicated(dat.0802$transectID),]
polar.dat.start.list$TAN1802 <- polar.dat.1802[!duplicated(dat.1802$transectID),]
polar.dat.start.list$TAN1901 <- polar.dat.1901[!duplicated(dat.1901$transectID),]
polar.dat.start.list$NBP1402 <- polar.dat.NBP1402[!duplicated(dat.NBP1402$transectID),]
polar.dat.start.list$NBP1502 <- polar.dat.NBP1502[!duplicated(dat.NBP1502$transectID),]
polar.dat.start.list$AA2011 <- polar.dat.AA2011[!duplicated(dat.AA2011$transectID),]
polar.dat.start.list$JR262 <- polar.dat.JR262[!duplicated(dat.JR262$transectID),]
polar.dat.start.list$JR15005 <- polar.dat.JR15005[!duplicated(dat.JR15005$transectID),]
polar.dat.start.list$JR17001 <- polar.dat.JR17001[!duplicated(dat.JR17001$transectID),]
polar.dat.start.list$JR17003 <- polar.dat.JR17003[!duplicated(dat.JR17003$transectID),]
polar.dat.start.list$LMG1311 <- polar.dat.LMG1311[!duplicated(dat.LMG1311$transectID),]

dat.list <- list()
dat.list$WAP <- dat.WAP
dat.list$PS06 <- dat.PS06
dat.list$PS14 <- dat.PS14
dat.list$PS18 <- dat.PS18
dat.list$PS61 <- dat.PS61
dat.list$PS81 <- dat.PS81
dat.list$PS81_shallow <- dat.PS81_shallow
dat.list$PS96 <- dat.PS96
dat.list$PS118 <- dat.PS118
dat.list$TAN0802 <- dat.0802
dat.list$TAN1802 <- dat.1802
dat.list$TAN1901 <- dat.1901
dat.list$NBP1402 <- dat.NBP1402
dat.list$NBP1502 <- dat.NBP1502
dat.list$AA2011 <- dat.AA2011
dat.list$JR262 <- dat.JR262
dat.list$JR15005 <- dat.JR15005
dat.list$JR17001 <- dat.JR17001
dat.list$JR17003 <- dat.JR17003
dat.list$LMG1311 <- dat.LMG1311

total.t.length.list <- list()
total.t.length.list$WAP <- total.t.length.v.WAP
total.t.length.list$PS06 <- NA
total.t.length.list$PS14 <- total.t.length.v.PS14
total.t.length.list$PS18 <- total.t.length.v.PS18
total.t.length.list$PS61 <- total.t.length.v.PS61
total.t.length.list$PS81 <- total.t.length.v.PS81
total.t.length.list$PS81_shallow <- total.t.length.v.PS81_shallow
total.t.length.list$PS96 <- total.t.length.v.PS96
total.t.length.list$PS118 <- total.t.length.v.PS118
total.t.length.list$TAN0802 <- total.t.length.v.0802
total.t.length.list$TAN1802 <- total.t.length.v.1802
total.t.length.list$TAN1901 <- total.t.length.v.1901
total.t.length.list$NBP1402 <- total.t.length.v.NBP1402
total.t.length.list$NBP1502 <- total.t.length.v.NBP1502
total.t.length.list$AA2011 <- NA
total.t.length.list$JR262 <- NA
total.t.length.list$JR15005 <- total.t.length.v.JR15005
total.t.length.list$JR17001 <- total.t.length.v.JR17001
total.t.length.list$JR17003 <- total.t.length.v.JR17003
total.t.length.list$LMG1311 <- total.t.length.v.LMG1311

########## annotated subset only
## following transects were disregarded:
## CRS 1103 & CRS 1325              # not yet removed
## PS06 288, 301-303, 307,309-311   # already removed
## PS81 159                         # not yet removed
## PS118 38                         # already removed
## TAN0802 181-310                  # not yet removed
## TAN1802 160-213                  # not yet removed
## TAN1901 209                      # not yet removed
## NBP1402 just a single image that is misplaced # not yet removed
## AA2011 CTD53,CTD55,CTD57-58,CTD61-64,CTD85,TD90,CTD93,CTD98,CTD107,CTD115,CTD138 # already removed
polar.dat.list.clean <- list(polar.dat.WAP[-which(dat.list$WAP$transectID%in%c(1103,1325))],
                       polar.dat.PS06,
                       polar.dat.PS14,
                       polar.dat.PS18,
                       polar.dat.PS61,
                       polar.dat.PS81[-which(dat.list$PS81$transectID=="PS81_159")],
                       polar.dat.PS81_shallow,
                       polar.dat.PS96,
                       polar.dat.PS118,
                       polar.dat.0802[-which(as.numeric(as.character(dat.list$TAN0802$transectID))>=181)],
                       polar.dat.1802[-which(as.numeric(as.character(dat.list$TAN1802$transectID))>=160)],
                       polar.dat.1901[-which(dat.list$TAN1901$transectID=="209")],
                       polar.dat.NBP1402,#[-2],
                       polar.dat.NBP1502,
                       polar.dat.AA2011,
                       polar.dat.JR262,
                       polar.dat.JR15005,
                       polar.dat.JR17001,
                       polar.dat.JR17003,
                       polar.dat.LMG1311)
names(polar.dat.list.clean) <- names(dat.list)
dat.list.clean <- list()
dat.list.clean$WAP <- dat.WAP[-which(dat.list$WAP$transectID%in%c(1103,1325)),]
dat.list.clean$WAP$transectID <- factor(dat.list.clean$WAP$transectID)
dat.list.clean$PS06 <- dat.PS06
dat.list.clean$PS14 <- dat.PS14
dat.list.clean$PS18 <- dat.PS18
dat.list.clean$PS61 <- dat.PS61
dat.list.clean$PS81 <- dat.PS81[-which(dat.list$PS81$transectID=="PS81_159"),]
dat.list.clean$PS81$transectID <- factor(dat.list.clean$PS81$transectID)
dat.list.clean$PS81_shallow <- dat.PS81_shallow
dat.list.clean$PS96 <- dat.PS96
dat.list.clean$PS118 <- dat.PS118
dat.list.clean$TAN0802 <- dat.0802[-which(as.numeric(as.character(dat.list$TAN0802$transectID))>=181),]
dat.list.clean$TAN0802$transectID <- factor(dat.list.clean$TAN0802$transectID)
dat.list.clean$TAN1802 <- dat.1802[-which(as.numeric(as.character(dat.list$TAN1802$transectID))>=160),]
dat.list.clean$TAN1802$transectID <- factor(dat.list.clean$TAN1802$transectID)
dat.list.clean$TAN1901 <- dat.1901[-which(dat.list$TAN1901$transectID=="209"),]
dat.list.clean$TAN1901$transectID <- factor(dat.list.clean$TAN1901$transectID)
dat.list.clean$NBP1402 <- dat.NBP1402#[-2,]
dat.list.clean$NBP1502 <- dat.NBP1502
dat.list.clean$AA2011 <- dat.AA2011
dat.list.clean$JR262 <- dat.JR262
dat.list.clean$JR15005 <- dat.JR15005
dat.list.clean$JR17001 <- dat.JR17001
dat.list.clean$JR17003 <- dat.JR17003
dat.list.clean$LMG1311 <- dat.LMG1311

total.t.length.list.clean <- total.t.length.list
total.t.length.list.clean$WAP <- total.t.length.list$WAP[-c(4,32)]
total.t.length.list.clean$PS81 <- total.t.length.list$PS81[-3]
total.t.length.list.clean$TAN0802 <- total.t.length.list$TAN0802[-c(18:53)]
total.t.length.list.clean$TAN1802 <- total.t.length.list$TAN1802[-c(16:31)]
total.t.length.list.clean$TAN1901 <- total.t.length.list$TAN1901[-34]

##### create selector lookup for each dataframe
col.select.list <- list(c("lon","lat","Filename.standardised","transectID","Filename","image.select","Date"),
                        c("lon","lat","Filename.standardised","transectID","Filename","image.select",'time_start'),
                        c("lon","lat","Filename.standardised","transectID","Filename","image.select",'time_start'),
                        c("lon","lat","Filename.standardised","transectID","Filename","image.select",'time_start'),
                        c("lon","lat","Filename.standardised","transectID","Filename","image.select",'time_start'),
                        c("Longitude","Latitude","Filename.standardised","transectID","Filename","image.select",'Date.Time'),
                        c("Longitude","Latitude","Filename.standardised","transectID","Filename","image.select",'Date.Time'),
                        c("Longitude","Latitude","Filename.standardised","transectID","Filename","image.select",'Date.Time'),
                        c("Longitude","Latitude","Filename.standardised","transectID","Filename","image.select",'Date.Time'),
                        c("GPS_lon","GPS_lat","Filename.standardised","transectID","FileName","image.select",'time'),
                        c("GPS_lon","GPS_lat","Filename.standardised","transectID","FileName","image.select",'time'),
                        c("GPS_lon","GPS_lat","Filename.standardised","transectID","FileName","image.select",'time'),
                        c("GPS_lon","GPS_lat","Filename.standardised","transectID","FileName","image.select",'time'),
                        c("GPS_lon","GPS_lat","Filename.standardised","transectID","FileName","image.select","Date"),
                        c("lon","lat","Filename.standardised","transectID","Filename","image.select","DateTime"),
                        c("lon","lat","Filename.standardised","transectID","filename","image.select",'time'),
                        c("lon","lat","Filename.standardised","transectID","filename","image.select",'time'),
                        c("lon","lat","Filename.standardised","transectID","filename","image.select",'time'),
                        c("lon","lat","Filename.standardised","transectID","filename","image.select",'time'),
                        c("lon","lat","Filename.standardised","transectID","filename","image.select","DateTime"))

# #### standardise date and time formats, add missing dates where necessary
# for(i in 1:20){
#   a <- select(dat.list.clean[[i]],col.select.list[[i]][7])[1,]
#   print(a)
# }
# dat.list.clean[[16]]$time <- 2011 ## JR262
# dat.list.clean[[19]]$time <- 2018 ## JR17003



#####                        
col.select2 <- c("lon","lat","Filename.standardised","transectID","Filename","image.select","DateTime")
polar.dat.subset.list <- list()
dat.subset.list <- list()
for(i in 1:length(dat.list.clean)){
  col.select <- col.select.list[[i]]
  # message(names(dat.list.clean)[i])
  t.images <- ceiling(total.t.length.list.clean[[i]]/100)
  if(names(dat.list.clean)[i]=="PS06"){t.images <- c(5,4)}
  if(names(dat.list.clean)[i]=="PS14"){t.images[t.images==0] <- 5}
  if(names(dat.list.clean)[i]=="PS18"){t.images[t.images==0] <- 5}
  if(names(dat.list.clean)[i]=="NBP1402"){t.images[7] <- 7}
  if(names(dat.list.clean)[i]=="AA2011"){t.images <- rep(1,71)}
  if(names(dat.list.clean)[i]=="PS118"){t.images <- t.images.straight}
  if(names(dat.list.clean)[i]=="JR262"){t.images <- c(5,5,2,5)}
  if(names(dat.list.clean)[i]=="JR17003"){t.images[1] <- 2}
  if(names(dat.list.clean)[i]=="LMG1311"){t.images[1] <- 17}
  dat <- dat.list.clean[[i]]
  dat$transectID <- factor(dat$transectID)
  # print(levels(dat$transectID))
  pol.subset.dat <- polar.dat.AA2011[-c(1:435)]
  subset.dat <- dat.AA2011[-c(1:435),col.select2]
  t.levels <- levels(dat$transectID)
  if(names(dat.list.clean)[i]=="PS81"){
    dat$t.ID_temp <- factor(dat$t.ID_temp)
    t.levels <- levels(dat$t.ID_temp)}  
  for(k in 1:length(t.levels)){
    message(t.levels[k])
    if(names(dat.list.clean)[i]=="PS81"){
      subset.v <- which(dat$t.ID_temp==t.levels[k])
    }else subset.v <- which(dat$transectID==t.levels[k])
    ## select the correct number of images for each transect
    sel <- order(dat$image.select[subset.v])[1:t.images[k]]
    if(names(dat.list.clean)[i]=="WAP" & t.images[i]>length(dat$image.select[subset.v])){
      sel <- order(dat$image.select[subset.v])
    }
    ## create vector of selected images
    if(i==1){dat.sel <<-  subset.v[sel]
    }else dat.sel <-  c(dat.sel,subset.v[sel])
    # print(polar.dat.list.clean[[i]][subset.v[sel]])
    pol.subset.dat <- spRbind(pol.subset.dat,polar.dat.list.clean[[i]][subset.v[sel]])
    subset.dat <- rbind(subset.dat,dat.list.clean[[i]][subset.v[sel],col.select])
  }
  polar.dat.subset.list[[i]] <- pol.subset.dat
  dat.subset.list[[i]] <- subset.dat
  names(dat.subset.list[[i]]) <- col.select2
}
names(polar.dat.subset.list) <- names(dat.subset.list) <- names(dat.list)

```


```{r}
## if available, add image area to data
for(i in 1:length(dat.list.clean)){
  dat.list.clean[[i]]$area <- NA
  ## fill if there is a standard area for the survey
  if(names(dat.list.clean)[i]=="PS14"){
    dat.list.clean[[i]]$area <- 0.56
  }
  if(names(dat.list.clean)[i]=="PS18"){
    dat.list.clean[[i]]$area <- 0.9
  }
  if(names(dat.list.clean)[i]=="PS61"){
    dat.list.clean[[i]]$area <- 1
  }
  if(names(dat.list.clean)[i]=="WAP"){
    dat.list.clean[[i]]$area <- 3
  }
  if(names(dat.list.clean)[i]=="NBP1402"){
    dat.list.clean[[i]]$area <- 4.8
  }
  if(names(dat.list.clean)[i]%in%c("JR262","JR15005","JR17001","JR17003")){
    dat.list.clean[[i]]$area <- 0.51
  }
  ## first find for which images some area estimates exist (and then select only the first appearance of that image)
  area.sel.raw <- which(area.raw$`image filename`%in%dat.list.clean[[i]]$Filename.standardised)
  if(length(area.sel.raw>0)){
    area.sel <- area.sel.raw[seq(1,length(area.sel.raw),by=2)]
    area.raw$`image filename`[area.sel]
    ## then find  matches between data and area
    dat.sel <- match(area.raw$`image filename`[area.sel], dat.list.clean[[i]]$Filename.standardised)
    ## then fill if available
    dat.list.clean[[i]]$area[dat.sel] <- area.raw$`image area in mÂ˛`[area.sel]
  }
}

```


### save output for analysis

```{r, eval=FALSE, warning=FALSE, message=FALSE}
save(polar.dat.0802,
     polar.dat.1802,
     polar.dat.1901,
     polar.dat.WAP,
     polar.dat.NBP1402,
     polar.dat.NBP1502,
     polar.dat.PS06,
     polar.dat.PS14,
     polar.dat.PS18,
     polar.dat.PS61,
     polar.dat.PS81,
     polar.dat.PS81_shallow,
     polar.dat.PS96,
     polar.dat.PS118,
     polar.dat.AA2011,
     polar.dat.JR262,
     polar.dat.JR15005,
     polar.dat.JR17001,
     polar.dat.JR17003,
     polar.dat.LMG1311,
     polar.dat.list.clean,
     polar.dat.start.list,
     polar.dat.subset.list,
     dat.list,
     dat.list.clean,
     dat.subset.list,
     dat.start.list,
     total.t.length.list,
     total.t.length.list.clean,
     stereo,
     file=paste0(ARC_Data.dir,"prep_image/Circumpolar_DownwardImages_metadata.Rdata"))
```

### standardise image metadata for csv-files IS DONE IN annotation/1_Circumpolar_Annotation_Data_prep.R TO INCLUDE LASER POINTS
