---
title: "Read and clean up circumpolar image metadata"
author: "Jan Jansen"
date: "19/07/2021"
output: html_document
---

This script is to create the summary data:  
__"Circumpolar_DownwardImages_metadata.Rdata"__  
and standardised/consistent __metadata csv-files for each individual survey__  

TO DO FOR THE CSV-FILES:  
- transect-dates for CRS-data
- license-information for CRS, all TAN, all JR except 17003
- data-links for CRS, all TAN, all JR except 17003

This summary file contains the following objects:  

*  starting positions for each transect in each survey:
    + unprojected dataframes: __dat.start.list__
    + projected SpatialPoints: __polar.dat.start.list__
  
*  metadata:
    + original dataframes for each survey: __dat.list__
    + total transect lengths: __total.t.length.list__
    + clean data only including useful images: __dat.list.clean__
    + & clean projected SpatialPoints: __polar.dat.list.clean__
    + & clean total transect lengths: __total.t.length.list.clean__
  
*  images selected for annotation only:
    + unprojected dataframes: __dat.subset.list__
    + projected SpatialPoints: __polar.dat.subset.list__
  
*  individual surveys:
    + __polar.dat.XXX__  

*  polar stereographic projection used: __stereo__


```{r, warning=FALSE, message=FALSE}
library(raadtools)
library(readxl)
library(raster)
library(maptools)
library(lubridate)
library(plyr)
```

### Setting up the environment

```{r, warning=FALSE, message=FALSE}
## Jan's local machine:
sci.dir <-      "C:/Users/jjansen/Desktop/science/"
env.derived <-  paste0(sci.dir,"data_environmental/derived/")

## remote repository (DOESN'T WORK YET):
# env.dir <- "https://data.imas.utas.edu.au/data_transfer/admin/files/EnvironmentalData/"

## common paths (after "sci.dir")
tools.dir <-    paste0(sci.dir,"SouthernOceanBiodiversityMapping/Useful_Functions_Tools/")
ARC_Data.dir <- paste0(sci.dir,"SouthernOceanBiodiversityMapping/ARC_Data/")
image.dir <-      paste0(ARC_Data.dir,"prep_image/ReadIn_Circumpolar_DownwardImage_Data_And_Subset_Surveys/")

## polar stereographic projection:
stereo <- "+proj=stere +lat_0=-90 +lat_ts=-71 +lon_0=0 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs"
# stereo <- "+proj=stere +lat_0=-90 +lat_ts=-71 +lon_0=0 +k=1 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0"
```

```{r, warning=FALSE, message=FALSE}
#### load depth
r2 <- raster(paste0(env.derived,"Circumpolar_EnvData_500m_shelf_bathy_gebco_depth.grd"))

# #### load diatom data
# load(file=paste0(image.dir,"Circumpolar_Diatom_locations.Rdata"))

#### load coastline
load(paste0(env.derived,"Circumpolar_Coastline.Rdata"))
```

### load image area information
```{r}
R.dir <- "R:/IMAS/Antarctic_Seafloor/Clean_Data_For_Permanent_Storage/"
area.xlsx <- list.files(R.dir,pattern=".xlsx", full.names=TRUE)
area.raw <- ldply(area.xlsx, read_excel, skip=1)
```


### prepare data

```{r, warning=FALSE, message=FALSE}
#### Entire continental shelf
# par(mar=c(2,2,1,1))
# #png("diatom_overview.png",width=1250,height=1250)
# plot(r2, xlim=c(-3000000,3000000), ylim=c(-3000000,3000000),col=terrain.colors(99), main="Sample sites for sediment cores/grabs")
# plot(coast.proj, add=TRUE)
# scalebar(500000, type="bar", label=c("0","250","500"), below="km")
# points(polar.dat.diatom, pch=16, col="blue")
# legend("topright","Sample-sites",pch=16, col="blue")
#dev.off()

#### load image data
## Ross Sea
dat.0802 <- get(load(paste0(image.dir,"TAN0802_dat.Rdata")))
total.t.length.v.0802 <- total.t.length.v
total.t.length.v.0802[c(13,16,17,28,32:34,45,46)] <- t.l[c(13,16,17,28,32:34,45,46)] #used distance to start for these transects, not distance along transect
dat.1802 <- get(load(paste0(image.dir,"TAN1802_dat.Rdata")))
t.l.1802 <- t.l
total.t.length.v.1802 <- total.t.length.v
dat.1901 <- get(load(paste0(image.dir,"TAN1901_dat.Rdata")))
t.l.1901 <- t.l
total.t.length.v.1901 <- total.t.length.v
dat.NBP1502 <- get(load(file=paste0(image.dir,"NBP1502_dat.Rdata")))
total.t.length.v.NBP1502 <- total.t.length.v
t.l.NBP1502 <- t.l
## West Antarctic
dat.PS81 <- get(load(file=paste0(image.dir,"PS81_dat_FINAL.Rdata")))
total.t.length.v.PS81 <- total.t.length.v
dat.PS81_shallow <- get(load(file=paste0(image.dir,"PS81_dat_shallow.Rdata")))
total.t.length.v.PS81_shallow <- total.t.length.v
dat.PS96 <- get(load(file=paste0(image.dir,"PS96_dat_FINAL.Rdata")))
total.t.length.v.PS96 <- total.t.length.v
dat.PS118 <- get(load(file=paste0(image.dir,"PS118_dat.Rdata")))
total.t.length.v.PS118 <- total.t.length.v
## CRS
dat.WAP <- get(load(file=paste0(image.dir,"CRS_dat.Rdata")))
total.t.length.v.WAP <- total.t.length.v
## East Antarctic
dat.NBP1402 <- get(load(file=paste0(image.dir,"NBP1402_dat.Rdata")))
total.t.length.v.NBP1402 <- total.t.length.v
rm(dat,t.l,total.t.length.v)
## old Weddell Sea surveys
dat.PS06 <- get(load(file=paste0(image.dir,"PS06_dat.Rdata")))
dat.PS14 <- get(load(file=paste0(image.dir,"PS14_dat.Rdata")))
total.t.length.v.PS14 <- total.t.length.v
dat.PS18 <- get(load(file=paste0(image.dir,"PS18_dat.Rdata")))
total.t.length.v.PS18 <- total.t.length.v
dat.PS61 <- get(load(file=paste0(image.dir,"PS61_dat.Rdata")))
total.t.length.v.PS61 <- total.t.length.v
## Mertz
dat.AA2011 <- get(load(file=paste0(image.dir,"AA2011_dat.Rdata")))
## SUCS
dat.JR262 <- get(load(file=paste0(image.dir,"JR262_dat.Rdata")))
dat.JR15005 <- get(load(file=paste0(image.dir,"JR15005_dat.Rdata")))
total.t.length.v.JR15005 <- total.t.length.v
dat.JR17001 <- get(load(file=paste0(image.dir,"JR17001_dat.Rdata")))
total.t.length.v.JR17001 <- total.t.length.v
dat.JR17003 <- get(load(file=paste0(image.dir,"JR17003_dat.Rdata")))
total.t.length.v.JR17003 <- total.t.length.v
## LMG1311
dat.LMG1311 <- get(load(file=paste0(image.dir,"LMG1311_dat.Rdata")))
total.t.length.v.LMG1311 <- total.t.length.v

##### PS81: something went wrong with the "image_select"-values when redoing the subsetting and replacing shitty images. Only happened to images from image.select>=50
dat.PS81$image.select[dat.PS81$Filename=="PS81_116-3_2013-01-26T12_51_09_5010.jpg"] <- 50
dat.PS81$image.select[dat.PS81$Filename=="PS81_116-3_2013-01-26T14_21_26_5190.jpg"] <- 53
dat.PS81$image.select[dat.PS81$Filename=="PS81_116-3_2013-01-26T13_58_10_5143.jpg"] <- 56
dat.PS81$image.select[dat.PS81$Filename=="PS81_116-3_2013-01-26T12_34_31_4977.jpg"] <- 62
dat.PS81$image.select[dat.PS81$Filename=="PS81_116-3_2013-01-26T13_09_49_5047.jpg"] <- 64
dat.PS81$image.select[dat.PS81$Filename=="PS81_116-3_2013-01-26T13_18_23_5064.jpg"] <- 65
dat.PS81$image.select[dat.PS81$Filename=="PS81_116-3_2013-01-26T12_28_30_4964.jpg"] <- 66

##### TAN1901: something went really wrong with saving the image subsetting.
#####   - Couldn't figure out why the difference between list of annotated images and stored TAN1901_dat.Rdata
#####   - Only one version of TAN1901_dat.Rdata exists and the reproducing the annotated subset using the R-code is not possible
#####   - Seems to be often off by 1 or 2 images, bu most transect are not affected

##
dat.1901$image.select[dat.1901$FileName=="tan1901_095_039.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_095_227.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_095_036.jpg"] <- 2
dat.1901$image.select[dat.1901$FileName=="tan1901_095_229.jpg"] <- 5
##
dat.1901$image.select[dat.1901$FileName=="tan1901_096_054.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_096_196.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_096_140.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_096_245.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_096_264.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_096_084.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_096_206.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_096_159.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_096_066.jpg"] <- 2
dat.1901$image.select[dat.1901$FileName=="tan1901_096_197.jpg"] <- 3
dat.1901$image.select[dat.1901$FileName=="tan1901_096_141.jpg"] <- 4
dat.1901$image.select[dat.1901$FileName=="tan1901_096_242.jpg"] <- 5
dat.1901$image.select[dat.1901$FileName=="tan1901_096_265.jpg"] <- 9
dat.1901$image.select[dat.1901$FileName=="tan1901_096_086.jpg"] <- 10
dat.1901$image.select[dat.1901$FileName=="tan1901_096_208.jpg"] <- 11
##
dat.1901$image.select[dat.1901$FileName=="tan1901_098_228.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_098_226.jpg"] <- 5
##
dat.1901$image.select[dat.1901$FileName=="tan1901_102_090.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_102_056.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_102_116.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_102_221.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_102_088.jpg"] <- 4
dat.1901$image.select[dat.1901$FileName=="tan1901_102_057.jpg"] <- 6
dat.1901$image.select[dat.1901$FileName=="tan1901_102_115.jpg"] <- 8
dat.1901$image.select[dat.1901$FileName=="tan1901_102_222.jpg"] <- 9
##
dat.1901$image.select[dat.1901$FileName=="tan1901_105_093.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_105_122.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_105_092.jpg"] <- 3
dat.1901$image.select[dat.1901$FileName=="tan1901_105_121.jpg"] <- 7
##
dat.1901$image.select[dat.1901$FileName=="tan1901_113_068.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_113_182.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_113_139.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_113_212.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_113_067.jpg"] <- 2
dat.1901$image.select[dat.1901$FileName=="tan1901_113_185.jpg"] <- 3
dat.1901$image.select[dat.1901$FileName=="tan1901_113_140.jpg"] <- 4
dat.1901$image.select[dat.1901$FileName=="tan1901_113_213.jpg"] <- 5
##
dat.1901$image.select[dat.1901$FileName=="tan1901_149_051.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_149_189.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_149_134.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_149_224.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_149_104.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_149_162.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_149_241.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_149_053.jpg"] <- 2
dat.1901$image.select[dat.1901$FileName=="tan1901_149_185.jpg"] <- 3
dat.1901$image.select[dat.1901$FileName=="tan1901_149_133.jpg"] <- 4
dat.1901$image.select[dat.1901$FileName=="tan1901_149_223.jpg"] <- 5
dat.1901$image.select[dat.1901$FileName=="tan1901_149_103.jpg"] <- 6
dat.1901$image.select[dat.1901$FileName=="tan1901_149_160.jpg"] <- 8
dat.1901$image.select[dat.1901$FileName=="tan1901_149_243.jpg"] <- 9
##
dat.1901$image.select[dat.1901$FileName=="tan1901_157_257.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_157_020.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_157_157.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_157_100.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_157_129.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_157_237.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_157_174.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_157_114.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_157_082.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_157_254.jpg"] <- 1
dat.1901$image.select[dat.1901$FileName=="tan1901_157_017.jpg"] <- 2
dat.1901$image.select[dat.1901$FileName=="tan1901_157_155.jpg"] <- 3
dat.1901$image.select[dat.1901$FileName=="tan1901_157_096.jpg"] <- 4
dat.1901$image.select[dat.1901$FileName=="tan1901_157_128.jpg"] <- 8
dat.1901$image.select[dat.1901$FileName=="tan1901_157_238.jpg"] <- 9
dat.1901$image.select[dat.1901$FileName=="tan1901_157_171.jpg"] <- 11
dat.1901$image.select[dat.1901$FileName=="tan1901_157_113.jpg"] <- 12
dat.1901$image.select[dat.1901$FileName=="tan1901_157_083.jpg"] <- 14
##
dat.1901$image.select[dat.1901$FileName=="tan1901_165_107.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_165_222.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_165_033.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_165_151.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_165_108.jpg"] <- 8
dat.1901$image.select[dat.1901$FileName=="tan1901_165_223.jpg"] <- 9
dat.1901$image.select[dat.1901$FileName=="tan1901_165_031.jpg"] <- 10
dat.1901$image.select[dat.1901$FileName=="tan1901_165_150.jpg"] <- 11
##
dat.1901$image.select[dat.1901$FileName=="tan1901_166_245.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_166_018.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_166_148.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_166_063.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_166_244.jpg"] <- 1
dat.1901$image.select[dat.1901$FileName=="tan1901_166_025.jpg"] <- 2
dat.1901$image.select[dat.1901$FileName=="tan1901_166_151.jpg"] <- 3
dat.1901$image.select[dat.1901$FileName=="tan1901_166_065.jpg"] <- 6
##
dat.1901$image.select[dat.1901$FileName=="tan1901_168_207.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_168_127.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_168_170.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_168_053.jpg"] <- 999
dat.1901$image.select[dat.1901$FileName=="tan1901_168_206.jpg"] <- 1
dat.1901$image.select[dat.1901$FileName=="tan1901_168_128.jpg"] <- 3
dat.1901$image.select[dat.1901$FileName=="tan1901_168_168.jpg"] <- 5
dat.1901$image.select[dat.1901$FileName=="tan1901_168_054.jpg"] <- 6
```

### add standardised filenames to data

```{r, warning=FALSE, message=FALSE}
## PS06,14,18,61,81,96:
dat <- dat.PS06
dat.PS06$Filename.standardised <- paste0(substr(dat$Filename,1,8),"_",sprintf("%04d",dat$image.select),"__",dat$Filename)
dat <- dat.PS14
dat.PS14$Filename.standardised <- paste0(substr(dat$Filename,1,8),"_",sprintf("%04d",dat$image.select),"__",dat$Filename)
dat <- dat.PS18
dat.PS18$Filename.standardised <- paste0(substr(dat$Filename,1,8),"_",sprintf("%04d",dat$image.select),"__",dat$Filename)
dat <- dat.PS61
dat.PS61$Filename.standardised <- paste0(substr(dat$Filename,1,8),"_",sprintf("%04d",dat$image.select),"__",dat$Filename)
dat <- dat.PS81
dat.PS81$Filename.standardised <- paste0(substr(dat$Filename,1,8),"_",sprintf("%04d",dat$image.select),"__",dat$Filename)
dat <- dat.PS81_shallow
dat.PS81_shallow$Filename.standardised <- paste0(substr(dat$Filename,1,8),"_",sprintf("%04d",dat$image.select),"__",dat$Filename)
dat <- dat.PS96
dat.PS96$Filename.standardised <- paste0(substr(dat$Filename,1,8),"_",sprintf("%04d",dat$image.select),"__",dat$Filename,".jpg")
## PS118:
dat <- dat.PS118
dat.PS118$Filename.standardised <- paste0("PS118_",dat$transectID.forfile,"_",sprintf("%04d",dat$image.select),"__",dat$Filename,".jpg")
## TAN0802 & 1901:
dat <- dat.0802
dat.0802$Filename.standardised <- paste0(substr(dat$FileName,1,12),sprintf("%04d",dat$image.select),"__",dat$FileName)
dat <- dat.1901
dat.1901$Filename.standardised <- paste0(substr(dat$FileName,1,12),sprintf("%04d",dat$image.select),"__",dat$FileName)
## TAN1802:
dat <- dat.1802
dat.1802$Filename.standardised <- paste0("TAN1802_",dat$transectID,"_",sprintf("%04d",dat$image.select),"__",dat$FileName)
## NBP1402
dat <- dat.NBP1402
dat.NBP1402$Filename.standardised <- paste0("NBP1402_",dat$transectID,"_",sprintf("%04d",dat$image.select),"__",dat$FileName)
## NBP1402
dat <- dat.NBP1502
dat.NBP1502$Filename.standardised <- paste0("NBP1502_",dat$transectID,"_",sprintf("%04d",dat$image.select),"__",dat$FileName)
## CRS:
dat <- dat.WAP
dat.WAP$Filename.standardised <- paste0("CRS_",dat$transectID,"_",sprintf("%04d",dat$image.select),"__",dat$Filename)
## AA2011
dat <- dat.AA2011
dat.AA2011$Filename.standardised <- paste0("AA2011_",dat$transectID_folder,"_",sprintf("%04d",dat$image.select),"__",dat$Filename)
## SUCS
dat <- dat.JR262
dat.JR262$Filename.standardised <- paste0("JR262_",dat$transectID,"_",sprintf("%04d",dat$image.select),"__",dat$filename)
dat <- dat.JR15005
dat.JR15005$Filename.standardised <- paste0("JR15005_",dat$transectID,"_",sprintf("%04d",dat$image.select),"__",dat$filename)
dat <- dat.JR17001
dat.JR17001$Filename.standardised <- paste0("JR17001_",dat$transectID,"_",sprintf("%04d",dat$image.select),"__",dat$filename)
dat <- dat.JR17003
dat.JR17003$Filename.standardised <- paste0("JR17003_",dat$transectID,"_",sprintf("%04d",dat$image.select),"__",dat$filename)
## LMG1311
dat <- dat.LMG1311
dat.LMG1311$Filename.standardised <- paste0("LMG1311_",dat$transectID,"_",sprintf("%04d",dat$image.select),"__",dat$filename)

rm(dat)
```

### reproject biological data

```{r, warning=FALSE, message=FALSE}
##
spatial.dat <- data.frame(cbind(dat.0802$GPS_lon, dat.0802$GPS_lat))
names(spatial.dat) <- c("Longitude","Latitude")
coordinates(spatial.dat) <- c("Longitude","Latitude")
proj4string(spatial.dat) <- CRS("+proj=longlat +datum=WGS84")
polar.dat.0802 <- spTransform(spatial.dat, CRS(stereo))
##
spatial.dat <- data.frame(cbind(dat.1802$GPS_lon, dat.1802$GPS_lat))
names(spatial.dat) <- c("Longitude","Latitude")
coordinates(spatial.dat) <- c("Longitude","Latitude")
proj4string(spatial.dat) <- CRS("+proj=longlat +datum=WGS84")
polar.dat.1802 <- spTransform(spatial.dat, CRS(stereo))
##
spatial.dat <- data.frame(cbind(dat.1901$GPS_lon, dat.1901$GPS_lat))
names(spatial.dat) <- c("Longitude","Latitude")
coordinates(spatial.dat) <- c("Longitude","Latitude")
proj4string(spatial.dat) <- CRS("+proj=longlat +datum=WGS84")
polar.dat.1901 <- spTransform(spatial.dat, CRS(stereo))
##
spatial.dat <- data.frame(cbind(dat.PS81$Longitude, dat.PS81$Latitude))
names(spatial.dat) <- c("Longitude","Latitude")
coordinates(spatial.dat) <- c("Longitude","Latitude")
proj4string(spatial.dat) <- CRS("+proj=longlat +datum=WGS84")
polar.dat.PS81 <- spTransform(spatial.dat, CRS(stereo))
##
spatial.dat <- data.frame(cbind(dat.PS81_shallow$Longitude, dat.PS81_shallow$Latitude))
names(spatial.dat) <- c("Longitude","Latitude")
coordinates(spatial.dat) <- c("Longitude","Latitude")
proj4string(spatial.dat) <- CRS("+proj=longlat +datum=WGS84")
polar.dat.PS81_shallow <- spTransform(spatial.dat, CRS(stereo))
##
spatial.dat <- data.frame(cbind(dat.PS96$Longitude, dat.PS96$Latitude))
names(spatial.dat) <- c("Longitude","Latitude")
coordinates(spatial.dat) <- c("Longitude","Latitude")
proj4string(spatial.dat) <- CRS("+proj=longlat +datum=WGS84")
polar.dat.PS96 <- spTransform(spatial.dat, CRS(stereo))
##
spatial.dat <- data.frame(cbind(dat.PS118$Longitude, dat.PS118$Latitude))
names(spatial.dat) <- c("Longitude","Latitude")
coordinates(spatial.dat) <- c("Longitude","Latitude")
proj4string(spatial.dat) <- CRS("+proj=longlat +datum=WGS84")
polar.dat.PS118 <- spTransform(spatial.dat, CRS(stereo))
##
spatial.dat <- data.frame(cbind(dat.WAP$lon, dat.WAP$lat))
names(spatial.dat) <- c("Longitude","Latitude")
coordinates(spatial.dat) <- c("Longitude","Latitude")
proj4string(spatial.dat) <- CRS("+proj=longlat +datum=WGS84")
polar.dat.WAP <- spTransform(spatial.dat, CRS(stereo))
##
spatial.dat <- data.frame(cbind(dat.NBP1402$GPS_lon, dat.NBP1402$GPS_lat))
names(spatial.dat) <- c("Longitude","Latitude")
coordinates(spatial.dat) <- c("Longitude","Latitude")
proj4string(spatial.dat) <- CRS("+proj=longlat +datum=WGS84")
polar.dat.NBP1402 <- spTransform(spatial.dat, CRS(stereo))
##
spatial.dat <- data.frame(cbind(dat.NBP1502$GPS_lon, dat.NBP1502$GPS_lat))
names(spatial.dat) <- c("Longitude","Latitude")
coordinates(spatial.dat) <- c("Longitude","Latitude")
proj4string(spatial.dat) <- CRS("+proj=longlat +datum=WGS84")
polar.dat.NBP1502 <- spTransform(spatial.dat, CRS(stereo))
##
spatial.dat <- data.frame(cbind(dat.PS06$lon, dat.PS06$lat))
names(spatial.dat) <- c("Longitude","Latitude")
coordinates(spatial.dat) <- c("Longitude","Latitude")
proj4string(spatial.dat) <- CRS("+proj=longlat +datum=WGS84")
polar.dat.PS06 <- spTransform(spatial.dat, CRS(stereo))
##
spatial.dat <- data.frame(cbind(dat.PS14$lon, dat.PS14$lat))
names(spatial.dat) <- c("Longitude","Latitude")
coordinates(spatial.dat) <- c("Longitude","Latitude")
proj4string(spatial.dat) <- CRS("+proj=longlat +datum=WGS84")
polar.dat.PS14 <- spTransform(spatial.dat, CRS(stereo))
##
spatial.dat <- data.frame(cbind(dat.PS18$lon, dat.PS18$lat))
names(spatial.dat) <- c("Longitude","Latitude")
coordinates(spatial.dat) <- c("Longitude","Latitude")
proj4string(spatial.dat) <- CRS("+proj=longlat +datum=WGS84")
polar.dat.PS18 <- spTransform(spatial.dat, CRS(stereo))
##
spatial.dat <- data.frame(cbind(dat.PS61$lon, dat.PS61$lat))
names(spatial.dat) <- c("Longitude","Latitude")
coordinates(spatial.dat) <- c("Longitude","Latitude")
proj4string(spatial.dat) <- CRS("+proj=longlat +datum=WGS84")
polar.dat.PS61 <- spTransform(spatial.dat, CRS(stereo))
##
spatial.dat <- data.frame(cbind(dat.AA2011$lon, dat.AA2011$lat))
names(spatial.dat) <- c("Longitude","Latitude")
coordinates(spatial.dat) <- c("Longitude","Latitude")
proj4string(spatial.dat) <- CRS("+proj=longlat +datum=WGS84")
polar.dat.AA2011 <- spTransform(spatial.dat, CRS(stereo))
##
spatial.dat <- data.frame(cbind(dat.JR262$lon, dat.JR262$lat))
names(spatial.dat) <- c("Longitude","Latitude")
coordinates(spatial.dat) <- c("Longitude","Latitude")
proj4string(spatial.dat) <- CRS("+proj=longlat +datum=WGS84")
polar.dat.JR262 <- spTransform(spatial.dat, CRS(stereo))
##
spatial.dat <- data.frame(cbind(dat.JR15005$lon, dat.JR15005$lat))
names(spatial.dat) <- c("Longitude","Latitude")
coordinates(spatial.dat) <- c("Longitude","Latitude")
proj4string(spatial.dat) <- CRS("+proj=longlat +datum=WGS84")
polar.dat.JR15005 <- spTransform(spatial.dat, CRS(stereo))
##
spatial.dat <- data.frame(cbind(dat.JR17001$lon, dat.JR17001$lat))
names(spatial.dat) <- c("Longitude","Latitude")
coordinates(spatial.dat) <- c("Longitude","Latitude")
proj4string(spatial.dat) <- CRS("+proj=longlat +datum=WGS84")
polar.dat.JR17001 <- spTransform(spatial.dat, CRS(stereo))
##
spatial.dat <- data.frame(cbind(dat.JR17003$lon, dat.JR17003$lat))
names(spatial.dat) <- c("Longitude","Latitude")
coordinates(spatial.dat) <- c("Longitude","Latitude")
proj4string(spatial.dat) <- CRS("+proj=longlat +datum=WGS84")
polar.dat.JR17003 <- spTransform(spatial.dat, CRS(stereo))
##
spatial.dat <- data.frame(cbind(dat.LMG1311$lon, dat.LMG1311$lat))
names(spatial.dat) <- c("Longitude","Latitude")
coordinates(spatial.dat) <- c("Longitude","Latitude")
proj4string(spatial.dat) <- CRS("+proj=longlat +datum=WGS84")
polar.dat.LMG1311 <- spTransform(spatial.dat, CRS(stereo))

####
# ## need start and end positions for PS14 and PS18
# spatial.dat <- data.frame(cbind(dat.PS14$Longitude_Start, dat.PS14$Latitude_Start))
# names(spatial.dat) <- c("Longitude","Latitude")
# coordinates(spatial.dat) <- c("Longitude","Latitude")
# proj4string(spatial.dat) <- CRS("+proj=longlat +datum=WGS84")
# polar.dat.PS14.start <- spTransform(spatial.dat, CRS(stereo))
# ##
# spatial.dat <- data.frame(cbind(dat.PS14$Longitude_End, dat.PS14$Latitude_End))
# names(spatial.dat) <- c("Longitude","Latitude")
# coordinates(spatial.dat) <- c("Longitude","Latitude")
# proj4string(spatial.dat) <- CRS("+proj=longlat +datum=WGS84")
# polar.dat.PS14.end <- spTransform(spatial.dat, CRS(stereo))
# ##
# spatial.dat <- data.frame(cbind(dat.PS18$Longitude_Start, dat.PS18$Latitude_Start))
# names(spatial.dat) <- c("Longitude","Latitude")
# spatial.dat <- spatial.dat[-which(is.na(rowSums(spatial.dat))),]
# coordinates(spatial.dat) <- c("Longitude","Latitude")
# proj4string(spatial.dat) <- CRS("+proj=longlat +datum=WGS84")
# polar.dat.PS18.start <- spTransform(spatial.dat, CRS(stereo))
# ##
# spatial.dat <- data.frame(cbind(dat.PS18$Longitude_End, dat.PS18$Latitude_End))
# names(spatial.dat) <- c("Longitude","Latitude")
# spatial.dat <- spatial.dat[-which(is.na(rowSums(spatial.dat))),]
# coordinates(spatial.dat) <- c("Longitude","Latitude")
# proj4string(spatial.dat) <- CRS("+proj=longlat +datum=WGS84")
# polar.dat.PS18.end <- spTransform(spatial.dat, CRS(stereo))
```

### create single files/lists for all surveys together

```{r, warning=FALSE, message=FALSE}
## start-points and ID for all transects:
dat.start.list <- list()
dat.start.list$WAP <- dat.WAP[!duplicated(dat.WAP$transectID),]
dat.start.list$PS06 <- dat.PS06[!duplicated(dat.PS06$transectID),]
dat.start.list$PS14 <- dat.PS14[!duplicated(dat.PS14$transectID),]
dat.start.list$PS18 <- dat.PS18[!duplicated(dat.PS18$transectID),]
dat.start.list$PS61 <- dat.PS61[!duplicated(dat.PS61$transectID),]
dat.start.list$PS81 <- dat.PS81[!duplicated(dat.PS81$transectID),]
dat.start.list$PS81_shallow <- dat.PS81_shallow[!duplicated(dat.PS81_shallow$transectID),]
dat.start.list$PS96 <- dat.PS96[!duplicated(dat.PS96$transectID),]
dat.start.list$PS118 <- dat.PS118[!duplicated(dat.PS118$transectID),]
dat.start.list$TAN0802 <- dat.0802[!duplicated(dat.0802$transectID),]
dat.start.list$TAN1802 <- dat.1802[!duplicated(dat.1802$transectID),]
dat.start.list$TAN1901 <- dat.1901[!duplicated(dat.1901$transectID),]
dat.start.list$NBP1402 <- dat.NBP1402[!duplicated(dat.NBP1402$transectID),]
dat.start.list$NBP1502 <- dat.NBP1502[!duplicated(dat.NBP1502$transectID),]
dat.start.list$AA2011 <- dat.AA2011[!duplicated(dat.AA2011$transectID),]
dat.start.list$JR262 <- dat.JR262[!duplicated(dat.JR262$transectID),]
dat.start.list$JR15005 <- dat.JR15005[!duplicated(dat.JR15005$transectID),]
dat.start.list$JR17001 <- dat.JR17001[!duplicated(dat.JR17001$transectID),]
dat.start.list$JR17003 <- dat.JR17003[!duplicated(dat.JR17003$transectID),]
dat.start.list$LMG1311 <- dat.LMG1311[!duplicated(dat.LMG1311$transectID),]

polar.dat.start.list <- list()
polar.dat.start.list$WAP <- polar.dat.WAP[!duplicated(dat.WAP$transectID),]
polar.dat.start.list$PS06 <- polar.dat.PS06[!duplicated(dat.PS06$transectID),]
polar.dat.start.list$PS14 <- polar.dat.PS14[!duplicated(dat.PS14$transectID),]
polar.dat.start.list$PS18 <- polar.dat.PS18[!duplicated(dat.PS18$transectID),]
polar.dat.start.list$PS61 <- polar.dat.PS61[!duplicated(dat.PS61$transectID),]
polar.dat.start.list$PS81 <- polar.dat.PS81[!duplicated(dat.PS81$transectID),]
polar.dat.start.list$PS81_shallow <- polar.dat.PS81_shallow[!duplicated(dat.PS81_shallow$transectID),]
polar.dat.start.list$PS96 <- polar.dat.PS96[!duplicated(dat.PS96$transectID),]
polar.dat.start.list$PS118 <- polar.dat.PS118[!duplicated(dat.PS118$transectID),]
polar.dat.start.list$TAN0802 <- polar.dat.0802[!duplicated(dat.0802$transectID),]
polar.dat.start.list$TAN1802 <- polar.dat.1802[!duplicated(dat.1802$transectID),]
polar.dat.start.list$TAN1901 <- polar.dat.1901[!duplicated(dat.1901$transectID),]
polar.dat.start.list$NBP1402 <- polar.dat.NBP1402[!duplicated(dat.NBP1402$transectID),]
polar.dat.start.list$NBP1502 <- polar.dat.NBP1502[!duplicated(dat.NBP1502$transectID),]
polar.dat.start.list$AA2011 <- polar.dat.AA2011[!duplicated(dat.AA2011$transectID),]
polar.dat.start.list$JR262 <- polar.dat.JR262[!duplicated(dat.JR262$transectID),]
polar.dat.start.list$JR15005 <- polar.dat.JR15005[!duplicated(dat.JR15005$transectID),]
polar.dat.start.list$JR17001 <- polar.dat.JR17001[!duplicated(dat.JR17001$transectID),]
polar.dat.start.list$JR17003 <- polar.dat.JR17003[!duplicated(dat.JR17003$transectID),]
polar.dat.start.list$LMG1311 <- polar.dat.LMG1311[!duplicated(dat.LMG1311$transectID),]

dat.list <- list()
dat.list$WAP <- dat.WAP
dat.list$PS06 <- dat.PS06
dat.list$PS14 <- dat.PS14
dat.list$PS18 <- dat.PS18
dat.list$PS61 <- dat.PS61
dat.list$PS81 <- dat.PS81
dat.list$PS81_shallow <- dat.PS81_shallow
dat.list$PS96 <- dat.PS96
dat.list$PS118 <- dat.PS118
dat.list$TAN0802 <- dat.0802
dat.list$TAN1802 <- dat.1802
dat.list$TAN1901 <- dat.1901
dat.list$NBP1402 <- dat.NBP1402
dat.list$NBP1502 <- dat.NBP1502
dat.list$AA2011 <- dat.AA2011
dat.list$JR262 <- dat.JR262
dat.list$JR15005 <- dat.JR15005
dat.list$JR17001 <- dat.JR17001
dat.list$JR17003 <- dat.JR17003
dat.list$LMG1311 <- dat.LMG1311

total.t.length.list <- list()
total.t.length.list$WAP <- total.t.length.v.WAP
total.t.length.list$PS06 <- NA
total.t.length.list$PS14 <- total.t.length.v.PS14
total.t.length.list$PS18 <- total.t.length.v.PS18
total.t.length.list$PS61 <- total.t.length.v.PS61
total.t.length.list$PS81 <- total.t.length.v.PS81
total.t.length.list$PS81_shallow <- total.t.length.v.PS81_shallow
total.t.length.list$PS96 <- total.t.length.v.PS96
total.t.length.list$PS118 <- total.t.length.v.PS118
total.t.length.list$TAN0802 <- total.t.length.v.0802
total.t.length.list$TAN1802 <- total.t.length.v.1802
total.t.length.list$TAN1901 <- total.t.length.v.1901
total.t.length.list$NBP1402 <- total.t.length.v.NBP1402
total.t.length.list$NBP1502 <- total.t.length.v.NBP1502
total.t.length.list$AA2011 <- NA
total.t.length.list$JR262 <- NA
total.t.length.list$JR15005 <- total.t.length.v.JR15005
total.t.length.list$JR17001 <- total.t.length.v.JR17001
total.t.length.list$JR17003 <- total.t.length.v.JR17003
total.t.length.list$LMG1311 <- total.t.length.v.LMG1311

########## annotated subset only
## following transects were disregarded:
## CRS 1103 & CRS 1325              # not yet removed
## PS06 288, 301-303, 307,309-311   # already removed
## PS81 159                         # not yet removed
## PS118 38                         # already removed
## TAN0802 181-310                  # not yet removed
## TAN1802 160-213                  # not yet removed
## TAN1901 209                      # not yet removed
## NBP1402 just a single image that is misplaced # not yet removed
## AA2011 CTD53,CTD55,CTD57-58,CTD61-64,CTD85,TD90,CTD93,CTD98,CTD107,CTD115,CTD138 # already removed
polar.dat.list.clean <- list(polar.dat.WAP[-which(dat.list$WAP$transectID%in%c(1103,1325))],
                       polar.dat.PS06,
                       polar.dat.PS14,
                       polar.dat.PS18,
                       polar.dat.PS61,
                       polar.dat.PS81[-which(dat.list$PS81$transectID=="PS81_159")],
                       polar.dat.PS81_shallow,
                       polar.dat.PS96,
                       polar.dat.PS118,
                       polar.dat.0802[-which(as.numeric(as.character(dat.list$TAN0802$transectID))>=181)],
                       polar.dat.1802[-which(as.numeric(as.character(dat.list$TAN1802$transectID))>=160)],
                       polar.dat.1901[-which(dat.list$TAN1901$transectID=="209")],
                       polar.dat.NBP1402,#[-2],
                       polar.dat.NBP1502,
                       polar.dat.AA2011,
                       polar.dat.JR262,
                       polar.dat.JR15005,
                       polar.dat.JR17001,
                       polar.dat.JR17003,
                       polar.dat.LMG1311)
names(polar.dat.list.clean) <- names(dat.list)
dat.list.clean <- list()
dat.list.clean$WAP <- dat.WAP[-which(dat.list$WAP$transectID%in%c(1103,1325)),]
dat.list.clean$WAP$transectID <- factor(dat.list.clean$WAP$transectID)
dat.list.clean$PS06 <- dat.PS06
dat.list.clean$PS14 <- dat.PS14
dat.list.clean$PS18 <- dat.PS18
dat.list.clean$PS61 <- dat.PS61
dat.list.clean$PS81 <- dat.PS81[-which(dat.list$PS81$transectID=="PS81_159"),]
dat.list.clean$PS81$transectID <- factor(dat.list.clean$PS81$transectID)
dat.list.clean$PS81_shallow <- dat.PS81_shallow
dat.list.clean$PS96 <- dat.PS96
dat.list.clean$PS118 <- dat.PS118
dat.list.clean$TAN0802 <- dat.0802[-which(as.numeric(as.character(dat.list$TAN0802$transectID))>=181),]
dat.list.clean$TAN0802$transectID <- factor(dat.list.clean$TAN0802$transectID)
dat.list.clean$TAN1802 <- dat.1802[-which(as.numeric(as.character(dat.list$TAN1802$transectID))>=160),]
dat.list.clean$TAN1802$transectID <- factor(dat.list.clean$TAN1802$transectID)
dat.list.clean$TAN1901 <- dat.1901[-which(dat.list$TAN1901$transectID=="209"),]
dat.list.clean$TAN1901$transectID <- factor(dat.list.clean$TAN1901$transectID)
dat.list.clean$NBP1402 <- dat.NBP1402#[-2,]
dat.list.clean$NBP1502 <- dat.NBP1502
dat.list.clean$AA2011 <- dat.AA2011
dat.list.clean$JR262 <- dat.JR262
dat.list.clean$JR15005 <- dat.JR15005
dat.list.clean$JR17001 <- dat.JR17001
dat.list.clean$JR17003 <- dat.JR17003
dat.list.clean$LMG1311 <- dat.LMG1311

total.t.length.list.clean <- total.t.length.list
total.t.length.list.clean$WAP <- total.t.length.list$WAP[-c(4,32)]
total.t.length.list.clean$PS81 <- total.t.length.list$PS81[-3]
total.t.length.list.clean$TAN0802 <- total.t.length.list$TAN0802[-c(18:53)]
total.t.length.list.clean$TAN1802 <- total.t.length.list$TAN1802[-c(16:31)]
total.t.length.list.clean$TAN1901 <- total.t.length.list$TAN1901[-34]

# "CRS",'Date'
# "PS06",'time_start'
# "PS14",'time_start'
# "PS18",'time_start'
# "PS61",'time_start'
# "PS81",'Date.Time'
# "PS96",'Date.Time'
# "PS118",'Date.Time'
# "TAN0802",'time'
# "TAN1802",'time'
# "TAN1901",'time'
# "NBP1402",'time'
# "NBP1502",'Date'
# "AA2011",'DateTime'
# "JR262",'time'
# "JR15005",'time'
# "JR17001",'time'
# "JR17003",'time'
# "LMG1311",'DateTime'

##### create selector lookup for each dataframe
col.select.list <- list(c("lon","lat","Filename.standardised","transectID","Filename","image.select","Date"),
                        c("lon","lat","Filename.standardised","transectID","Filename","image.select",'time_start'),
                        c("lon","lat","Filename.standardised","transectID","Filename","image.select",'time_start'),
                        c("lon","lat","Filename.standardised","transectID","Filename","image.select",'time_start'),
                        c("lon","lat","Filename.standardised","transectID","Filename","image.select",'time_start'),
                        c("Longitude","Latitude","Filename.standardised","transectID","Filename","image.select",'Date.Time'),
                        c("Longitude","Latitude","Filename.standardised","transectID","Filename","image.select",'Date.Time'),
                        c("Longitude","Latitude","Filename.standardised","transectID","Filename","image.select",'Date.Time'),
                        c("Longitude","Latitude","Filename.standardised","transectID","Filename","image.select",'Date.Time'),
                        c("GPS_lon","GPS_lat","Filename.standardised","transectID","FileName","image.select",'time'),
                        c("GPS_lon","GPS_lat","Filename.standardised","transectID","FileName","image.select",'time'),
                        c("GPS_lon","GPS_lat","Filename.standardised","transectID","FileName","image.select",'time'),
                        c("GPS_lon","GPS_lat","Filename.standardised","transectID","FileName","image.select",'time'),
                        c("GPS_lon","GPS_lat","Filename.standardised","transectID","FileName","image.select","Date"),
                        c("lon","lat","Filename.standardised","transectID","Filename","image.select","DateTime"),
                        c("lon","lat","Filename.standardised","transectID","filename","image.select",'time'),
                        c("lon","lat","Filename.standardised","transectID","filename","image.select",'time'),
                        c("lon","lat","Filename.standardised","transectID","filename","image.select",'time'),
                        c("lon","lat","Filename.standardised","transectID","filename","image.select",'time'),
                        c("lon","lat","Filename.standardised","transectID","filename","image.select","DateTime"))
#####                        
col.select2 <- c("lon","lat","Filename.standardised","transectID","Filename","image.select","DateTime")
polar.dat.subset.list <- list()
dat.subset.list <- list()
for(i in 1:length(dat.list.clean)){
  col.select <- col.select.list[[i]]
  # message(names(dat.list.clean)[i])
  t.images <- ceiling(total.t.length.list.clean[[i]]/100)
  if(names(dat.list.clean)[i]=="PS06"){t.images <- c(5,4)}
  if(names(dat.list.clean)[i]=="PS14"){t.images[t.images==0] <- 5}
  if(names(dat.list.clean)[i]=="PS18"){t.images[t.images==0] <- 5}
  if(names(dat.list.clean)[i]=="NBP1402"){t.images[7] <- 7}
  if(names(dat.list.clean)[i]=="AA2011"){t.images <- rep(1,71)}
  if(names(dat.list.clean)[i]=="PS118"){t.images <- t.images.straight}
  if(names(dat.list.clean)[i]=="JR262"){t.images <- c(5,5,2,5)}
  if(names(dat.list.clean)[i]=="JR17003"){t.images[1] <- 2}
  if(names(dat.list.clean)[i]=="LMG1311"){t.images[1] <- 17}
  dat <- dat.list.clean[[i]]
  dat$transectID <- factor(dat$transectID)
  # print(levels(dat$transectID))
  pol.subset.dat <- polar.dat.AA2011[-c(1:435)]
  subset.dat <- dat.AA2011[-c(1:435),col.select2]
  t.levels <- levels(dat$transectID)
  if(names(dat.list.clean)[i]=="PS81"){
    dat$t.ID_temp <- factor(dat$t.ID_temp)
    t.levels <- levels(dat$t.ID_temp)}  
  for(k in 1:length(t.levels)){
    message(t.levels[k])
    if(names(dat.list.clean)[i]=="PS81"){
      subset.v <- which(dat$t.ID_temp==t.levels[k])
    }else subset.v <- which(dat$transectID==t.levels[k])
    ## select the correct number of images for each transect
    sel <- order(dat$image.select[subset.v])[1:t.images[k]]
    if(names(dat.list.clean)[i]=="WAP" & t.images[i]>length(dat$image.select[subset.v])){
      sel <- order(dat$image.select[subset.v])
    }
    ## create vector of selected images
    if(i==1){dat.sel <<-  subset.v[sel]
    }else dat.sel <-  c(dat.sel,subset.v[sel])
    # print(polar.dat.list.clean[[i]][subset.v[sel]])
    pol.subset.dat <- spRbind(pol.subset.dat,polar.dat.list.clean[[i]][subset.v[sel]])
    subset.dat <- rbind(subset.dat,dat.list.clean[[i]][subset.v[sel],col.select])
  }
  polar.dat.subset.list[[i]] <- pol.subset.dat
  dat.subset.list[[i]] <- subset.dat
  names(dat.subset.list[[i]]) <- col.select2
}
names(polar.dat.subset.list) <- names(dat.subset.list) <- names(dat.list)

## if available, add image area to data
for(i in 1:length(dat.subset.list)){
  dat.subset.list[[i]]$area <- NA
  ## fill if there is a standard area for the survey
  if(names(dat.subset.list)[i]=="PS14"){
    dat.subset.list[[i]]$area <- 0.56
  }
  if(names(dat.subset.list)[i]=="PS18"){
    dat.subset.list[[i]]$area <- 0.9
  }
  if(names(dat.subset.list)[i]=="PS61"){
    dat.subset.list[[i]]$area <- 1
  }
  if(names(dat.subset.list)[i]=="WAP"){
    dat.subset.list[[i]]$area <- 3
  }
  if(names(dat.subset.list)[i]=="NBP1402"){
    dat.subset.list[[i]]$area <- 4.8
  }
  if(names(dat.subset.list)[i]%in%c("JR262","JR15005","JR17001","JR17003")){
    dat.subset.list[[i]]$area <- 0.51
  }
  ## first find for which images some area estimates exist (and then select only the first appearance of that image)
  area.sel.raw <- which(area.raw$`image filename`%in%dat.subset.list[[i]]$Filename.standardised)
  if(length(area.sel.raw>0)){
    area.sel <- area.sel.raw[seq(1,length(area.sel.raw),by=2)]
    area.raw$`image filename`[area.sel]
    ## then find  matches between data and area
    dat.sel <- match(area.raw$`image filename`[area.sel], dat.subset.list[[i]]$Filename.standardised)
    ## then fill if available
    dat.subset.list[[i]]$area[dat.sel] <- area.raw$`image area in m²`[area.sel]
  }
}
```

```{r, warning=FALSE, message=FALSE}
for(i in 1:length(dat.subset.list)){
  # Convert area values from string to float
  dat.subset.list[[i]]$area <- as.numeric(dat.subset.list[[i]]$area)
  print(names(dat.subset.list)[[i]])
  # Check if missing area values
  if (sum(is.na(dat.subset.list[[i]]$area))) {
    # No data for PS06
    if (names(dat.subset.list)[[i]] == "PS06") {
      print("No area available for PS06... filling with NA...")
      dat.subset.list[[i]]$estimated_area <- NA
    }
    else {
      # For some surveys, we fill missing values with the average value across the survey
      if (names(dat.subset.list)[[i]] %in% c("PS96", "AA2011", "NBP1502", "TAN1802")) {
        # Get indexes
        idx_nonNA_values <- which(!is.na(dat.subset.list[[i]]$area))
        filling_value <- mean(dat.subset.list[[i]]$area[idx_nonNA_values])
      }
      else {
        filling_value <- NA
      }
      # Iterate through all transects
      transect_ids <- as.vector(unique(dat.subset.list[[i]]$transectID))
      for (transect_id in transect_ids) {
        # Get transect data
        idxes_transet <- which(dat.subset.list[[i]]$transectID == transect_id)
        dat_transect <- dat.subset.list[[i]][idxes_transet, ]
        if (sum(is.na(dat_transect$area)) == nrow(dat_transect)) {
          # If no area value available at all, estimated_area filled with NA
          dat.subset.list[[i]][idxes_transet, ]$estimated_area <- filling_value
          print(paste("Missing data for survey: ", names(dat.subset.list)[[i]] , " / Transect: ", as.vector(unique(dat_transect$transectID)), " ... filling with ", filling_value))
        }
        else if (sum(is.na(dat_transect$area)) == 0) {
          dat.subset.list[[i]][idxes_transet, ]$estimated_area <- dat.subset.list[[i]][idxes_transet, ]$area
        }
        else {
          # Get indexes
          idx_NA_values <- which(is.na(dat_transect$area))
          idx_nonNA_values <- which(!is.na(dat_transect$area))
     
          # Init column
          dat_transect$estimated_area <- NA

          # Compute average value of non NA images
          avg_value <- mean(dat_transect$area[idx_nonNA_values])

          # Assign this value to images with NA area
          dat_transect$estimated_area[idx_NA_values] <- avg_value

          # Assign true area for images with non NA area
          dat_transect$estimated_area[idx_nonNA_values] <- dat.subset.list[[i]][idxes_transet, ]$area[idx_nonNA_values]
          
          # Restore values
          dat.subset.list[[i]][idxes_transet, ] <- dat_transect
        }
      }
    }
  }
  else {
    # If no missing values, the estimated area = true area
    dat.subset.list[[i]]$estimated_area <- dat.subset.list[[i]]$area
  }
}
```

### save output for analysis

```{r, eval=FALSE, warning=FALSE, message=FALSE}
save(polar.dat.0802,
     polar.dat.1802,
     polar.dat.1901,
     polar.dat.WAP,
     polar.dat.NBP1402,
     polar.dat.NBP1502,
     polar.dat.PS06,
     polar.dat.PS14,
     polar.dat.PS18,
     polar.dat.PS61,
     polar.dat.PS81,
     polar.dat.PS81_shallow,
     polar.dat.PS96,
     polar.dat.PS118,
     polar.dat.AA2011,
     polar.dat.JR262,
     polar.dat.JR15005,
     polar.dat.JR17001,
     polar.dat.JR17003,
     polar.dat.LMG1311,
     polar.dat.list.clean,
     polar.dat.start.list,
     polar.dat.subset.list,
     dat.list,
     dat.list.clean,
     dat.subset.list,
     dat.start.list,
     total.t.length.list,
     total.t.length.list.clean,
     stereo,
     file=paste0(ARC_Data.dir,"Circumpolar_DownwardImages_metadata.Rdata"))
```

### standardise image metadata for csv-files

We need:
* SurveyID  
* TransectID  
* Filename  
* Filename-standardised  
* Longitude  
* Latitude  
* Date taken 
* Measured depth
* Link to original source  
* License  

```{r, eval=FALSE, warning=FALSE}
##### first, read in all metadata and store as a list
csv.list <- list()
csv.names <- c("Filename","Filename.standardised","Survey.ID","Transect.ID","Longitude","Latitude",
               "Date","Date_start","Date_end","Source.link","License","Depth","Depth_start","Depth_end","Area","Area_from_annotations")

csv.list$CRS <- cbind(dat.list$WAP[,c('Filename','Filename.standardised')],
                      "CRS",dat.list$WAP[,c('transectID','lon','lat')],
                       NA,NA,NA,
                       NA,NA,
                       dat.list$WAP[,c('depth.mean')],NA,NA,3,NA)

csv.list$PS06 <- cbind(dat.list$PS06[,c('Filename','Filename.standardised')],
                       "PS06",dat.list$PS06[,c('transectID','lon','lat')],
                       NA,dat.list$PS06[,c('time_start','time_end')],
                       NA,"CC-BY-3.0",
                       dat.list$PS06[,c('depth')],NA,NA,NA,NA)

csv.list$PS14 <- cbind(dat.list$PS14[,c('Filename','Filename.standardised')],
                       "PS14",dat.list$PS14[,c('transectID','lon','lat')],
                       NA,dat.list$PS14[,c('time_start','time_end')],
                       NA,"CC-BY-3.0",
                       NA,dat.list$PS14[,c('depth_start','depth_end')],0.56,NA)

csv.list$PS18 <- cbind(dat.list$PS18[,c('Filename','Filename.standardised')],
                       "PS18",dat.list$PS18[,c('transectID','lon','lat')],
                       NA,dat.list$PS18[,c('time_start','time_end')],
                       NA,"CC-BY-3.0",
                       NA,dat.list$PS18[,c('depth_start','depth_end')],0.9,NA)

csv.list$PS61 <- cbind(dat.list$PS61[,c('Filename','Filename.standardised')],
                       "PS61",dat.list$PS61[,c('transectID','lon','lat')],
                       NA,dat.list$PS61[,c('time_start','time_end')],
                       NA,"CC-BY-3.0",
                       NA,dat.list$PS61[,c('depth_start','depth_end')],1,NA)


temp.PS81 <- cbind(dat.list$PS81[,c('Filename','Filename.standardised')],
                             "PS81",dat.list$PS81[,c('transectID','Longitude','Latitude','Date.Time')],
                             NA,NA,
                             "doi.pangaea.de/10.1594/PANGAEA.872719","CC-BY-3.0",
                             dat.list$PS81[,c('Bathy.depth..m.')],NA,NA,dat.list$PS81[,c('Area_from_metadata','Area')])
temp.PS81_s <- cbind(dat.list$PS81_shallow[,c('Filename','Filename.standardised')],
                             "PS81",dat.list$PS81_shallow[,c('transectID','Longitude','Latitude','Date.Time')],
                             NA,NA,
                             "doi.pangaea.de/10.1594/PANGAEA.872719","CC-BY-3.0",
                             dat.list$PS81_shallow[,c('Bathy.depth..m.')],NA,NA,dat.list$PS81_shallow[,c('Area_from_metadata','Area')])
names(temp.PS81) <- names(temp.PS81_s) <- csv.names
csv.list$PS81 <- rbind(temp.PS81,temp.PS81_s)

csv.list$PS96 <- cbind(dat.list$PS96[,c('Filename','Filename.standardised')],
                             "PS96",dat.list$PS96[,c('transectID','Longitude','Latitude','Date.Time')],
                             NA,NA,
                             "doi.pangaea.de/10.1594/PANGAEA.862097","CC-BY-3.0",
                             dat.list$PS96[,c('Depth.water..m.')],NA,NA,NA,dat.list$PS96[,c('Area')])

csv.list$PS118 <- cbind(dat.list$PS118[,c('Filename','Filename.standardised')],
                             "PS118",dat.list$PS118[,c('transectID','Longitude','Latitude','Date.Time')],
                             NA,NA,
                             "doi.pangaea.de/10.1594/PANGAEA.911904","CC-BY-4.0",
                             dat.list$PS118[,c('Depth.water..m.')],NA,NA,NA,dat.list$PS118[,c('Area')])

csv.list$TAN0802 <- cbind(dat.list$TAN0802[,c('FileName','Filename.standardised')],
                             "TAN0802",dat.list$TAN0802[,c('transectID','GPS_lon','GPS_lat','time')],
                             NA,NA,
                             NA,NA,
                             dat.list$TAN0802[,c('depth')],NA,NA,NA,NA)

csv.list$TAN1802 <- cbind(dat.list$TAN1802[,c('FileName','Filename.standardised')],
                             "TAN1802",dat.list$TAN1802[,c('transectID','GPS_lon','GPS_lat','time')],
                             NA,NA,
                             NA,NA,
                             dat.list$TAN1802[,c('depth')],NA,NA,NA,NA)

csv.list$TAN1901 <- cbind(dat.list$TAN1901[,c('FileName','Filename.standardised')],
                             "TAN1901",dat.list$TAN1901[,c('transectID','GPS_lon','GPS_lat','time')],
                             NA,NA,
                             NA,NA,
                             dat.list$TAN1901[,c('depth')],NA,NA,NA,NA)

csv.list$NBP1402 <- cbind(dat.list$NBP1402[,c('FileName','Filename.standardised')],
                             "NBP1402",dat.list$NBP1402[,c('transectID','GPS_lon','GPS_lat','time')],
                             NA,NA,
                             "www.usap-dc.org/view/dataset/601310","CC-BY-NC 4.0",
                             dat.list$NBP1402[,c('Depth')],NA,NA,4.8,NA)

csv.list$NBP1502 <- cbind(dat.list$NBP1502[,c('FileName','Filename.standardised')],
                             "NBP1502",dat.list$NBP1502[,c('transectID','GPS_lon','GPS_lat')],
                             NA,NA,NA,
                             "doi.org/10.15784/601182","CC-BY-NC 4.0",
                             NA,NA,NA,NA,NA)

csv.list$AA2011 <- cbind(dat.list$AA2011[,c('Filename','Filename.standardised')],
                             "AA2011",dat.list$AA2011[,c('transectID','lon','lat','DateTime')],
                             NA,NA,
                             "doi.org/doi:10.4225/15/59acda196ccfb","CC-BY-4.0",
                             dat.list$AA2011[,c('altimeter')],NA,NA,NA,NA)

csv.list$JR262 <- cbind(dat.list$JR262[,c('filename','Filename.standardised')],
                             "JR262",dat.list$JR262[,c('transectID','lon','lat','time')],
                             NA,NA,
                             NA,NA,
                             dat.list$JR262[,c('depth')],NA,NA,0.51,NA)

csv.list$JR15005 <- cbind(dat.list$JR15005[,c('filename','Filename.standardised')],
                             "JR15005",dat.list$JR15005[,c('transectID','lon','lat','time')],
                             NA,NA,
                             NA,NA,
                             dat.list$JR15005[,c('depth')],NA,NA,0.51,NA)

csv.list$JR17001 <- cbind(dat.list$JR17001[,c('filename','Filename.standardised')],
                             "JR17001",dat.list$JR17001[,c('transectID','lon','lat','time')],
                             NA,NA,
                             NA,NA,
                             dat.list$JR17001[,c('depth')],NA,NA,0.51,NA)

csv.list$JR17003 <- cbind(dat.list$JR17003[,c('filename','Filename.standardised')],
                             "JR17003",dat.list$JR17003[,c('transectID','lon','lat','time')],
                             NA,NA,
                             "doi.org/10.5285/48dcef16-6719-45e5-a335-3a97f099e451","http://www.nationalarchives.gov.uk/doc/open-government-licence/version/3/",
                             dat.list$JR17003[,c('depth')],NA,NA,0.51,NA)

csv.list$LMG1311 <- cbind(dat.list$LMG1311[,c('filename','Filename.standardised')],
                             "LMG1311",dat.list$LMG1311[,c('transectID','lon','lat','DateTime')],
                             NA,NA,
                             "doi.org/10.15784/601311","CC-BY-NC 4.0",
                             NA,dat.list$LMG1311[,c('depth.start','depth.end')],NA,NA)

for(i in 1:length(csv.list)){
  names(csv.list[[i]]) <- csv.names
}

##### some surveys need special attention
## PS06, PS14, PS18, PS61 all have a separate doi/link for each transect
unique(csv.list$PS06$Transect.ID)
csv.list$PS06$Source.link[csv.list$PS06$Transect.ID=="289"] <- "doi.pangaea.de/10.1594/PANGAEA.713324"
csv.list$PS06$Source.link[csv.list$PS06$Transect.ID=="292"] <- "doi.pangaea.de/10.1594/PANGAEA.713325"
unique(csv.list$PS14$Transect.ID)
csv.list$PS14$Source.link[csv.list$PS14$Transect.ID=="245"] <- "doi.pangaea.de/10.1594/PANGAEA.691559"
csv.list$PS14$Source.link[csv.list$PS14$Transect.ID=="246"] <- "doi.pangaea.de/10.1594/PANGAEA.691560"
csv.list$PS14$Source.link[csv.list$PS14$Transect.ID=="250"] <- "doi.pangaea.de/10.1594/PANGAEA.691561"
csv.list$PS14$Source.link[csv.list$PS14$Transect.ID=="256"] <- "doi.pangaea.de/10.1594/PANGAEA.691562"
csv.list$PS14$Source.link[csv.list$PS14$Transect.ID=="259"] <- "doi.pangaea.de/10.1594/PANGAEA.691563"
csv.list$PS14$Source.link[csv.list$PS14$Transect.ID=="260"] <- "doi.pangaea.de/10.1594/PANGAEA.691564"
csv.list$PS14$Source.link[csv.list$PS14$Transect.ID=="261"] <- "doi.pangaea.de/10.1594/PANGAEA.691565"
csv.list$PS14$Source.link[csv.list$PS14$Transect.ID=="270"] <- "doi.pangaea.de/10.1594/PANGAEA.691566"
csv.list$PS14$Source.link[csv.list$PS14$Transect.ID=="275"] <- "doi.pangaea.de/10.1594/PANGAEA.691568"
csv.list$PS14$Source.link[csv.list$PS14$Transect.ID=="276"] <- "doi.pangaea.de/10.1594/PANGAEA.691569"
csv.list$PS14$Source.link[csv.list$PS14$Transect.ID=="277"] <- "doi.pangaea.de/10.1594/PANGAEA.691570"
csv.list$PS14$Source.link[csv.list$PS14$Transect.ID=="278-2"] <- "doi.pangaea.de/10.1594/PANGAEA.691571"
csv.list$PS14$Source.link[csv.list$PS14$Transect.ID=="278-3"] <- "doi.pangaea.de/10.1594/PANGAEA.691572"
csv.list$PS14$Source.link[csv.list$PS14$Transect.ID=="280"] <- "doi.pangaea.de/10.1594/PANGAEA.691573"
csv.list$PS14$Source.link[csv.list$PS14$Transect.ID=="285"] <- "doi.pangaea.de/10.1594/PANGAEA.691574"
csv.list$PS14$Source.link[csv.list$PS14$Transect.ID=="293"] <- "doi.pangaea.de/10.1594/PANGAEA.691575"
csv.list$PS14$Source.link[csv.list$PS14$Transect.ID=="294"] <- "doi.pangaea.de/10.1594/PANGAEA.691576"
csv.list$PS14$Source.link[csv.list$PS14$Transect.ID=="304"] <- "doi.pangaea.de/10.1594/PANGAEA.691577"
csv.list$PS14$Source.link[csv.list$PS14$Transect.ID=="305"] <- "doi.pangaea.de/10.1594/PANGAEA.691578"
csv.list$PS14$Source.link[csv.list$PS14$Transect.ID=="306"] <- "doi.pangaea.de/10.1594/PANGAEA.691579"
csv.list$PS14$Source.link[csv.list$PS14$Transect.ID=="307"] <- "doi.pangaea.de/10.1594/PANGAEA.691580"
unique(csv.list$PS18$Transect.ID)
csv.list$PS18$Source.link[csv.list$PS18$Transect.ID=="126-2"] <- "doi.pangaea.de/10.1594/PANGAEA.667008"
csv.list$PS18$Source.link[csv.list$PS18$Transect.ID=="126-6"] <- "doi.pangaea.de/10.1594/PANGAEA.667009"
csv.list$PS18$Source.link[csv.list$PS18$Transect.ID=="129-1"] <- "doi.pangaea.de/10.1594/PANGAEA.667010"
csv.list$PS18$Source.link[csv.list$PS18$Transect.ID=="131-1"] <- "doi.pangaea.de/10.1594/PANGAEA.667011"
csv.list$PS18$Source.link[csv.list$PS18$Transect.ID=="134-1"] <- "doi.pangaea.de/10.1594/PANGAEA.667012"
csv.list$PS18$Source.link[csv.list$PS18$Transect.ID=="135-1"] <- "doi.pangaea.de/10.1594/PANGAEA.667013"
csv.list$PS18$Source.link[csv.list$PS18$Transect.ID=="136-4"] <- "doi.pangaea.de/10.1594/PANGAEA.667014"
csv.list$PS18$Source.link[csv.list$PS18$Transect.ID=="160-2"] <- "doi.pangaea.de/10.1594/PANGAEA.667015"
csv.list$PS18$Source.link[csv.list$PS18$Transect.ID=="165-1"] <- "doi.pangaea.de/10.1594/PANGAEA.667017"
csv.list$PS18$Source.link[csv.list$PS18$Transect.ID=="169-2"] <- "doi.pangaea.de/10.1594/PANGAEA.667018"
csv.list$PS18$Source.link[csv.list$PS18$Transect.ID=="171-1"] <- "doi.pangaea.de/10.1594/PANGAEA.667019"
csv.list$PS18$Source.link[csv.list$PS18$Transect.ID=="174-1"] <- "doi.pangaea.de/10.1594/PANGAEA.667021"
csv.list$PS18$Source.link[csv.list$PS18$Transect.ID=="175-2"] <- "doi.pangaea.de/10.1594/PANGAEA.667022"
csv.list$PS18$Source.link[csv.list$PS18$Transect.ID=="179-1"] <- "doi.pangaea.de/10.1594/PANGAEA.667023"
csv.list$PS18$Source.link[csv.list$PS18$Transect.ID=="180-2"] <- "doi.pangaea.de/10.1594/PANGAEA.667024"
csv.list$PS18$Source.link[csv.list$PS18$Transect.ID=="182-1"] <- "doi.pangaea.de/10.1594/PANGAEA.667025"
csv.list$PS18$Source.link[csv.list$PS18$Transect.ID=="189-2"] <- "doi.pangaea.de/10.1594/PANGAEA.667026"
csv.list$PS18$Source.link[csv.list$PS18$Transect.ID=="206-2"] <- "doi.pangaea.de/10.1594/PANGAEA.667027"
csv.list$PS18$Source.link[csv.list$PS18$Transect.ID=="211-1"] <- "doi.pangaea.de/10.1594/PANGAEA.667029"
csv.list$PS18$Source.link[csv.list$PS18$Transect.ID=="212-2"] <- "doi.pangaea.de/10.1594/PANGAEA.667030"
csv.list$PS18$Source.link[csv.list$PS18$Transect.ID=="220-1"] <- "doi.pangaea.de/10.1594/PANGAEA.667031"
csv.list$PS18$Source.link[csv.list$PS18$Transect.ID=="222-1"] <- "doi.pangaea.de/10.1594/PANGAEA.667032"
unique(csv.list$PS61$Transect.ID)
csv.list$PS61$Source.link[csv.list$PS61$Transect.ID=="235-1"] <- "doi.org/10.1594/PANGAEA.220746"
csv.list$PS61$Source.link[csv.list$PS61$Transect.ID=="249-1"] <- "doi.org/10.1594/PANGAEA.220747"



#### CRS survey information 
## Anvord, Flanders and Barilla in 2010 during NBP10-01 (https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5214585/)
unique(csv.list$CRS$Transect.ID)
csv.list$CRS$Date[csv.list$CRS$Transect.ID=="1289"] <- ymd("2010:01:20")
csv.list$CRS$Date[csv.list$CRS$Transect.ID=="1290"] <- ymd("2010:01:20")
csv.list$CRS$Date[csv.list$CRS$Transect.ID=="1283"] <- ymd("2010:01:18")
csv.list$CRS$Date[csv.list$CRS$Transect.ID=="1284"] <- ymd("2010:01:19")
csv.list$CRS$Date[csv.list$CRS$Transect.ID=="1337"] <- ymd("2010:02:20")
csv.list$CRS$Date[csv.list$CRS$Transect.ID=="1338"] <- ymd("2010:02:20")
csv.list$CRS$Date[csv.list$CRS$Transect.ID=="1285"] <- ymd("2010:01:19")
csv.list$CRS$Date[csv.list$CRS$Transect.ID=="1286"] <- ymd("2010:01:19")
csv.list$CRS$Date[csv.list$CRS$Transect.ID=="1281"] <- ymd("2010:01:18")
csv.list$CRS$Date[csv.list$CRS$Transect.ID=="1282"] <- ymd("2010:01:18")
csv.list$CRS$Date[csv.list$CRS$Transect.ID=="1279"] <- ymd("2010:01:17")
csv.list$CRS$Date[csv.list$CRS$Transect.ID=="1280"] <- ymd("2010:01:17")
csv.list$CRS$Date[csv.list$CRS$Transect.ID=="1276"] <- ymd("2010:01:14")
csv.list$CRS$Date[csv.list$CRS$Transect.ID=="1278"] <- ymd("2010:01:17")
csv.list$CRS$Date[csv.list$CRS$Transect.ID=="1300"] <- ymd("2010:01:27")
csv.list$CRS$Date[csv.list$CRS$Transect.ID=="1295"] <- ymd("2010:01:24")
csv.list$CRS$Date[csv.list$CRS$Transect.ID=="1297"] <- ymd("2010:01:26")
## LMG0902: https://www.marine-geo.org/tools/entry/LMG0902
csv.list$CRS$Date[csv.list$CRS$Transect.ID=="1207"] <- ymd("2009:02:27")
csv.list$CRS$Date[csv.list$CRS$Transect.ID=="1208"] <- ymd("2009:02:27")
csv.list$CRS$Date[csv.list$CRS$Transect.ID=="1217"] <- ymd("2009:02:28")
csv.list$CRS$Date[csv.list$CRS$Transect.ID=="1219"] <- ymd("2009:02:28")
csv.list$CRS$Date[csv.list$CRS$Transect.ID=="1255"] <- ymd("2009:03:05")
csv.list$CRS$Date[csv.list$CRS$Transect.ID=="1267"] <- ymd("2009:03:06")

## still missing:
# 1069
# 1072
# 1091
# 1103
# 1130
# 1132
# 1315
# 1320
# 1323
# 1324
# 1325



```

```{r}
for(i in 1:19){
  message(names(csv.list)[i])
  print(head(csv.list[[i]]))
}
```
