---
title: "ROMS FAM 10km floats"
author: "Jan Jansen"
output: html_document
keep_md: yes
---

```{r setup, include=FALSE}
## Evaluate ROMS outputs
library(raster)
library(colorspace)
library(viridis)
library(ncdf4)
library(raadtools)
library(rgdal)

## data-directories
env.raw <- "E:/science/data_environmental/raw/"
env.dir <- "E:/science/data_environmental/"
env.derived <- "C:/Users/jjansen/Desktop/science/data_environmental/derived/"
AAD_dir <- paste0(env.raw,"accessed_through_R")
ARC_data_dir <- "C:/Users/jjansen/Desktop/science/SouthernOceanBiodiversityMapping/ARC_Data/"
ant.proj <- "+proj=stere +lat_0=-90 +lat_ts=-71 +lon_0=0 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs"
```

```{r, results="hide"}
## load projected coastline for plotting
load(paste0(env.derived,"Circumpolar_Coastline.Rdata"))

## ROMS-runs:
data.dat100 <- paste0(env.dir,"Circumpolar_ROMS/10km_outputs/output_sed_float_test3/")
# data.dat <- paste0("Circumpolar_ROMS/10km_outputs/sed_test1/")
# data.dat <- paste0(env.dir,"Circumpolar_ROMS/10km_outputs/sed_test5/")

## load lon/lat information from ROMS-grid
grd10k_nc <- nc_open(paste0(env.raw,"waom10extend_grd.nc"))
lon_rho <- ncvar_get(grd10k_nc, varid="lon_rho")
lat_rho <- ncvar_get(grd10k_nc, varid="lat_rho")

#### Prepare empty rasters to assign correct projected values to
roms.coords.proj <- project(cbind(c(lon_rho), c(lat_rho)), proj=ant.proj)
x.range <- c(min(roms.coords.proj[,1])-5000,max(roms.coords.proj[,1])+5000)
y.range <- c(min(roms.coords.proj[,2])-5000,max(roms.coords.proj[,2])+5000)
empty.roms.ra <- raster(extent(c(x.range,y.range)), crs=ant.proj, resolution=10000)

#### load other ROMS data (currents & depth)
#depth
h <- raster(paste0(data.dat100,"ocean_avg_0003.nc"), varname="h", level=1)
#seafloor currents (seafloor-layer is 1)
u.raw <- brick(paste0(data.dat100,"ocean_his_0003.nc"), varname="u", level=1)
v.raw <- brick(paste0(data.dat100,"ocean_his_0003.nc"), varname="v", level=1)
#seasurface currents (surface-layer is 31)
u_31.raw <- brick(paste0(data.dat100,"ocean_avg_0003.nc"), varname="u", level=31)
v_31.raw <- brick(paste0(data.dat100,"ocean_avg_0003.nc"), varname="v", level=31)
## sum up monthly values for a climatology (THIS SHOULD BE DONE ON HIGH RESOLUTION HISTORY FILES)
u.sum <- sum(u.raw)
v.sum <- sum(v.raw)
u.sum.abs <- sum(abs(u.raw))
v.sum.abs <- sum(abs(v.raw))
u31.sum <- sum(u_31.raw)
v31.sum <- sum(v_31.raw)
## extract current speeds at rho-points where depth is defined
## (in ROMS they are all at different locations):
## u has one less column than the grid, so all coordinates need to be moved half a cell to the right for the grids to match up
coord.grd.u <- coordinates(h)
coord.grd.u[,1] <- coordinates(h)[,1]-0.5
## v has one less row than the grid, so all coordinates need to be moved half a cell up for the grids to match up
coord.grd.v <- coordinates(h)
coord.grd.v[,2] <- coordinates(h)[,2]-0.5
## now extract values at the rho-points and interpolate (because they are 2km away from the nearest original point), and place into projected raster
u <- v <- u.abs <- v.abs <- u31 <- v31 <- empty.roms.ra
u[] <- extract(u.sum, coord.grd.u, method="bilinear")
v[] <- extract(v.sum, coord.grd.v, method="bilinear")
u.abs[] <- extract(u.sum.abs, coord.grd.u, method="bilinear")
v.abs[] <- extract(v.sum.abs, coord.grd.v, method="bilinear")
u31[] <- extract(u31.sum, coord.grd.u, method="bilinear")
v31[] <- extract(v31.sum, coord.grd.v, method="bilinear")
#seasurface current speeds
uv_31 <- sqrt(u31^2+v31^2)
#temporal mean seafloor current speed
mean.uv <- sqrt(u^2+v^2)
#absolute mean seafloor current speed
abs.uv <- sqrt(u.abs^2+v.abs^2)
#residual seafloor current speed
res.uv <- abs.uv-mean.uv
## remove inland values for depth
h2 <- empty.roms.ra
h2[] <- h[]
h2[is.na(mean.uv)] <- NA
```

########## Overview plots - Ocean currents ##########

```{r, fig.width=10, fig.height=10}
## ocean current data
#cols <- rev(terrain.colors(96))
cols <- rev(magma(96)) #12, 24, 48, 96
breaks <- seq(0,6,length.out=97)
par(mfrow=c(2,2))
plot(mean.uv, breaks=breaks, col=cols, main="seafloor currents - mean")
plot(abs.uv, breaks=breaks, col=cols, main="seafloor currents - absolute speed")
plot(res.uv, breaks=breaks, col=cols, main="seafloor currents - residual")
plot(uv_31, breaks=breaks, col=cols, main="seasurface currents")
```


```{r, fig.width=10, fig.height=8}
## ocean current data in the Mertz region
cols <- rev(magma(84)) #12, 24, 48, 96
breaks <- seq(0,2.1,length.out=85)
xlim <- c(1300000,1900000)
ylim <- c(-2300000,-1900000)
par(mfrow=c(2,2))
plot(mean.uv, breaks=breaks, col=cols, main="seafloor currents - mean", xlim=xlim, ylim=ylim)
#plot(coast.proj, add=TRUE)
plot(abs.uv, breaks=breaks, col=cols, main="seafloor currents - absolute speed", xlim=xlim, ylim=ylim)
#plot(coast.proj, add=TRUE)
plot(res.uv, breaks=breaks, col=cols, main="seafloor currents - residual", xlim=xlim, ylim=ylim)
#plot(coast.proj, add=TRUE)
plot(h2, col=cols, main="depth", xlim=xlim, ylim=ylim)
#plot(coast.proj, add=TRUE)
```

########## EVALUATE particle-tracks ##########

```{r, results="hide"}
## load trajectories
#it's being stored in a rather weird way...
#if you read in as a raster, the rows are the time-steps
#everything between 153-183 is NA for some reason
flts.z100 <- brick(paste0(data.dat100,"ocean_flt.nc"), varname="depth", level=1)
#### translate grid points into projected lon/lats
## one cell is 10000m and the grid starts at -3000000/-3000000, so need to start from there and multiply by 10000
grd.x100 <- -3000000+brick(paste0(data.dat100,"ocean_flt.nc"), varname="Xgrid", level=1)*10000
grd.y100 <- -3000000+brick(paste0(data.dat100,"ocean_flt.nc"), varname="Ygrid", level=1)*10000
# flts.z200 <- brick(paste0(data.dat200,"ocean_flt.nc"), varname="depth", level=1)
# grd.x200 <- brick(paste0(data.dat200,"ocean_flt.nc"), varname="Xgrid", level=1)
# grd.y200 <- brick(paste0(data.dat200,"ocean_flt.nc"), varname="Ygrid", level=1)
# flts.z400 <- brick(paste0(data.dat400,"ocean_flt.nc"), varname="depth", level=1)
# grd.x400 <- brick(paste0(data.dat400,"ocean_flt.nc"), varname="Xgrid", level=1)
# grd.y400 <- brick(paste0(data.dat400,"ocean_flt.nc"), varname="Ygrid", level=1)

#### particle trajectories (only use a subset of 1000 traces)
set.seed(1)
s <- sample(1:length(which(!is.na(grd.x100[151,]))), 1000)

start.vals <- cbind(grd.x100[152,s], grd.y100[152,s])
traj.vals <- cbind(grd.x100[1:151,s], grd.y100[1:151,s])
traj.depthvals <- flts.z100[1:151,s]

traj.1 <- which(traj.depthvals>=-500)
traj.2 <- which(traj.depthvals<(-500))
traj.3 <- which(traj.depthvals<(-1000))
traj.4 <- which(traj.depthvals<(-2000))
col <- c("purple","lightskyblue","blue","black")
```

```{r, fig.width=8, fig.height=10}
## trajectories of 1000 floats (MERTZ)
xlim <- c(1300000,1900000)
ylim <- c(-2300000,-1900000)
cex=0.2
pch=16
cols <- rev(magma(74))
breaks <- seq(0,3.7,length.out=75)
par(mfrow=c(2,1), mar=c(2,3,3,1), oma=c(0,0,0,0))
plot(mean.uv, xlim=xlim, ylim=ylim, col=cols, breaks=breaks, main="trajectories - 100m/day - seafloor currents")
# points(start.vals, pch=16, col="blue")
# points(traj.vals, pch=16, col="blue", cex=0.2)
points(start.vals, pch=pch, col=col[1])
points(traj.vals[traj.1,], pch=pch, cex=cex, col=col[1])
points(traj.vals[traj.2,], pch=pch, cex=cex, col=col[2])
points(traj.vals[traj.3,], pch=pch, cex=cex, col=col[3])
points(traj.vals[traj.4,], pch=pch, cex=cex, col=col[4])
legend("topleft", legend=c("0 to 500m","500 to 1000m","1000 to 2000m",">2000m"), col=col, pch =15, bty="n")

plot(uv_31, xlim=xlim, ylim=ylim, col=cols, breaks=breaks, main="trajectories - 100m/day - surface currents")
# points(start.vals, pch=16, col="blue")
# points(traj.vals, pch=16, col="blue", cex=0.2)
points(start.vals, pch=pch, col=col[1])
points(traj.vals[traj.1,], pch=pch, cex=cex, col=col[1])
points(traj.vals[traj.2,], pch=pch, cex=cex, col=col[2])
points(traj.vals[traj.3,], pch=pch, cex=cex, col=col[3])
points(traj.vals[traj.4,], pch=pch, cex=cex, col=col[4])
```

```{r, fig.width=7, fig.height=10}
## trajectories of 1000 floats (Ross Sea)
xlim <- c(-1000000,600000)
ylim <- c(-2300000,-1000000)
par(mfrow=c(2,1), mar=c(2,3,3,1), oma=c(0,0,0,0))
plot(mean.uv, xlim=xlim, ylim=ylim, col=cols, breaks=breaks, main="trajectories - 100m/day - seafloor currents")
# points(start.vals, pch=16, col="blue")
# points(traj.vals, pch=16, col="blue", cex=0.2)
points(start.vals, pch=16, col=col[1])
points(traj.vals[traj.1,], pch=16, cex=0.2, col=col[1])
points(traj.vals[traj.2,], pch=16, cex=0.2, col=col[2])
points(traj.vals[traj.3,], pch=16, cex=0.2, col=col[3])
points(traj.vals[traj.4,], pch=16, cex=0.2, col=col[4])
plot(uv_31, xlim=xlim, ylim=ylim, col=cols, breaks=breaks, main="trajectories - 100m/day - surface currents")
# points(start.vals, pch=16, col="blue")
# points(traj.vals, pch=16, col="blue", cex=0.2)
points(start.vals, pch=16, col=col[1])
points(traj.vals[traj.1,], pch=16, cex=0.2, col=col[1])
points(traj.vals[traj.2,], pch=16, cex=0.2, col=col[2])
points(traj.vals[traj.3,], pch=16, cex=0.2, col=col[3])
points(traj.vals[traj.4,], pch=16, cex=0.2, col=col[4])
```

```{r, fig.width=10, fig.height=9}
par(mfrow=c(1,1), mar=c(2,1,3,0), oma=c(0,0,0,0))
plot(mean.uv, col=cols, breaks=breaks)
points(cbind(grd.x100[91,], grd.y100[91,]), pch=16, col="blue", cex=0.2)
```

```{r, fig.width=10, fig.height=10, eval=FALSE}
par(mfrow=c(3,3), mar=c(2,1,3,0), oma=c(0,0,0,0))
xlim <- c(-3000000,-1000000)
ylim <- c(1000000,3000000)
plot(mean.uv, xlim=xlim, ylim=ylim, col=cols, breaks=breaks)
points(cbind(grd.x100[91,], grd.y100[91,]), pch=16, col="blue", cex=0.2)
xlim <- c(-1000000,1000000)
ylim <- c(1000000,3000000)
plot(mean.uv, xlim=xlim, ylim=ylim, col=cols, breaks=breaks)
points(cbind(grd.x100[91,], grd.y100[91,]), pch=16, col="blue", cex=0.2)
xlim <- c(1000000,3000000)
ylim <- c(1000000,3000000)
plot(mean.uv, xlim=xlim, ylim=ylim, col=cols, breaks=breaks)
points(cbind(grd.x100[91,], grd.y100[91,]), pch=16, col="blue", cex=0.2)
xlim <- c(-3000000,-1000000)
ylim <- c(-1000000,1000000)
plot(mean.uv, xlim=xlim, ylim=ylim, col=cols, breaks=breaks)
points(cbind(grd.x100[91,], grd.y100[91,]), pch=16, col="blue", cex=0.2)
plot.new()
xlim <- c(1000000,3000000)
ylim <- c(-1000000,1000000)
plot(mean.uv, xlim=xlim, ylim=ylim, col=cols, breaks=breaks)
points(cbind(grd.x100[91,], grd.y100[91,]), pch=16, col="blue", cex=0.2)
xlim <- c(-3000000,-1000000)
ylim <- c(-3000000,-1000000)
plot(mean.uv, xlim=xlim, ylim=ylim, col=cols, breaks=breaks)
points(cbind(grd.x100[91,], grd.y100[91,]), pch=16, col="blue", cex=0.2)
xlim <- c(-1000000,1000000)
ylim <- c(-3000000,-1000000)
plot(mean.uv, xlim=xlim, ylim=ylim, col=cols, breaks=breaks)
points(cbind(grd.x100[91,], grd.y100[91,]), pch=16, col="blue", cex=0.2)
xlim <- c(1000000,3000000)
ylim <- c(-3000000,-1000000)
plot(mean.uv, xlim=xlim, ylim=ylim, col=cols, breaks=breaks)
points(cbind(grd.x100[91,], grd.y100[91,]), pch=16, col="blue", cex=0.2)
```

```{r, fig.width=10, fig.height=9}
## end-points of all floats
par(mfrow=c(2,2), mar=c(2,1,3,0), oma=c(1,1,1,1))
#after 1 months
plot(mean.uv, xlim=xlim, ylim=ylim, col=cols, breaks=breaks, main="end-points - 1 month - 100m/day", legend=FALSE)
points(cbind(grd.x100[121,], grd.y100[121,]), pch=16, col="blue", cex=0.2)
#after 2 months
plot(mean.uv, xlim=xlim, ylim=ylim, col=cols, breaks=breaks, main="end-points - 2 month  - 100m/day")
points(cbind(grd.x100[91,], grd.y100[91,]), pch=16, col="blue", cex=0.2)
#after 3 months
plot(mean.uv, xlim=xlim, ylim=ylim, col=cols, breaks=breaks, main="end-points - 3 month  - 100m/day", legend=FALSE)
points(cbind(grd.x100[61,], grd.y100[61,]), pch=16, col="blue", cex=0.2)
#after 4 months
plot(mean.uv, xlim=xlim, ylim=ylim, col=cols, breaks=breaks, main="end-points - 4 month  - 100m/day")
points(cbind(grd.x100[31,], grd.y100[31,]), pch=16, col="blue", cex=0.2)
```

## how far have particles travelled?

```{r}
## should be able to just substract positions from another at different points in time and calculate the distance
## but, some values are off, why? (thousands of kilometers travelled in 1 day...?) -. some boundary effects happening
weird <- which(sqrt((grd.x100[121,]-grd.x100[122,])^2+(grd.y100[121,]-grd.y100[122,])^2)>200000)
# par(mfrow=c(2,1))
# plot(mean.uv, col=cols, breaks=breaks, main="end-points - 1 month - 100m/day")
# points(cbind(grd.x100[122,weird], grd.y100[122,weird]), pch=16, col="blue", cex=1)
# plot(mean.uv, col=cols, breaks=breaks, main="end-points - 1 month - 100m/day")
# points(cbind(grd.x100[121,weird], grd.y100[121,weird]), pch=16, col="blue", cex=1)

## best approach is to just disregard the boundary one in the plotting
## movement since start:
move.s.1st <- sqrt((grd.x100[121,]-grd.x100[151,])^2+(grd.y100[121,]-grd.y100[151,])^2)
move.s.2nd <- sqrt((grd.x100[91,]-grd.x100[151,])^2+(grd.y100[91,]-grd.y100[151,])^2)
move.s.3rd <- sqrt((grd.x100[61,]-grd.x100[151,])^2+(grd.y100[61,]-grd.y100[151,])^2)
move.s.4th <- sqrt((grd.x100[31,]-grd.x100[151,])^2+(grd.y100[31,]-grd.y100[151,])^2)

ms1.1 <- which(move.s.1st >= 0)
ms1.2 <- which(move.s.1st >= 10000)
ms1.3 <- which(move.s.1st >= 100000)
ms1.4 <- which(move.s.1st >= 1000000)

ms2.1 <- which(move.s.2nd >= 0)
ms2.2 <- which(move.s.2nd >= 10000)
ms2.3 <- which(move.s.2nd >= 100000)
ms2.4 <- which(move.s.2nd >= 1000000)

ms4.1 <- which(move.s.4th >= 0)
ms4.2 <- which(move.s.4th >= 10000)
ms4.3 <- which(move.s.4th >= 100000)
ms4.4 <- which(move.s.4th >= 1000000)

## movement in 1 month:
move.m.1st <- sqrt((grd.x100[121,]-grd.x100[151,])^2+(grd.y100[121,]-grd.y100[151,])^2)
move.m.2nd <- sqrt((grd.x100[91,]-grd.x100[121,])^2+(grd.y100[91,]-grd.y100[121,])^2)
move.m.3rd <- sqrt((grd.x100[61,]-grd.x100[91,])^2+(grd.y100[61,]-grd.y100[91,])^2)
move.m.4th <- sqrt((grd.x100[31,]-grd.x100[61,])^2+(grd.y100[31,]-grd.y100[61,])^2)

mm1.1 <- which(move.m.1st >= 0)
mm1.2 <- which(move.m.1st >= 5000)
mm1.3 <- which(move.m.1st >= 10000)
mm1.4 <- which(move.m.1st >= 50000)

```

## endpoints coloured by how far they travelled since the start

```{r, fig.width=10, fig.height=9}
## after how many months?
x <- 121 ## 1 month
y <- list(ms1.1, ms1.2, ms1.3, ms1.4)

col <- c("black", "grey40", "grey60","grey80")
par(mfrow=c(1,1), mar=c(2,1,3,0), oma=c(0,0,0,0))
#after 1 months
plot(mean.uv, col=cols, breaks=breaks, main="end-points - 1 month - 100m/day")
points(cbind(grd.x100[x,y[[1]]], grd.y100[x,y[[1]]]), pch=16, col=col[1], cex=0.2) ## less than 10km
points(cbind(grd.x100[x,y[[2]]], grd.y100[x,y[[2]]]), pch=16, col=col[2], cex=0.2) ## less than 100km
points(cbind(grd.x100[x,y[[3]]], grd.y100[x,y[[3]]]), pch=16, col=col[3], cex=0.2) ## less than 1000km
points(cbind(grd.x100[x,y[[4]]], grd.y100[x,y[[4]]]), pch=16, col=col[4], cex=0.2) ## more than 1000km
legend("center", legend=c("0 to 10km","10 to 100km","100 to 1000km",">1000km"), col=col, pch =15, bty="n")
```

```{r, fig.width=10, fig.height=9}
## after how many months?
# x <- 121 ## 1 month
# y <- list(ms1.1, ms1.2, ms1.3, ms1.4)
# x <- 91 ## 2 month
# y <- list(ms2.1,ms2.2,ms2.3,ms2.4)
# x <- 61 ## 3 month
x <- 31 ## 4 month
y <- list(ms4.1,ms4.2,ms4.3,ms4.4)

col <- c("black", "grey40", "grey60","grey80")
par(mfrow=c(1,1), mar=c(2,1,3,0), oma=c(0,0,0,0))
#after 1 months
plot(mean.uv, col=cols, breaks=breaks, main="end-points - 4 month - 100m/day")
points(cbind(grd.x100[x,y[[1]]], grd.y100[x,y[[1]]]), pch=16, col=col[1], cex=0.2) ## less than 10km
points(cbind(grd.x100[x,y[[2]]], grd.y100[x,y[[2]]]), pch=16, col=col[2], cex=0.2) ## less than 100km
points(cbind(grd.x100[x,y[[3]]], grd.y100[x,y[[3]]]), pch=16, col=col[3], cex=0.2) ## less than 1000km
points(cbind(grd.x100[x,y[[4]]], grd.y100[x,y[[4]]]), pch=16, col=col[4], cex=0.2) ## more than 1000km
legend("center", legend=c("0 to 10km","10 to 100km","100 to 1000km",">1000km"), col=col, pch =15, bty="n")
```

```{r, fig.width=10, fig.height=9.5}
xlim=c(-500000,500000)
ylim=c(-2100000,-1150000)

col <- c("black", "grey40", "grey60","grey80")
par(mfrow=c(1,1), mar=c(2,1,3,0), oma=c(0,0,0,0))
#after 1 months
plot(mean.uv, xlim=xlim, ylim=ylim, col=cols, breaks=breaks, main="end-points - 4 month - 100m/day")
points(cbind(grd.x100[x,y[[1]]], grd.y100[x,y[[1]]]), pch=16, col=col[1], cex=1) ## less than 10km
points(cbind(grd.x100[x,y[[2]]], grd.y100[x,y[[2]]]), pch=16, col=col[2], cex=1) ## less than 100km
points(cbind(grd.x100[x,y[[3]]], grd.y100[x,y[[3]]]), pch=16, col=col[3], cex=1) ## less than 1000km
points(cbind(grd.x100[x,y[[4]]], grd.y100[x,y[[4]]]), pch=16, col=col[4], cex=1) ## more than 1000km
legend("top", legend=c("0 to 10km","10 to 100km","100 to 1000km",">1000km"), col=col, pch =15)
```

## endpoints coloured by how far they travelled in a month

### number and distances travelled TO EACH CELL:

```{r}
## create a dataframe with coordinates and distance traveled, one row for each point
dat.month1 <- data.frame(cbind(grd.x100[121,], grd.y100[121,], move.s.1st))
names(dat.month1) <- c("x","y","distance")

## identify into which cell each particle belongs
dat.month1$cellID <- extract(mean.uv,cbind(dat.month1$x,dat.month1$y), cellnumbers=TRUE)[,1]

## NAs and unique IDs
cellID_nas <- which(is.na(dat.month1$cellID))
cellID_unique <- unique(dat.month1$cellID[-cellID_nas][order(dat.month1$cellID[-cellID_nas])])

## number of particles in each cell
tab.count <- table(dat.month1$cellID)

## summed distances travelled for all particles in each cell
#xtabs(distance~cellID, data=dat.month1)
tab.sum <- with(dat.month1, tapply(distance, list(cellID), FUN=sum))

dat.m1 <- data.frame(cbind(cellID_unique, tab.count, tab.sum))

## set NAs to zero
count.ra <- dist.ra <- dist.ave.ra <- raster(mean.uv)
count.ra[!is.na(mean.uv)] <- 0
dist.ra[!is.na(mean.uv)] <- 0
dist.ave.ra[!is.na(mean.uv)] <- 0

## counts per cell
count.ra[cellID_unique] <- dat.m1$tab.count
## total distance per cell
dist.ra[cellID_unique] <- dat.m1$tab.sum/1000
## mean distance per cell
dist.ave.ra[cellID_unique] <- (dat.m1$tab.sum/1000)/dat.m1$tab.count

```

### number and distances travelled FROM EACH CELL

```{r}
dat.month0 <- data.frame(cbind(grd.x100[151,], grd.y100[151,], move.s.1st))
names(dat.month0) <- c("x","y","distance")

## identify into which cell each particle belongs
dat.month0$cellID <- extract(mean.uv,cbind(dat.month0$x,dat.month0$y), cellnumbers=TRUE)[,1]

## NAs and unique IDs
cellID_nas <- which(is.na(dat.month0$cellID))
cellID_unique <- unique(dat.month0$cellID[-cellID_nas][order(dat.month0$cellID[-cellID_nas])])

## number of particles in each cell
tab.count <- table(dat.month0$cellID)

## summed distances travelled for all particles in each cell
#xtabs(distance~cellID, data=dat.month1)
tab.sum <- with(dat.month0, tapply(distance, list(cellID), FUN=sum))

dat.m0 <- data.frame(cbind(cellID_unique, tab.count, tab.sum))

## set NAs to zero
count0.ra <- dist0.ra <- dist0.ave.ra <- raster(mean.uv)
count0.ra[!is.na(mean.uv)] <- 0
dist0.ra[!is.na(mean.uv)] <- 0
dist0.ave.ra[!is.na(mean.uv)] <- 0

## counts per cell
count0.ra[cellID_unique] <- dat.m0$tab.count
## total distance per cell
dist0.ra[cellID_unique] <- dat.m0$tab.sum/1000
## mean distance per cell
dist0.ave.ra[cellID_unique] <- (dat.m0$tab.sum/1000)/dat.m0$tab.count
```

### Ross Sea only:

```{r, fig.height=7}
xlim=c(-500000,500000)
ylim=c(-2100000,-1150000)
par(mfrow=c(2,2), mar=c(4,3,3,2))
plot(count.ra, xlim=xlim, ylim=ylim, main="counts per cell")
plot(dist.ra, xlim=xlim, ylim=ylim, main="total distance per cell")
plot(dist.ave.ra, xlim=xlim, ylim=ylim, main="sink: average distance to each cell (seafloor)")
plot(dist0.ave.ra, xlim=xlim, ylim=ylim, main="source: average distance from each cell (surface)")
```

```{r, fig.height=6}
par(mfrow=c(1,1), mar=c(4,3,3,2))
plot(dist0.ave.ra-dist.ave.ra, xlim=xlim, ylim=ylim, main="source minus sink")
```

### Mertz only:

```{r, fig.height=6}
xlim <- c(1300000,1900000)
ylim <- c(-2300000,-1900000)
par(mfrow=c(2,2), mar=c(4,3,3,2))
plot(count.ra, xlim=xlim, ylim=ylim, main="counts per cell")
plot(dist.ra, xlim=xlim, ylim=ylim, main="total distance per cell")
plot(dist.ave.ra, xlim=xlim, ylim=ylim, main="sink: average distance to each cell (seafloor)")
plot(dist0.ave.ra, xlim=xlim, ylim=ylim, main="source: average distance from each cell (surface)")
```

```{r, fig.height=5}
par(mfrow=c(1,1), mar=c(4,3,3,2))
plot(dist0.ave.ra-dist.ave.ra, xlim=xlim, ylim=ylim, main="source minus sink")
```


```{r, fig.width=10, fig.height=9, eval=FALSE}
## end-points of all floats
par(mfrow=c(2,2), mar=c(2,1,3,0), oma=c(0,0,0,0))
#after 1 months
plot(mean.uv, xlim=xlim, ylim=ylim, col=cols, breaks=breaks, main="end-points - 1 month - 100m/day")
points(cbind(grd.x100[121,mm1.1], grd.y100[121,mm1.1]), pch=16, col=col[1], cex=0.2)
points(cbind(grd.x100[121,mm1.2], grd.y100[121,mm1.2]), pch=16, col=col[2], cex=0.2)
points(cbind(grd.x100[121,mm1.3], grd.y100[121,mm1.3]), pch=16, col=col[3], cex=0.2)
points(cbind(grd.x100[121,mm1.4], grd.y100[121,mm1.4]), pch=16, col=col[4], cex=0.2)

#after 2 months
plot(mean.uv, xlim=xlim, ylim=ylim, col=cols, breaks=breaks, main="end-points - 2 month  - 100m/day")
points(cbind(grd.x100[91,], grd.y100[91,]), pch=16, col="blue", cex=0.2)
#after 3 months
plot(mean.uv, xlim=xlim, ylim=ylim, col=cols, breaks=breaks, main="end-points - 3 month  - 100m/day")
points(cbind(grd.x100[61,], grd.y100[61,]), pch=16, col="blue", cex=0.2)
#after 4 months
plot(mean.uv, xlim=xlim, ylim=ylim, col=cols, breaks=breaks, main="end-points - 4 month  - 100m/day")
points(cbind(grd.x100[31,], grd.y100[31,]), pch=16, col="blue", cex=0.2)
```

```{r}
########## TODO: 
## calculate movement speed using changes in m in grid.x and grid.y per day

## plot depth against movement speed

## 3D plot of particle subset

```


# Analysing particle pathways (This chunk takes about 20min to run)

```{r, eval=FALSE}
## x and y locations are stored in "grd.x" and "grd.y" as time-steps (rows) and individual particles (columns), and their depths in "flts.z"
x.raw1 <- matrix(NA, nrow=152, ncol=ncol(grd.x100))
y.raw1 <- x.raw1
z.raw1 <- x.raw1
for(i in 1:152){
  x.raw1[i,] <- grd.x100[i,,]
  y.raw1[i,] <- grd.y100[i,,]
  z.raw1[i,] <- flts.z100[i,,]
}

## set points that move into the bottom corner to NA!!! (x.raw=-3000000, y.raw=-3000000)
#all(which(x.raw1==-3000000)==which(y.raw1==-3000000))
sel <- which(x.raw1==-3000000)
x.raw1[sel] <- NA
y.raw1[sel] <- NA
z.raw1[sel] <- NA

x.raw2 <- data.frame(x.raw1)
y.raw2 <- data.frame(y.raw1)
z.raw2 <- data.frame(z.raw1)

## remove particles that only have NAs
sel.x <- which(sapply(x.raw2, function(x)all(is.na(x))))
sel.y <- which(sapply(y.raw2, function(x)all(is.na(x))))
#all(sel.y==sel.x)
x.raw <- x.raw2[,-sel.x]
y.raw <- y.raw2[,-sel.x]
z.raw <- z.raw2[,-sel.x]

## calculate ocean depth at each particle location (this takes about 15min)
z.h <- data.frame(matrix(NA, nrow=nrow(z.raw), ncol=ncol(z.raw)))
for(i in 1:ncol(x.raw)){
  z.h[,i] <- extract(h2, cbind(x.raw[,i],y.raw[,i]))
}

## calculate distance to seafloor at each particle location
z.diff <- z.h+z.raw

dim(x.raw)
dim(y.raw)
dim(z.raw)
dim(z.diff)
dim(z.h)

## save output
save(x.raw,y.raw,z.raw,z.h,z.diff, file=paste0(ARC_data_dir,"FAM/Floats_10km_100mday_test.Rdata"))
```

```{r, eval=FALSE}
#######################
## CREATE LINES FOR 6-MONTHS OF TRAVELLING TIME

load(paste0(ARC_data_dir,"FAM/Floats_10km_100mday_test.Rdata"))
##########################

## subset first
# set.seed(1)
# s <- sample(1:ncol(x.raw), 10000)
s <- 1:ncol(x.raw)

## split points into (1) in the water column and (2) on the bottom

## prepare data objects and translate particle coordinates into lines
line_list_s <- vector(mode="list", length=length(s))
line_list_b <- vector(mode="list", length=length(s))
xyz_list <- vector(mode="list", length=length(s))
xyz_spatial <- vector(mode="list", length=length(s))
names(line_list_s) <- names(line_list_b) <- s

for(i in 1:length(s)){
  # message(i)
  ## coordinates etc for each particle
  xyz.raw <- data.frame(cbind(x.raw[,s[i]],y.raw[,s[i]],z.diff[,s[i]],z.raw[,s[i]],z.h[,s[i]],1:152,NA))
  names(xyz.raw) <- c("x","y","z_diff","z_particle","z_cell","time","bottom")
  ## identify any NA locations for that particle
  na.sel <- which(is.na(rowSums(xyz.raw[,1:2])))
  ## remove single NA locations
  if(length(na.sel)>0) {
    xyz <- xyz.raw[-na.sel,]
  }else
    xyz <- xyz.raw
  ## find first instance of where the particle is within 100m of the seafloor
  sel.first <- tail(which(xyz[,3]<100),1)
  ## select between surface and seafloor
  if(length(sel.first)>0){
    sel.surface <- (sel.first+1):nrow(xyz)
    sel.bottom <- 1:sel.first
    if(sel.first==nrow(xyz)){
      #message(paste0("Particles close to seafloor here: ",i))
      line_list_b[[i]] <- Lines(list(Line(xyz[,1:2])), ID=s[i])
    } else{
      xyz$bottom[sel.surface] <- "no"
      xyz$bottom[sel.bottom] <- "yes"
      line_list_s[[i]] <- Lines(list(Line(xyz[which(xyz$bottom=="no"),1:2])), ID=s[i])
      line_list_b[[i]] <- Lines(list(Line(xyz[which(xyz$bottom=="yes"),1:2])), ID=s[i])
    }   
  }else {
    xyz$bottom <- "no"
    line_list_s[[i]] <- Lines(list(Line(xyz[which(xyz$bottom=="no"),1:2])), ID=s[i])
  }
  ## create line from the locations of each particle
  # if(length(sel.not.bottom)==152){
  #   line_list_s[[i]] <- Lines(list(Line(xyz[,1:2])), ID=s[i])
  # }else{
  xyz_list[[i]] <- xyz
  xyz_spatial[[i]] <- SpatialPointsDataFrame(xyz[,1:2],xyz[,3:7])
  # }
}

# ## remove particles which only have NA locations
line_list_s2 <- line_list_s[-which(sapply(line_list_s, is.null))]
line_list_b2 <- line_list_b[-which(sapply(line_list_b, is.null))]

## make SpatialLines from Lines
track_lines_s <- SpatialLines(line_list_s2)
track_lines_b <- SpatialLines(line_list_b2)

save(track_lines_s, track_lines_b, xyz_spatial, xyz_list, file=paste0(ARC_data_dir,"FAM/Floats_10km_100mday_test_lines.Rdata"))
```


```{r, eval=FALSE}
#######################
## CREATE LINES FOR 1-MONTH OF TRAVELLING TIME
load(paste0(ARC_data_dir,"FAM/Floats_10km_100mday_test.Rdata"))

m <- 122:152
##########################

## subset first
# set.seed(1)
# s <- sample(1:ncol(x.raw), 1000)
s <- 1:ncol(x.raw)

## split points into (1) in the water column and (2) on the bottom

## prepare data objects and translate particle coordinates into lines
line_list_s <- vector(mode="list", length=length(s))
line_list_b <- vector(mode="list", length=length(s))
xyz_list <- vector(mode="list", length=length(s))
xyz_spatial <- vector(mode="list", length=length(s))
names(line_list_s) <- names(line_list_b) <- s

for(i in 1:length(s)){
  # message(i)
  ## coordinates etc for each particle
  xyz.raw <- data.frame(cbind(x.raw[m,s[i]],y.raw[m,s[i]],z.diff[m,s[i]],z.raw[m,s[i]],z.h[m,s[i]],m,NA))
  names(xyz.raw) <- c("x","y","z_diff","z_particle","z_cell","time","bottom")
  ## identify any NA locations for that particle
  na.sel <- which(is.na(rowSums(xyz.raw[,1:2])))
  ## remove single NA locations
  if(length(na.sel)>0) {
    xyz <- xyz.raw[-na.sel,]
  }else
    xyz <- xyz.raw
  ## find first instance of where the particle is within 100m of the seafloor
  sel.first <- tail(which(xyz[,3]<100),1)
  ## select between surface and seafloor
  if(length(sel.first)>0){
    sel.surface <- (sel.first+1):nrow(xyz)
    sel.bottom <- 1:sel.first
    if(sel.first==nrow(xyz)){
      #message(paste0("Particles close to seafloor here: ",i))
      line_list_b[[i]] <- Lines(list(Line(xyz[,1:2])), ID=s[i])
    } else{
      xyz$bottom[sel.surface] <- "no"
      xyz$bottom[sel.bottom] <- "yes"
      line_list_s[[i]] <- Lines(list(Line(xyz[which(xyz$bottom=="no"),1:2])), ID=s[i])
      line_list_b[[i]] <- Lines(list(Line(xyz[which(xyz$bottom=="yes"),1:2])), ID=s[i])
    }   
  }else {
    xyz$bottom <- "no"
    line_list_s[[i]] <- Lines(list(Line(xyz[which(xyz$bottom=="no"),1:2])), ID=s[i])
  }
  ## create line from the locations of each particle
  # if(length(sel.not.bottom)==152){
  #   line_list_s[[i]] <- Lines(list(Line(xyz[,1:2])), ID=s[i])
  # }else{
  xyz_list[[i]] <- xyz
  xyz_spatial[[i]] <- SpatialPointsDataFrame(xyz[,1:2],xyz[,3:7])
  # }
}

# ## remove particles which only have NA locations
line_list_s2 <- line_list_s[-which(sapply(line_list_s, is.null))]
line_list_b2 <- line_list_b[-which(sapply(line_list_b, is.null))]

## make SpatialLines from Lines
track_lines_s <- SpatialLines(line_list_s2)
track_lines_b <- SpatialLines(line_list_b2)

save(track_lines_s, track_lines_b, xyz_spatial, xyz_list, file=paste0(ARC_data_dir,"FAM/Floats_10km_100mday_test_lines_1month.Rdata"))
```

```{r}
load(paste0(ARC_data_dir,"FAM/Floats_10km_100mday_test_lines.Rdata"))
#load(paste0(ARC_data_dir,"FAM/Floats_10km_100mday_test_lines.Rdata_1month"))
```


# PAPER OUTPUTS:

1. seafloor connectivity map
- which areas are connected to each other
- maybe a contour plot?

1. under-ice particles


## full particle trajectories

### Mertz Area

```{r, fig.height=7.5, fig.width=10}
xlim <- c(1300000,1900000)
ylim <- c(-2300000,-1900000)
## plot sinking only:
plot(uv_31, xlim=xlim, ylim=ylim, main="sinking movements", breaks=breaks, col=cols)
# plot(track_lines_s, add=TRUE)
plot(track_lines_s[seq(1,length(track_lines_s), by=2)], add=TRUE)
scalebar(100000, xy=c(1350000,-1930000), type="bar", label=c(0,50,100), below="km")
```

```{r, fig.height=7.5, fig.width=10}
## plot bottom only:
plot(mean.uv, xlim=xlim, ylim=ylim, main="seafloor movements", breaks=breaks, col=cols)
# plot(track_lines_b, add=TRUE)
plot(track_lines_b[seq(1,length(track_lines_b), by=2)], add=TRUE)
scalebar(100000, xy=c(1350000,-1930000), type="bar", label=c(0,50,100), below="km")
```


### Ross Sea

```{r, fig.height=8, fig.width=9}
xlim=c(-500000,500000)
ylim=c(-2100000,-1150000)
## plot sinking only:
plot(uv_31, xlim=xlim, ylim=ylim, main="sinking movements", breaks=breaks, col=cols)
# plot(track_lines_s, add=TRUE)
plot(track_lines_s[seq(1,length(track_lines_s), by=2)], add=TRUE)
scalebar(100000, xy=c(0,-1200000), type="bar", label=c(0,50,100), below="km")
```

```{r, fig.height=8, fig.width=9}
## plot bottom only:
plot(mean.uv, xlim=xlim, ylim=ylim, main="seafloor movements", breaks=breaks, col=cols)
# plot(track_lines_b, add=TRUE)
plot(track_lines_b, add=TRUE)
scalebar(100000, xy=c(0,-1200000), type="bar", label=c(0,50,100), below="km")
```

```{r, fig.height=8, fig.width=9}
## plot bottom only:
plot(mean.uv, xlim=xlim, ylim=ylim, main="seafloor movements", breaks=breaks, col=cols)
# plot(track_lines_b, add=TRUE)
plot(track_lines_b[seq(1,length(track_lines_b), by=2)], add=TRUE)
scalebar(100000, xy=c(0,-1200000), type="bar", label=c(0,50,100), below="km")
```

```{r, fig.height=8, fig.width=9}
## plot bottom only:
plot(mean.uv, xlim=xlim, ylim=ylim, main="seafloor movements", breaks=breaks, col=cols)
# plot(track_lines_b, add=TRUE)
plot(track_lines_b[seq(1,length(track_lines_b), by=4)], add=TRUE)
scalebar(100000, xy=c(0,-1200000), type="bar", label=c(0,50,100), below="km")
```

### Weddell Sea

```{r, fig.height=9, fig.width=10}
xlim=c(-2000000,-500000)
ylim=c(200000,1500000)
## plot bottom only:
plot(uv_31, xlim=xlim, ylim=ylim, main="sinking movements", breaks=breaks, col=cols)
# plot(track_lines_s, add=TRUE)
plot(track_lines_s[seq(1,length(track_lines_s), by=2)], add=TRUE)
scalebar(100000, xy=c(-800000,300000), type="bar", label=c(0,50,100), below="km")
```

```{r, fig.height=9, fig.width=10}
## plot bottom only:
plot(mean.uv, xlim=xlim, ylim=ylim, main="seafloor movements", breaks=breaks, col=cols)
# plot(track_lines_b, add=TRUE)
plot(track_lines_b[seq(1,length(track_lines_b), by=2)], add=TRUE)
scalebar(100000, xy=c(-800000,300000), type="bar", label=c(0,50,100), below="km")
```











```{r, eval=FALSE, echo=FALSE}
## x and y locations are stored in "grd.x" and "grd.y" as time-steps (rows) and individual particles (columns)
x.raw <- matrix(NA, nrow=152, ncol=ncol(grd.x100))
y.raw <- x.raw
for(i in 1:152){
  x.raw[i,] <- grd.x100[i,,]
  y.raw[i,] <- grd.y100[i,,]
}
## remove points that move into the bottom corner!!! (x.raw=-3000000, y.raw=-3000000)
x.raw[which(x.raw==-3000000)] <- NA
y.raw[which(y.raw==-3000000)] <- NA

## remove NAs
# all(which(is.na(colSums(x)))==which(is.na(colSums(y))))
x <- x.raw[,-which(is.na(colSums(x.raw)))]
y <- y.raw[,-which(is.na(colSums(x.raw)))]

#### particle trajectories (only use a subset of 100 traces)
set.seed(1)
s <- sample(1:ncol(x), 100)
line_list_s <- list()
for(i in 1:length(s)){
  line_list_s[[i]] <- Lines(list(Line(cbind(x[,s[i]],y[,s[i]]))), ID=s[i])
}
names(line_list_s) <- s
track_lines_s <- SpatialLines(line_list_s)
plot(track_lines_s)

#### particle trajectories (all traces)
line_list <- list()
for(i in 1:ncol(x)){
  line_list[[i]] <- Lines(list(Line(cbind(x[,i],y[,i]))), ID=i)
}
names(line_list) <- 1:ncol(x)
track_lines <- SpatialLines(line_list)


## density of lines per raster-cell


## first remove points at the sea-surface:



###############################
# ## plot floating traces
# ## not near the bottom
# plot(track_lines_s, xlim=xlim, ylim=ylim)
# ## near the bottom
# plot(track_lines_b, col="green", add=TRUE)
# ## add individual particle locations
# for(i in 1:length(s)){
#   ## all individual particle locations
#   points(xyz_spatial[[i]], col="grey")
#   ## particle locations of bottom particles only
#   points(xyz_spatial[[i]][xyz_spatial[[i]]$bottom=="yes",], col="brown")
# }
# #points(cbind(unlist(c(x.raw[,s])),unlist(c(y.raw[,s])))) ## all particles

# #################
# ## this works below:
# 
# ## prepare data objects and translate particle coordinates into lines
# line_list_n <- vector(mode="list", length=length(s))
# names(line_list_n) <- s
# for(i in 1:length(s)){
#   print(i)
#   ## coordinates and depth
#   xyz.raw <- cbind(x.raw[,s[i]],y.raw[,s[i]],z.raw[,s[i]]) 
#   ## identify particles with NA locations
#   na.sel <- which(is.na(rowSums(xyz.raw[,1:2])))
#   ## remove single NA locations
#   if(length(na.sel)>0) {
#     xyz <- xyz.raw[-na.sel,] 
#   }else xyz <- xyz.raw
#   ## identify if the particle is near the seafloor or not
#   
#   
#   ## set particles to NULL which only have NA locations
#   if(nrow(xyz)==0){
#       line_list_n[[i]] <- NULL 
#   }else
#     ## create line from the locations of each particle
#     line_list_n[[i]] <- Lines(list(Line(xyz[,1:2])), ID=s[i])
# }
# ## remove particles which only have NA locations
# line_list_n2 <- line_list_s[-which(sapply(line_list_s, is.null))]
# ## make SpatialLines from Lines
# track_lines_n <- SpatialLines(line_list_n2)
# 
# ## plot
# plot(track_lines_n)
```





