---
title: "ROMS FAM 4km analysis"
author: "Jan Jansen"
date: "20/01/2022"
output: html_document
---

###########################################
########## Evaluate ROMS outputs ##########
###########################################

```{r setup, include=FALSE}
#### setup environment
library(raster)
library(colorspace)
library(viridis)
library(ncdf4)
library(raadtools)
library(rgdal)
library(scico)
## data-directories
env.dir <- "E:/science/data_environmental/"
env.raw <- paste0(env.dir,"raw/")
env.derived <- "C:/Users/jjansen/Desktop/science/data_environmental/derived/"
AAD_dir <- paste0(env.raw,"accessed_through_R")
ARC_data_dir <- "C:/Users/jjansen/Desktop/science/SouthernOceanBiodiversityMapping/ARC_Data/"
ant.proj <- "+proj=stere +lat_0=-90 +lat_ts=-71 +lon_0=0 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs"
```

########## Prepare data ##########

```{r}
#### load non-ROMS data
## load projected coastline for plotting
load(paste0(env.derived,"Circumpolar_Coastline.Rdata"))
r2 <- raster(paste0(env.derived,"Circumpolar_EnvData_500m_shelf_bathy_gebco_depth.grd"))
## load circumpolar Diatom data
load(paste0(ARC_data_dir,"FAM/Circumpolar_Diatom_metadata.Rdata"))
## load Mertz Diatom data
mertz.diatom.dat <- read.csv("C:/Users/jjansen/Desktop/science/PhD/1 bentho-pelagic-coupling/DiatomAbundance.csv")
mertz.diatom.locs <- project(as.matrix(mertz.diatom.dat[,c(4,3)]),proj=ant.proj)

## load TOC data
sed.dir <- "C:/Users/jjansen/Desktop/science/data_biological/sediment_data/"
mertz.toc.dat <- read.csv(paste0(sed.dir,"Circumpolar_TOC_data_CEAMARC_sediment_isotopes.csv"))
mertz.toc.locs <- project(as.matrix(mertz.toc.dat[,2:3]),proj=ant.proj)
ross.toc.dat <- read.csv(paste0(sed.dir,"Circumpolar_TOC_data_villinski_2008.csv"))
ross.toc.locs <- project(as.matrix(ross.toc.dat[,c(3,2)]),proj=ant.proj)
ant.toc.dat <- read.csv(paste0(sed.dir,"Circumpolar_TOC_data_Leventer.csv"), skip=4)[-c(90,95),] ## rows 90 and 95 have no coordinated
ant.toc.locs.raw <- read.csv(paste0(sed.dir,"Circumpolar_TOC_coords_Leventer.csv"))
ant.toc.locs <- project(as.matrix(ant.toc.locs.raw[-c(90,95),c(5,4)]),proj=ant.proj)

#### load ROMS-runs:
data.dat100 <- paste0(env.dir,"Circumpolar_ROMS/4km_outputs/output_sed_test2a/")
# data.dat100 <- paste0(env.dir,"Circumpolar_ROMS/4km_outputs/output_sed_float_test_6/")
# data.dat100 <- paste0(env.dir,"Circumpolar_ROMS/4km_outputs/output_sed_float_test8/")
# data.dat100 <- paste0(env.dir,"Circumpolar_ROMS/10km_outputs/output_sed_float_test3/")
# data.dat200 <- paste0(env.dir,"Circumpolar_ROMS/10km_outputs/output_sed_float_test4/")
# data.dat400 <- paste0(env.dir,"Circumpolar_ROMS/10km_outputs/output_sed_float_test5/")
# data.dat <- paste0("Circumpolar_ROMS/10km_outputs/sed_test1/")
# data.dat <- paste0(env.dir,"Circumpolar_ROMS/10km_outputs/sed_test5/")

#### load lon/lat information from ROMS-grid
grd4k_nc <- nc_open(paste0(env.raw,"waom4extend_grd.nc"))
lon_rho <- ncvar_get(grd4k_nc, varid="lon_rho")
lat_rho <- ncvar_get(grd4k_nc, varid="lat_rho")

#### Prepare empty rasters to assign correct projected values to
roms.coords.proj <- project(cbind(c(lon_rho), c(lat_rho)), proj=ant.proj)
x.range <- c(min(roms.coords.proj[,1])-2000,max(roms.coords.proj[,1])+2000)
y.range <- c(min(roms.coords.proj[,2])-2000,max(roms.coords.proj[,2])+2000)
empty.roms.ra <- raster(extent(c(x.range,y.range)), crs=ant.proj, resolution=4000)

#### load other ROMS data (currents & depth)
#depth
h <- raster(paste0(data.dat100,"ocean_avg_0001.nc"), varname="h", level=1)
#salinity
salt <- raster(paste0(data.dat100,"ocean_avg_0001.nc"), varname="salt", level=1)
#temperature
temp <- raster(paste0(data.dat100,"ocean_avg_0001.nc"), varname="temp", level=1)
#seafloor currents (seafloor-layer is 1)
u.raw <- brick(paste0(data.dat100,"ocean_his_0001.nc"), varname="u", level=1)
v.raw <- brick(paste0(data.dat100,"ocean_his_0001.nc"), varname="v", level=1)
#seasurface currents (surface-layer is 31)
u_31.raw <- brick(paste0(data.dat100,"ocean_avg_0001.nc"), varname="u", level=31)
v_31.raw <- brick(paste0(data.dat100,"ocean_avg_0001.nc"), varname="v", level=31)
## sum up monthly values for a climatology (THIS SHOULD BE DONE ON HIGH RESOLUTION HISTORY FILES)
u.mean <- sum(u.raw)/nlayers(u.raw)
v.mean <- sum(v.raw)/nlayers(v.raw)
u.mean.abs <- sum(abs(u.raw))/nlayers(u.raw)
v.mean.abs <- sum(abs(v.raw))/nlayers(v.raw)
u31.mean <- sum(u_31.raw)/nlayers(u_31.raw)
v31.mean <- sum(v_31.raw)/nlayers(v_31.raw)
## extract current speeds at rho-points where depth is defined
## (in ROMS they are all at different locations):
## u has one less column than the grid, so all coordinates need to be moved half a cell to the right for the grids to match up
coord.grd.u <- coordinates(h)
coord.grd.u[,1] <- coordinates(h)[,1]-0.5
## v has one less row than the grid, so all coordinates need to be moved half a cell up for the grids to match up
coord.grd.v <- coordinates(h)
coord.grd.v[,2] <- coordinates(h)[,2]-0.5
## now extract values at the rho-points and interpolate (because they are 2km away from the nearest original point), and place into projected raster
u <- v <- u.abs <- v.abs <- u31 <- v31 <- sa <- te <- empty.roms.ra
u[] <- extract(u.mean, coord.grd.u, method="bilinear")
v[] <- extract(v.mean, coord.grd.v, method="bilinear")
u.abs[] <- extract(u.mean.abs, coord.grd.u, method="bilinear")
v.abs[] <- extract(v.mean.abs, coord.grd.v, method="bilinear")
u31[] <- extract(u31.mean, coord.grd.u, method="bilinear")
v31[] <- extract(v31.mean, coord.grd.v, method="bilinear")
#seasurface current speeds
uv_31 <- sqrt(u31^2+v31^2)
#temporal mean seafloor current speed
mean.uv <- sqrt(u^2+v^2)
#absolute mean seafloor current speed
abs.uv <- sqrt(u.abs^2+v.abs^2)
#residual seafloor current speed
res.uv <- abs.uv-mean.uv
## remove inland values for depth
h2 <- empty.roms.ra
h2[] <- h[]
h2[is.na(mean.uv)] <- NA
## salt & temperature
sa[] <- extract(salt, coordinates(h), method="bilinear")
te[] <- extract(temp, coordinates(h), method="bilinear")

# #temporal mean seafloor current speed
# mean.u <- sum(u)
# mean.v <- sum(v)
# #absolute mean seafloor current speed
# abs.u <- sum(abs(u))
# abs.v <- sum(abs(v))
# #residual seafloor current speed
# res.u <- abs.u-mean.u
# res.v <- abs.v-mean.v
# mean.uv2 <- sqrt(mean.u^2+mean.v^2)
# abs.uv2 <- sqrt(abs.u^2+abs.v^2)
# res.uv2 <- abs.uv2-mean.uv2

```

########## Overview plots - Ocean currents ##########

```{r, fig.width=10, fig.height=10}
## ocean current data
#cols <- rev(terrain.colors(96))
cols <- rev(magma(81)) #12, 24, 48, 96
breaks <- seq(0,0.81,length.out=82)
par(mfrow=c(2,2))
plot(mean.uv, breaks=breaks, col=cols, main="seafloor currents - mean")
plot(abs.uv, breaks=breaks, col=cols, main="seafloor currents - absolute speed")
plot(res.uv, breaks=breaks, col=cols, main="seafloor currents - residual")
plot(uv_31, breaks=breaks, col=cols, main="seasurface currents")
## ocean current data in the Mertz region
cols <- rev(magma(50)) #12, 24, 48, 96
breaks <- seq(0,0.5,length.out=51)
xlim <- c(1300000,1900000)
ylim <- c(-2300000,-1900000)
par(mfrow=c(2,2))
plot(mean.uv, breaks=breaks, col=cols, main="seafloor currents - mean", xlim=xlim, ylim=ylim)
#plot(coast.proj, add=TRUE)
plot(abs.uv, breaks=breaks, col=cols, main="seafloor currents - absolute speed", xlim=xlim, ylim=ylim)
#plot(coast.proj, add=TRUE)
plot(res.uv, breaks=breaks, col=cols, main="seafloor currents - residual", xlim=xlim, ylim=ylim)
#plot(coast.proj, add=TRUE)
plot(h2, col=cols, main="depth", xlim=xlim, ylim=ylim)
#plot(coast.proj, add=TRUE)
```

########## EVALUATE FAM ##########

#### Check input surface layer:
```{r}
test <- brick("C:/Users/jjansen/Desktop/science/data_environmental/derived/waom4extend_nudge_climatology20022019_shelf.nc")

input_01 <- brick(empty.roms.ra,nl=7)

```




```{r, message=FALSE, warning=FALSE, results="hide"}
## values originally in kg/m2/s -> change to kg/m2/day (*86400)
### NOTE THAT CURRENTLY DIFFERENT MODEL RUNS HAVE ALL THE SAME OUTPUT APART FROM THE "ocean_flt.nc" file
surf_01 <- surf_02 <- surf_03 <- surf_04 <- surf_05 <- surf_06 <- surf_07 <- surf_08 <- brick(empty.roms.ra,nl=7)
susp_01 <- susp_02 <- susp_03 <- susp_04 <- susp_05 <- susp_06 <- susp_07 <- susp_08 <- brick(empty.roms.ra,nl=7)
surf_his_01 <- surf_his_02 <- surf_his_03 <- surf_his_04 <- surf_his_05 <- surf_his_06 <- surf_his_07 <- surf_his_08 <-  brick(empty.roms.ra,nl=nlayers(u.raw))
susp_his_01 <- susp_his_02 <- susp_his_03 <- susp_his_04 <- susp_his_05 <- susp_his_06 <- susp_his_07 <- susp_his_08 <-  brick(empty.roms.ra,nl=nlayers(u.raw))
settle_01 <- settle_02 <- settle_03 <- settle_04 <- settle_05 <- settle_06 <- settle_07 <- settle_08 <- brick(empty.roms.ra,nl=nlayers(u.raw))
## surface production
surf_01[] <- brick(paste0(data.dat100,"ocean_avg_0001.nc"), varname="sand_01", level=31)[]*86400
surf_02[] <- brick(paste0(data.dat100,"ocean_avg_0001.nc"), varname="sand_02", level=31)[]*86400
surf_03[] <- brick(paste0(data.dat100,"ocean_avg_0001.nc"), varname="sand_03", level=31)[]*86400
surf_04[] <- brick(paste0(data.dat100,"ocean_avg_0001.nc"), varname="sand_04", level=31)[]*86400
surf_05[] <- brick(paste0(data.dat100,"ocean_avg_0001.nc"), varname="sand_05", level=31)[]*86400
surf_06[] <- brick(paste0(data.dat100,"ocean_avg_0001.nc"), varname="sand_06", level=31)[]*86400
surf_07[] <- brick(paste0(data.dat100,"ocean_avg_0001.nc"), varname="sand_07", level=31)[]*86400
surf_08[] <- brick(paste0(data.dat100,"ocean_avg_0001.nc"), varname="sand_08", level=31)[]*86400
# surf_his_01[] <- brick(paste0(data.dat100,"ocean_his_0001.nc"), varname="sand_01", level=31)[]*86400
# surf_his_02[] <- brick(paste0(data.dat100,"ocean_his_0001.nc"), varname="sand_02", level=31)[]*86400
# surf_his_03[] <- brick(paste0(data.dat100,"ocean_his_0001.nc"), varname="sand_03", level=31)[]*86400
surf_his_04[] <- brick(paste0(data.dat100,"ocean_his_0001.nc"), varname="sand_04", level=31)[]*86400
# surf_his_05[] <- brick(paste0(data.dat100,"ocean_his_0001.nc"), varname="sand_05", level=31)[]*86400
# surf_his_06[] <- brick(paste0(data.dat100,"ocean_his_0001.nc"), varname="sand_06", level=31)[]*86400
# surf_his_07[] <- brick(paste0(data.dat100,"ocean_his_0001.nc"), varname="sand_07", level=31)[]*86400
surf_his_08[] <- brick(paste0(data.dat100,"ocean_his_0001.nc"), varname="sand_08", level=31)[]*86400
## bottom ocean layer
susp_01[] <- brick(paste0(data.dat100,"ocean_avg_0001.nc"), varname="sand_01", level=1)[]*86400
susp_02[] <- brick(paste0(data.dat100,"ocean_avg_0001.nc"), varname="sand_02", level=1)[]*86400
susp_03[] <- brick(paste0(data.dat100,"ocean_avg_0001.nc"), varname="sand_03", level=1)[]*86400
susp_04[] <- brick(paste0(data.dat100,"ocean_avg_0001.nc"), varname="sand_04", level=1)[]*86400
susp_05[] <- brick(paste0(data.dat100,"ocean_avg_0001.nc"), varname="sand_05", level=1)[]*86400
susp_06[] <- brick(paste0(data.dat100,"ocean_avg_0001.nc"), varname="sand_06", level=1)[]*86400
susp_07[] <- brick(paste0(data.dat100,"ocean_avg_0001.nc"), varname="sand_07", level=1)[]*86400
susp_08[] <- brick(paste0(data.dat100,"ocean_avg_0001.nc"), varname="sand_08", level=1)[]*86400
# susp_his_01[] <- brick(paste0(data.dat100,"ocean_his_0001.nc"), varname="sand_01", level=1)[]*86400
# susp_his_02[] <- brick(paste0(data.dat100,"ocean_his_0001.nc"), varname="sand_02", level=1)[]*86400
# susp_his_03[] <- brick(paste0(data.dat100,"ocean_his_0001.nc"), varname="sand_03", level=1)[]*86400
susp_his_04[] <- brick(paste0(data.dat100,"ocean_his_0001.nc"), varname="sand_04", level=1)[]*86400
# susp_his_05[] <- brick(paste0(data.dat100,"ocean_his_0001.nc"), varname="sand_05", level=1)[]*86400
# susp_his_06[] <- brick(paste0(data.dat100,"ocean_his_0001.nc"), varname="sand_06", level=1)[]*86400
# susp_his_07[] <- brick(paste0(data.dat100,"ocean_his_0001.nc"), varname="sand_07", level=1)[]*86400
susp_his_08[] <- brick(paste0(data.dat100,"ocean_his_0001.nc"), varname="sand_08", level=1)[]*86400
## settled particles
settle_01[] <- brick(paste0(data.dat100,"ocean_his_0001.nc"), varname="sandfrac_01", level=1)[]*86400
settle_02[] <- brick(paste0(data.dat100,"ocean_his_0001.nc"), varname="sandfrac_02", level=1)[]*86400
settle_03[] <- brick(paste0(data.dat100,"ocean_his_0001.nc"), varname="sandfrac_03", level=1)[]*86400
settle_04[] <- brick(paste0(data.dat100,"ocean_his_0001.nc"), varname="sandfrac_04", level=1)[]*86400
settle_05[] <- brick(paste0(data.dat100,"ocean_his_0001.nc"), varname="sandfrac_05", level=1)[]*86400
settle_06[] <- brick(paste0(data.dat100,"ocean_his_0001.nc"), varname="sandfrac_06", level=1)[]*86400
settle_07[] <- brick(paste0(data.dat100,"ocean_his_0001.nc"), varname="sandfrac_07", level=1)[]*86400
settle_08[] <- brick(paste0(data.dat100,"ocean_his_0001.nc"), varname="sandfrac_08", level=1)[]*86400
## bedload u
#bedu_01 <- brick(paste0(data.dat100,"ocean_avg_0001.nc"), varname="bedload_Usand_01", level=31)

```

########## Overview plots - FAM ##########

# why the weird values in the sandfrac output?
i.e. not comparable to surface or suspended fraction

# is there a way to calculate flux along the seafloor?

```{r, fig.width=10, fig.height=10}
## surface productivity 
# plot(surf_08*86400, breaks=seq(0,0.24,length.out=25), col=rev(heat.colors(25)))
# #plot(susp_08*86400, breaks=seq(0,0.24,length.out=25), col=rev(heat.colors(25)))
# plot(susp_08*86400, breaks=round(seq(-0.1,0.24,length.out=35),3), col=c(rev(blues9),rev(heat.colors(25))))
# plot(settle_08*86400, breaks=c(10780,seq(10795,10810,length.out=25),10935), col=c("green",rev(heat.colors(24)),"black"))

#### comparisons
breaks1 <- seq(0,0.25,length.out=51)
breaks2 <- c(-0.5,round(seq(0,0.25,length.out=51),3))
#breaks3 <- c(10780,seq(10795,10810,length.out=25),10935)
breaks3 <- c(10795,seq(10800,10805,length.out=51))
cols1 <- rev(viridis(50))#rev(magma(48))
cols2 <- c("grey",rev(viridis(50)))
cols3 <- c("grey",rev(viridis(50)))
par(mfrow=c(3,3), mar=c(3,4,2,1))
plot(surf_08[[1]], breaks=breaks1, col=cols1, main=paste0("surf_08 ",names(surf_08)[1]))
plot(surf_08[[4]], breaks=breaks1, col=cols1, main=paste0("surf_08 ",names(surf_08)[4]))
plot(surf_08[[7]], breaks=breaks1, col=cols1, main=paste0("surf_08 ",names(surf_08)[7]))
plot(susp_08[[1]], breaks=breaks2, col=cols2, main=paste0("susp_08 ",names(susp_08)[1]))
plot(susp_08[[4]], breaks=breaks2, col=cols2, main=paste0("susp_08 ",names(susp_08)[4]))
plot(susp_08[[7]], breaks=breaks2, col=cols2, main=paste0("susp_08 ",names(susp_08)[7]))
# plot(settle_08[[10]], breaks=breaks3, col=cols3, main=paste0("settle_08 ",names(settle_08)[2]))
# plot(settle_08[[20]], breaks=breaks3, col=cols3, main=paste0("settle_08 ",names(settle_08)[3]))
# plot(settle_08[[30]], breaks=breaks3, col=cols3, main=paste0("settle_08 ",names(settle_08)[4]))
plot(settle_08[[2]], breaks=breaks3, col=cols3, main=paste0("settle_08 ",names(settle_08)[2]))
plot(settle_08[[5]], breaks=breaks3, col=cols3, main=paste0("settle_08 ",names(settle_08)[5]))
plot(settle_08[[8]], breaks=breaks3, col=cols3, main=paste0("settle_08 ",names(settle_08)[8]))

#### comparisons Mertz
par(mfrow=c(3,3))
plot(surf_08[[1]], xlim=xlim, ylim=ylim, breaks=breaks1, col=cols1, main=paste0("surf_08 ",names(surf_08)[1]))
plot(surf_08[[4]], xlim=xlim, ylim=ylim, breaks=breaks1, col=cols1, main=paste0("surf_08 ",names(surf_08)[4]))
plot(surf_08[[7]], xlim=xlim, ylim=ylim, breaks=breaks1, col=cols1, main=paste0("surf_08 ",names(surf_08)[7]))
plot(susp_08[[1]], xlim=xlim, ylim=ylim, breaks=breaks2, col=cols2, main=paste0("susp_08 ",names(susp_08)[1]))
plot(susp_08[[4]], xlim=xlim, ylim=ylim, breaks=breaks2, col=cols2, main=paste0("susp_08 ",names(susp_08)[4]))
plot(susp_08[[7]], xlim=xlim, ylim=ylim, breaks=breaks2, col=cols2, main=paste0("susp_08 ",names(susp_08)[7]))
# plot(settle_08[[10]], xlim=xlim, ylim=ylim, breaks=breaks3, col=cols3, main=paste0("settle_08 ",names(settle_08)[10]))
# plot(settle_08[[20]], xlim=xlim, ylim=ylim, breaks=breaks3, col=cols3, main=paste0("settle_08 ",names(settle_08)[20]))
# plot(settle_08[[30]], xlim=xlim, ylim=ylim, breaks=breaks3, col=cols3, main=paste0("settle_08 ",names(settle_08)[30]))
plot(settle_08[[2]], xlim=xlim, ylim=ylim, breaks=breaks3, col=cols3, main=paste0("settle_08 ",names(settle_08)[2]))
plot(settle_08[[5]], xlim=xlim, ylim=ylim, breaks=breaks3, col=cols3, main=paste0("settle_08 ",names(settle_08)[5]))
plot(settle_08[[8]], xlim=xlim, ylim=ylim, breaks=breaks3, col=cols3, main=paste0("settle_08 ",names(settle_08)[8]))

#### comparisons surf / sink / settle for Mertz
par(mfrow=c(3,3))
# plot(settle_04[[10]], xlim=xlim, ylim=ylim, breaks=breaks3, col=cols1, main=paste0("settle_04 ",names(settle_04)[10]))
# plot(settle_04[[20]], xlim=xlim, ylim=ylim, breaks=breaks3, col=cols1, main=paste0("settle_04 ",names(settle_04)[20]))
# plot(settle_04[[30]], xlim=xlim, ylim=ylim, breaks=breaks3, col=cols1, main=paste0("settle_04 ",names(settle_04)[30]))
plot(settle_04[[2]], xlim=xlim, ylim=ylim, breaks=breaks3, col=cols1, main=paste0("settle_04 ",names(settle_04)[2]))
plot(settle_04[[5]], xlim=xlim, ylim=ylim, breaks=breaks3, col=cols1, main=paste0("settle_04 ",names(settle_04)[5]))
plot(settle_04[[8]], xlim=xlim, ylim=ylim, breaks=breaks3, col=cols1, main=paste0("settle_04 ",names(settle_04)[8]))
# plot(settle_06[[10]], xlim=xlim, ylim=ylim, breaks=breaks3, col=cols1, main=paste0("settle_06 ",names(settle_06)[10]))
# plot(settle_06[[20]], xlim=xlim, ylim=ylim, breaks=breaks3, col=cols1, main=paste0("settle_06 ",names(settle_06)[20]))
# plot(settle_06[[30]], xlim=xlim, ylim=ylim, breaks=breaks3, col=cols1, main=paste0("settle_06 ",names(settle_06)[30]))
# plot(settle_08[[10]], xlim=xlim, ylim=ylim, breaks=breaks3, col=cols1, main=paste0("settle_08 ",names(settle_08)[10]))
# plot(settle_08[[20]], xlim=xlim, ylim=ylim, breaks=breaks3, col=cols1, main=paste0("settle_08 ",names(settle_08)[20]))
# plot(settle_08[[30]], xlim=xlim, ylim=ylim, breaks=breaks3, col=cols1, main=paste0("settle_08 ",names(settle_08)[30]))
plot(settle_08[[2]], xlim=xlim, ylim=ylim, breaks=breaks3, col=cols1, main=paste0("settle_08 ",names(settle_08)[2]))
plot(settle_08[[5]], xlim=xlim, ylim=ylim, breaks=breaks3, col=cols1, main=paste0("settle_08 ",names(settle_08)[5]))
plot(settle_08[[8]], xlim=xlim, ylim=ylim, breaks=breaks3, col=cols1, main=paste0("settle_08 ",names(settle_08)[8]))
```

```{r, fig.width=10, fig.height=10}
#### comparisons surf / sink / settle for Mertz
par(mfrow=c(3,2))
plot(susp_08[[3]], xlim=xlim, ylim=ylim, main=paste0("susp_08 ",names(susp_08)[3]))#, breaks=breaks2, col=cols2)
plot(settle_08[[8]], xlim=xlim, ylim=ylim, main=paste0("settle_08 ",names(settle_08)[8]))#, breaks=breaks3, col=cols1)
#plot(settle_08[[6]]-susp_08[[4]], xlim=xlim, ylim=ylim)
plot(settle_08[[8]]-settle_08[[1]], xlim=xlim, ylim=ylim, main="sedimentation across 8 timesteps")
plot(susp_08[[3]]-(settle_08[[8]]-settle_08[[1]]), xlim=xlim, ylim=ylim, main="mean-suspended minus sedimentation across 8 timesteps")

# plot((settle_08[[31]]-settle_08[[21]])/10, xlim=xlim, ylim=ylim, main="mean sedimentation last 10 timesteps")
plot(susp_08[[3]]-(settle_08[[8]]-settle_08[[1]])/10, xlim=xlim, ylim=ylim, main="mean-suspended minus mean sedimentation for last third")
```


```{r, fig.height=8.6, fig.width=10}
plot(settle_08[[31]]-settle_08[[1]], main="sedimentation across 1 month at 100m/day particle sinking speed", breaks=seq(-5.2,5.2, length.out=53), col=scico(52,palette="vik"))
```

```{r, fig.height=8.6, fig.width=10}
plot(abs.uv, main="seafloor currents - absolute speed", breaks=seq(0,0.8,length.out=81), col=rev(magma(80)))
```


```{r, fig.width=10, fig.height=10}
#### comparisons surf / sink / settle for Mertz
par(mfrow=c(3,2))
plot(susp_his_08[[31]], xlim=xlim, ylim=ylim, main="suspended timestep 31")#, breaks=breaks2, col=cols2)
plot(settle_08[[31]], xlim=xlim, ylim=ylim, main="settled timestep 31")#, breaks=breaks3, col=cols1)
plot(susp_his_08[[31]]-(settle_08[[31]]-settle_08[[1]]), xlim=xlim, ylim=ylim, main="suspended at 31 minus settled across 31 timesteps")
plot(settle_08[[31]]-settle_08[[1]], xlim=xlim, ylim=ylim, main="settled across 31 timesteps")

plot((susp_his_08[[31]]-susp_his_08[[21]])/10, xlim=xlim, ylim=ylim, main="suspended across last 10 timesteps")
plot((settle_08[[31]]-settle_08[[21]])/10, xlim=xlim, ylim=ylim, main="settled across last 10 timesteps")

# plot(susp_his_08[[31]]-surf_his_08[[31]], xlim=xlim, ylim=ylim, main="suspended at 31 minus surf at 31")

```


```{r, fig.height=10, fig.width=10}
h.mertz <- crop(h2, extent(c(xlim,ylim)))
surf.circum <- surf_his_08[[31]]-surf_his_08[[1]]
surf.mertz <- crop(surf.circum, extent(c(xlim,ylim)))
susp.circum <- susp_his_08[[31]]-susp_his_08[[1]]
susp.mertz <- crop(susp.circum, extent(c(xlim,ylim)))
settled.circum <- settle_08[[31]]-settle_08[[1]]
settled.mertz <- crop(settled.circum, extent(c(xlim,ylim)))

par(mfrow=c(2,2))
plot(surf.mertz[]~susp.mertz[])
plot(surf.mertz[]~settled.mertz[])
plot(susp.mertz[]~settled.mertz[])
plot(h.mertz[]~settled.mertz[])
```

## Comparison to Diatom values ##########

```{r, fig.height=8.5, fig.width=10}
plot(h2)
plot(polar.dat.diatom, add=TRUE)
```

```{r, fig.width=10, fig.height=6}
#### Extract flux values at Diatom locations
## circumpolar
surf.vals <- extract(surf_08, polar.dat.diatom)
susp.vals <- extract(susp_08, polar.dat.diatom)
settle.vals <- extract(settle_08, polar.dat.diatom)
diatom.vals <- diatom.dat$Absolute_diatom_abundance_in_million_valves_per_gram_dry_sediment
## Mertz region
surf.vals.m <- extract(surf_08,mertz.diatom.locs)
susp.vals.m <- extract(susp_08,mertz.diatom.locs)
settle.vals.m <- extract(settle_08,mertz.diatom.locs)
diatom.vals.m <- mertz.diatom.dat$Abundance_NO_benthics__valves_g_dry_weight_

#### circumpolar relationship
na.sel <- c(62,75,80,81,105,108,181)
par(mfrow=c(2,3))
plot(surf.vals[,3], diatom.vals)
legend("topleft", legend=paste0("cor=",round(cor(diatom.vals[-na.sel],surf.vals[-na.sel,3]),3)))
plot(susp.vals[,3], diatom.vals)
legend("topleft", legend=paste0("cor=",round(cor(diatom.vals[-na.sel],susp.vals[-na.sel,3]),3)))
plot(settle.vals[,30], diatom.vals)
legend("topleft", legend=paste0("cor=",round(cor(diatom.vals[-na.sel],settle.vals[-na.sel,30]),3)))
#### compare to Mertz data:
plot(surf.vals.m[,3], diatom.vals.m)
legend("topleft", legend=paste0("cor=",round(cor(diatom.vals.m,surf.vals.m[,3]),3)))
plot(susp.vals.m[,3], diatom.vals.m)
legend("topleft", legend=paste0("cor=",round(cor(diatom.vals.m,susp.vals.m[,3]),3)))
plot(settle.vals.m[,30], diatom.vals.m)
legend("topleft", legend=paste0("cor=",round(cor(diatom.vals.m,settle.vals.m[,30]),3)))

# #### circumpolar relationship
# library(modEvA)
# library(MASS)
# par(mfrow=c(2,3))
# plot(surf.vals[,5], log(diatom.vals))
# plot(susp.vals[,5], log(diatom.vals))
# plot(settle.vals[,5], log(diatom.vals))
# #### compare to Mertz data:
# plot(surf.vals.m[,5], log(diatom.vals.m))
# plot(susp.vals.m[,5], log(diatom.vals.m))
# plot(settle.vals.m[,5], log(diatom.vals.m))
# library(MASS)
# library(modEvA)
# summary(glm.nb(diatom.vals.m~surf.vals.m))
# summary(glm.nb(diatom.vals.m~susp.vals.m))
# summary(glm.nb(diatom.vals.m~settle.vals.m))
# Dsquared(glm.nb(diatom.vals~surf.vals))
# Dsquared(glm.nb(diatom.vals~susp.vals))
# Dsquared(glm.nb(diatom.vals~settle.vals))
```

## Comparison to TOC values ##########

```{r}
## all locations
polar.dat.toc <- SpatialPoints(rbind(mertz.toc.locs, ross.toc.locs, ant.toc.locs), proj4string = crs(polar.dat.diatom))
## all values
toc.vals <- c(mertz.toc.dat[,7],ross.toc.dat[,5],ant.toc.dat[,8])
depth.vals <- extract(h2, polar.dat.toc)
curr.vals <- extract(mean.uv, polar.dat.toc)
surf.vals.list <- list()
susp.vals.list <- list()
settle.vals.list <- list()
for(i in 1:8){
  surf.vals.list[[i]] <- extract(get(paste0("surf_0",i)),polar.dat.toc)[,3]
  susp.vals.list[[i]] <- extract(get(paste0("susp_0",i)),polar.dat.toc)[,3]
  settle.vals.list[[i]] <- extract(get(paste0("settle_0",i)),polar.dat.toc)[,30]-extract(get(paste0("settle_0",i)),polar.dat.toc)[,1]
}
```

### overview of sampling locations

```{r, fig.height=8.5, fig.width=10}
## plot overview
plot(h2)
plot(polar.dat.toc, add=TRUE)
```

### relationship with depth & curr

```{r}
par(mfrow=c(1,2), mar=c(5,4,4,1))
plot(depth.vals, toc.vals, main="toc vs depth")
plot(curr.vals, toc.vals, main="toc vs mean current speed")
```


### relationship with surf, susp & settle

```{r, fig.width=10, fig.height=8}
o <- toc.vals
par(mfrow=c(3,5), mar=c(3,2,2,1))
for(i in 4:8){
  p <- surf.vals.list[[i]]
  m <- "surf"
  plot(p, o, main=m)
  legend("topleft", legend=c(paste0("cor=",round(cor(o, p),3)),
                             paste0("r2=",round(summary(lm(o~p))$r.squared,3))),
         bty="n")
}
for(i in 4:8){
  p <- susp.vals.list[[i]]
  m <- "susp"
  plot(p, o, main=m)
  legend("topleft", legend=c(paste0("cor=",round(cor(o, p),3)),
                             paste0("r2=",round(summary(lm(o~p))$r.squared,3))),
         bty="n")
}
for(i in 4:8){
  p <- settle.vals.list[[i]]
  m <- "settle"
  plot(p, o, main=m)
  legend("topleft", legend=c(paste0("cor=",round(cor(o, p),3)),
                             paste0("r2=",round(summary(lm(o~p))$r.squared,3))),
         bty="n")
}
```

### relationship with surf, susp & settle - by region - settling speed 5

East Antarctica
Mertz
Ross Sea
Amundsen Sea
Antarctic Peninsula

```{r, fig.width=10, fig.height=8}
susp <- susp.vals.list[[5]]
settle <- settle.vals.list[[5]]

reg.east <- which(polar.dat.toc@coords[,1]>2000000)
reg.mertz <- which(polar.dat.toc@coords[,1]>1000000 & polar.dat.toc@coords[,1]<2000000)
reg.ross <- which(polar.dat.toc@coords[,1]>(-1000000) & polar.dat.toc@coords[,1]<1000000)
reg.amund <- which(polar.dat.toc@coords[,1]>(-2000000) & polar.dat.toc@coords[,1]<(-1000000))
reg.ap <- which(polar.dat.toc@coords[,1]<(-2100000))
reg.list <- list(reg.east, reg.mertz, reg.ross, reg.amund, reg.ap)
names(reg.list) <- c("East Antarctica", "Mertz", "Ross Sea", "Amundsen", "AP")

par(mfrow=c(3,5), mar=c(3,2,2,1))
for(i in 1:5){
  p <- surf.vals.list[[5]][reg.list[[i]]]
  o <- toc.vals[reg.list[[i]]]
  m <- paste0("surf, ",names(reg.list)[i])
  plot(p, o, main=m)
  legend("topleft", legend=c(paste0("cor=",round(cor(o, p),3)),
                             paste0("r2=",round(summary(lm(o~p))$r.squared,3))),
         bty="n")
}
for(i in 1:5){
  p <- susp[reg.list[[i]]]
  o <- toc.vals[reg.list[[i]]]
  m <- paste0("susp, ",names(reg.list)[i])
  plot(p, o, main=m)
  legend("topleft", legend=c(paste0("cor=",round(cor(o, p),3)),
                             paste0("r2=",round(summary(lm(o~p))$r.squared,3))),
         bty="n")
}
for(i in 1:5){
  p <- settle[reg.list[[i]]]
  o <- toc.vals[reg.list[[i]]]
  m <- paste0("settle, ",names(reg.list)[i])
  plot(p, o, main=m)
  legend("topleft", legend=c(paste0("cor=",round(cor(o, p),3)),
                             paste0("r2=",round(summary(lm(o~p))$r.squared,3))),
         bty="n")
}
```


## East Antarctica - underpredicting near the Amery Ice Shelf

```{r, fig.height=8, fig.width=10}
sel <- reg.east
xlim <- c(2000000,3000000)
ylim1 <- c(-1500000,0)
ylim2 <- c(0,1500000)
##

cex <- toc.vals[sel]*2
p <- settle.vals.list[[5]][sel]
o <- toc.vals[sel]
fit <- lm(o~p)
r2 <- summary(fit)$r.squared
top5 <- tail(order(sqrt(fit$residuals^2)))

par(mfrow=c(2,4), mar=c(2,2,2,0), oma=c(0,0,0,2))

plot(h2, xlim=xlim, ylim=ylim1, main="depth", legend=FALSE)
points(polar.dat.toc[sel], cex=cex)
points(polar.dat.toc[sel[top5]], cex=cex[top5], col="red")

plot(h2, xlim=xlim, ylim=ylim2, main="depth")
points(polar.dat.toc[sel], cex=cex)
points(polar.dat.toc[sel[top5]], cex=cex[top5], col="red")

plot(settle_05[[30]]-settle_05[[1]], xlim=xlim, ylim=ylim1, main="settle", legend=FALSE)
points(polar.dat.toc[sel], cex=cex)
points(polar.dat.toc[sel[top5]], cex=cex[top5], col="red")

plot(settle_05[[30]]-settle_05[[1]], xlim=xlim, ylim=ylim2, main="settle")
points(polar.dat.toc[sel], cex=cex)
points(polar.dat.toc[sel[top5]], cex=cex[top5], col="red")

plot(mean.uv, xlim=xlim, ylim=ylim1, main="mean current", legend=FALSE)
points(polar.dat.toc[sel], cex=cex)
points(polar.dat.toc[sel[top5]], cex=cex[top5], col="red")

plot(mean.uv, xlim=xlim, ylim=ylim2, main="mean current")
points(polar.dat.toc[sel], cex=cex)
points(polar.dat.toc[sel[top5]], cex=cex[top5], col="red")

plot(p, o, main="toc vs settle")
points(p[top5], o[top5], col="red")
abline(fit, lty=2)
legend("topleft", legend=c(paste0("cor=",round(cor(o, p),3)),
                             paste0("r2=",round(r2,3))),
         bty="n")
```

```{r, fig.height=8, fig.width=10}
sel <- reg.east
xlim <- c(2100000,2500000)
ylim <- c(800000,1200000)
##

cex <- toc.vals[sel]*2
p <- settle.vals.list[[5]][sel]
o <- toc.vals[sel]
fit <- lm(o~p)
r2 <- summary(fit)$r.squared
top5 <- tail(order(sqrt(fit$residuals^2)))

par(mfrow=c(2,2), mar=c(2,2,2,0), oma=c(0,0,0,2))

plot(h2, xlim=xlim, ylim=ylim, main="depth")
points(polar.dat.toc[sel], cex=cex)
points(polar.dat.toc[sel[top5]], cex=cex[top5], col="red")

plot(settle_05[[30]]-settle_05[[1]], xlim=xlim, ylim=ylim, main="settle")
points(polar.dat.toc[sel], cex=cex)
points(polar.dat.toc[sel[top5]], cex=cex[top5], col="red")

plot(mean.uv, xlim=xlim, ylim=ylim, main="mean current")
points(polar.dat.toc[sel], cex=cex)
points(polar.dat.toc[sel[top5]], cex=cex[top5], col="red")

plot(p, o, main="toc vs settle")
points(p[top5], o[top5], col="red")
abline(fit, lty=2)
legend("topleft", legend=c(paste0("cor=",round(cor(o, p),3)),
                             paste0("r2=",round(r2,3))),
         bty="n")
```

## Mertz - underpredicting just off the adelie bank

```{r, fig.height=7, fig.width=10}
sel <- reg.mertz
xlim <- c(1350000,1800000)
ylim <- c(-2200000,-1900000)
##

cex <- toc.vals[sel]*2
p <- settle.vals.list[[5]][sel]
o <- toc.vals[sel]
fit <- lm(o~p)
r2 <- summary(fit)$r.squared
top5 <- tail(order(sqrt(fit$residuals^2)))

par(mfrow=c(2,2), mar=c(2,3,2,1))

plot(h2, xlim=xlim, ylim=ylim, main="depth")
points(polar.dat.toc[sel], cex=cex)
points(polar.dat.toc[sel[top5]], cex=cex[top5], col="red")

plot(settle_05[[30]]-settle_05[[1]], xlim=xlim, ylim=ylim, main="settle")
points(polar.dat.toc[sel], cex=cex)
points(polar.dat.toc[sel[top5]], cex=cex[top5], col="red")

plot(mean.uv, xlim=xlim, ylim=ylim, main="mean current")
points(polar.dat.toc[sel], cex=cex)
points(polar.dat.toc[sel[top5]], cex=cex[top5], col="red")

plot(p, o, main="toc vs settle")
points(p[top5], o[top5], col="red")
abline(fit, lty=2)
legend("topleft", legend=c(paste0("cor=",round(cor(o, p),3)),
                             paste0("r2=",round(r2,3))),
         bty="n")
```

## Ross Sea - underpredicted in the bay - overpredicted on the banks

```{r, fig.height=6, fig.width=10}
sel <- reg.ross
xlim <- c(-600000,600000)
ylim <- c(-1800000,-1200000)
##

cex <- toc.vals[sel]*2
p <- settle.vals.list[[5]][sel]
o <- toc.vals[sel]
fit <- lm(o~p)
r2 <- summary(fit)$r.squared
top5 <- tail(order(sqrt(fit$residuals^2)))

par(mfrow=c(2,2), mar=c(2,3,2,1))

plot(h2, xlim=xlim, ylim=ylim, main="depth")
points(polar.dat.toc[sel], cex=cex)
points(polar.dat.toc[sel[top5]], cex=cex[top5], col="red")

plot(settle_05[[30]]-settle_05[[1]], xlim=xlim, ylim=ylim, main="settle")
points(polar.dat.toc[sel], cex=cex)
points(polar.dat.toc[sel[top5]], cex=cex[top5], col="red")

plot(mean.uv, xlim=xlim, ylim=ylim, main="mean current")
points(polar.dat.toc[sel], cex=cex)
points(polar.dat.toc[sel[top5]], cex=cex[top5], col="red")

plot(p, o, main="toc vs settle")
points(p[top5], o[top5], col="red")
abline(fit, lty=2)
legend("topleft", legend=c(paste0("cor=",round(cor(o, p),3)),
                             paste0("r2=",round(r2,3))),
         bty="n")
```

## Amundsen: - vastly overestimating along the shelf break

```{r, fig.height=10, fig.width=10}
sel <- reg.amund
xlim <- c(-2100000,-1200000)
ylim <- c(-1300000,-250000)
##

cex <- toc.vals[sel]*2
p <- settle.vals.list[[5]][sel]
o <- toc.vals[sel]
fit <- lm(o~p)
r2 <- summary(fit)$r.squared
top5 <- tail(order(sqrt(fit$residuals^2)),10)

par(mfrow=c(2,2), mar=c(2,3,2,1))

plot(h2, xlim=xlim, ylim=ylim, main="depth")
points(polar.dat.toc[sel], cex=cex)
points(polar.dat.toc[sel[top5]], cex=cex[top5], col="red")

plot(settle_05[[30]]-settle_05[[1]], xlim=xlim, ylim=ylim, main="settle")
points(polar.dat.toc[sel], cex=cex)
points(polar.dat.toc[sel[top5]], cex=cex[top5], col="red")

plot(mean.uv, xlim=xlim, ylim=ylim, main="mean current")
points(polar.dat.toc[sel], cex=cex)
points(polar.dat.toc[sel[top5]], cex=cex[top5], col="red")

plot(p, o, main="toc vs settle")
points(p[top5], o[top5], col="red")
abline(fit, lty=2)
legend("topleft", legend=c(paste0("cor=",round(cor(o, p),3)),
                             paste0("r2=",round(r2,3))),
         bty="n")
```

```{r, fig.height=10, fig.width=10}
sel <- reg.amund
xlim <- c(-2100000,-1600000)
ylim <- c(-1000000,-500000)
##

cex <- toc.vals[sel]*2
p <- settle.vals.list[[8]][sel]
o <- toc.vals[sel]
fit <- lm(o~p)
r2 <- summary(fit)$r.squared
top5 <- tail(order(sqrt(fit$residuals^2)),10)

par(mfrow=c(2,2), mar=c(2,3,2,1))

plot(terrain(mean.uv), xlim=xlim, ylim=ylim, main="current gradient")
points(polar.dat.toc[sel], cex=cex)
points(polar.dat.toc[sel[top5]], cex=cex[top5], col="red")

plot(settle_08[[30]]-settle_08[[1]], xlim=xlim, ylim=ylim, main="settle")
points(polar.dat.toc[sel], cex=cex)
points(polar.dat.toc[sel[top5]], cex=cex[top5], col="red")

plot(mean.uv, xlim=xlim, ylim=ylim, main="mean current")
points(polar.dat.toc[sel], cex=cex)
points(polar.dat.toc[sel[top5]], cex=cex[top5], col="red")

plot(p, o, main="toc vs settle")
points(p[top5], o[top5], col="red")
abline(fit, lty=2)
legend("topleft", legend=c(paste0("cor=",round(cor(o, p),3)),
                             paste0("r2=",round(r2,3))),
         bty="n")
```

```{r, fig.height=10, fig.width=10}
sel <- reg.amund
xlim <- c(-2050000,-1900000)
ylim <- c(-750000,-500000)
##

cex <- toc.vals[sel]*2
p <- settle.vals.list[[8]][sel]
o <- toc.vals[sel]
fit <- lm(o~p)
r2 <- summary(fit)$r.squared
top5 <- tail(order(sqrt(fit$residuals^2)),10)

par(mfrow=c(2,2), mar=c(2,3,2,1))

plot(terrain(mean.uv), xlim=xlim, ylim=ylim, main="current gradient")
points(polar.dat.toc[sel], cex=cex)
points(polar.dat.toc[sel[top5]], cex=cex[top5], col="red")

plot(settle_08[[30]]-settle_08[[1]], xlim=xlim, ylim=ylim, main="settle")
points(polar.dat.toc[sel], cex=cex)
points(polar.dat.toc[sel[top5]], cex=cex[top5], col="red")

plot(mean.uv, xlim=xlim, ylim=ylim, main="mean current")
points(polar.dat.toc[sel], cex=cex)
points(polar.dat.toc[sel[top5]], cex=cex[top5], col="red")

plot(p, o, main="toc vs settle")
points(p[top5], o[top5], col="red")
abline(fit, lty=2)
legend("topleft", legend=c(paste0("cor=",round(cor(o, p),3)),
                             paste0("r2=",round(r2,3))),
         bty="n")
```


## AP: - mostly bad or underpredicted

```{r, fig.height=10, fig.width=10}
sel <- reg.ap
xlim <- c(-2700000,-2000000)
ylim <- c(500000,1800000)
##

cex <- toc.vals[sel]*2
p <- settle.vals.list[[5]][sel]
o <- toc.vals[sel]
fit <- lm(o~p)
r2 <- summary(fit)$r.squared
top5 <- tail(order(sqrt(fit$residuals^2)))

par(mfrow=c(2,2), mar=c(2,3,2,1))

plot(h2, xlim=xlim, ylim=ylim, main="depth")
points(polar.dat.toc[sel], cex=cex)
points(polar.dat.toc[sel[top5]], cex=cex[top5], col="red")

plot(settle_05[[30]]-settle_05[[1]], xlim=xlim, ylim=ylim, main="settle")
points(polar.dat.toc[sel], cex=cex)
points(polar.dat.toc[sel[top5]], cex=cex[top5], col="red")

plot(mean.uv, xlim=xlim, ylim=ylim, main="mean current")
points(polar.dat.toc[sel], cex=cex)
points(polar.dat.toc[sel[top5]], cex=cex[top5], col="red")

plot(p, o, main="toc vs settle")
points(p[top5], o[top5], col="red")
abline(fit, lty=2)
legend("topleft", legend=c(paste0("cor=",round(cor(o, p),3)),
                             paste0("r2=",round(r2,3))),
         bty="n")
```

## AP - tip: generally underpredicted

```{r, fig.height=10, fig.width=10}
sel <- reg.ap
xlim <- c(-2700000,-2100000)
ylim <- c(1200000,1800000)
##

cex <- toc.vals[sel]*2
p <- settle.vals.list[[5]][sel]
o <- toc.vals[sel]
fit <- lm(o~p)
r2 <- summary(fit)$r.squared
top5 <- tail(order(sqrt(fit$residuals^2)))

par(mfrow=c(2,2), mar=c(2,3,2,1))

plot(h2, xlim=xlim, ylim=ylim, main="depth")
points(polar.dat.toc[sel], cex=cex)
points(polar.dat.toc[sel[top5]], cex=cex[top5], col="red")

plot(settle_05[[30]]-settle_05[[1]], xlim=xlim, ylim=ylim, main="settle")
points(polar.dat.toc[sel], cex=cex)
points(polar.dat.toc[sel[top5]], cex=cex[top5], col="red")

plot(mean.uv, xlim=xlim, ylim=ylim, main="mean current")
points(polar.dat.toc[sel], cex=cex)
points(polar.dat.toc[sel[top5]], cex=cex[top5], col="red")

plot(p, o, main="toc vs settle")
points(p[top5], o[top5], col="red")
abline(fit, lty=2)
legend("topleft", legend=c(paste0("cor=",round(cor(o, p),3)),
                             paste0("r2=",round(r2,3))),
         bty="n")
```

## All:

```{r, fig.height=10, fig.width=10}
##
cex <- toc.vals*2
p <- settle.vals.list[[5]]
o <- toc.vals
fit <- lm(o~p)
r2 <- summary(fit)$r.squared
over10 <- tail(order(fit$residuals),10)
under10 <- head(order(fit$residuals),10)

par(mfrow=c(2,2), mar=c(2,3,2,1))

plot(h2, main="depth")
points(polar.dat.toc, cex=cex)
points(polar.dat.toc[under10], cex=cex[under10], col="blue")
points(polar.dat.toc[over10], cex=cex[over10], col="red")

plot(settle_05[[30]]-settle_05[[1]], main="settle")
points(polar.dat.toc, cex=cex)
points(polar.dat.toc[under10], cex=cex[under10], col="blue")
points(polar.dat.toc[over10], cex=cex[over10], col="red")

plot(mean.uv, main="mean current")
points(polar.dat.toc, cex=cex)
points(polar.dat.toc[under10], cex=cex[under10], col="blue")
points(polar.dat.toc[over10], cex=cex[over10], col="red")

plot(p, o, main="toc vs settle")
points(p[over10], o[over10], col="red")
points(p[under10], o[under10], col="blue")
abline(fit, lty=2)
legend("topleft", legend=c(paste0("cor=",round(cor(o, p),3)),
                             paste0("r2=",round(r2,3))),
         bty="n")
```

## Residuals vs depth:

```{r, fig.height=10, fig.width=10}
par(mfrow=c(2,2), mar=c(5,4,4,1))
plot(depth.vals,o, main="depth")
points(depth.vals[under10],o[under10], col="blue")
points(depth.vals[over10],o[over10], col="red")
plot(curr.vals, o, main="current speed")
points(curr.vals[under10],o[under10], col="blue")
points(curr.vals[over10], o[over10], col="red")
plot(p,o)
points(p[under10],o[under10], col="blue")
points(p[over10] ,o[over10],  col="red")


```


## Current speeds vs settle:

```{r, fig.height=6, fig.width=6}
settle.ra.vals <- settle_05[[30]][]-settle_05[[1]][]
par(mfrow=c(1,1), mar=c(5,4,4,1))
plot(mean.uv[], settle.ra.vals, main="settle vs current speed", cex=0.1)
```

## Current speeds vs settle: on the shelf only

```{r, fig.height=6, fig.width=10}
shelf.sel <- which(h2[]>2000)
settle.ra.vals.shelf <- settle.ra.vals[][-shelf.sel]
uv.shelf <- mean.uv[][-shelf.sel]
uv.shelf.abs <- abs.uv[][-shelf.sel]

## maximum current speeds
nas <- which(is.na(settle.ra.vals.shelf)|is.na(uv.shelf.abs))
s <- sample((1:length(settle.ra.vals.shelf))[-nas], 200000)
abs.line <- smooth.spline(uv.shelf.abs[s], settle.ra.vals.shelf[s])

## mean current speeds
nas <- which(is.na(settle.ra.vals.shelf)|is.na(uv.shelf))
s <- sample((1:length(settle.ra.vals.shelf))[-nas], 200000)
mean.line <- smooth.spline(uv.shelf[s], settle.ra.vals.shelf[s])
```

```{r, fig.height=6, fig.width=10}
par(mfrow=c(1,1), mar=c(5,4,4,1))
plot(uv.shelf, settle.ra.vals.shelf, main="settle vs current speed (mean)", cex=0.1)
lines(mean.line, col = "red", lwd = 2)
```


```{r, fig.height=6, fig.width=10}
par(mfrow=c(1,1), mar=c(5,4,4,1))
plot(uv.shelf.abs, settle.ra.vals.shelf, main="settle vs current speed (abs)", cex=0.1)
lines(abs.line, col="red", lwd=2)
```



































```{r, fig.width=10, fig.height=8, eval=FALSE, echo=FALSE}
## sort diatom cruises by region
dc.ap.w <- c("DF82","DF86","LMG1311","NBP1001","PD88-III","PD90-7")
dc.ap.e <- c("LMG0502","NBP0003","NBP0603","NBP1203")
dc.ap.both <- c("NBP0107")
dc.eant <- c("NBP1402", "NBP0101")
dc.ross <- c("DF83 III", "NBP9501", "DF83 II", "NBP9401", "PD92")
dc.amund <- "NBP0702"

dat <- data.frame("surf.vals"=surf.vals[,5], "diatom.vals"=diatom.vals)
dat$cruise <- diatom.dat$Cruise
dat$cruise.short <- strtrim(diatom.dat$Cruise,3)
dat$region <- "empty"
dat[diatom.dat$Cruise%in%dc.ap.w,5] <- "ap.w"
dat[diatom.dat$Cruise%in%dc.ap.e,5] <- "ap.e"
dat[diatom.dat$Cruise%in%dc.ap.both,5] <- "ap.both"
dat[diatom.dat$Cruise%in%dc.eant,5] <- "e.eant"
dat[diatom.dat$Cruise%in%dc.ross,5] <- "e.ross"
dat[diatom.dat$Cruise%in%dc.amund,5] <- "e.amund"
dat$reg.cruise <- paste0(dat$region,"_",dat$cruise)

library(dplyr)
dat2 <- arrange(dat, dat$reg.cruise)

library(ggplot2)
ggplot(dat2, aes(x=surf.vals, y=log(diatom.vals), color=region)) +
  geom_point() +
  facet_wrap(~reg.cruise)

ggplot(dat2, aes(x=surf.vals, y=diatom.vals, color=region)) +
  geom_point() +
  facet_wrap(~reg.cruise)

```
