---
title: "ROMS FAM 2km analysis"
author: "Jan Jansen"
output: html_document
---

###########################################
########## Evaluate ROMS outputs ##########
###########################################

```{r setup, include=FALSE}
#### setup environment
library(terra)
library(viridis)
fam.dir <- "C:/Users/jjansen/OneDrive - University of Tasmania/science/data_environmental/FAM_outputs"
env.derived <- "C:/Users/jjansen/OneDrive - University of Tasmania/science/data_environmental/derived/"
ant.proj <- "+proj=stere +lat_0=-90 +lat_ts=-71 +lon_0=0 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs"

## load bathymetry and current speed
r.stack <- rast(paste0(env.derived,"Circumpolar_EnvData_2km_shelf_mask_unscaled_variables.tif"))
r.d <- r.stack$depth
r.c <- r.stack$seafloorcurrents_mean
r.t <- r.stack$tpi11
r.npp <- r.stack$npp_mean

# env.dir <- "E:/science/data_environmental/"
# env.raw <- paste0(env.dir,"raw/")
# AAD_dir <- paste0(env.raw,"accessed_through_R")

load("C:/Users/jjansen/OneDrive - University of Tasmania/science/data_environmental/Circumpolar_coastline.Rdata")
```

# Prepare data

```{r, fig.dim=10}
npp.ra.list <- list()
track3D.ra.list <- list()
sed.list1 <- list()
sed.list2 <- list()
sed.list5 <- list()
sed.list05 <- list()
sed.list.sum <- list()
sed.list.sum2 <- list()
flux.list1 <- list()
flux.list2 <- list()
flux.list5 <- list()
flux.list05 <- list()
flux.list.sum <- list()

tif.names <- list.files(fam.dir, full.names=TRUE, pattern=".tif")
# flux.names <- list.files(fam.dir, full.names=TRUE, pattern="flux.tif")
# sed.names <- list.files(fam.dir, full.names=TRUE, pattern="sed.tif")
npp.names <- list.files(fam.dir, full.names=TRUE, pattern= "_NPP_")
npp3D.names <- list.files(fam.dir, full.names=TRUE, pattern= "bottomnpp.tif")
flux.names1  <- list.files(fam.dir, full.names=TRUE, pattern= "r0001_28days_flux.tif")
flux.names2  <- list.files(fam.dir, full.names=TRUE, pattern= "r0002_28days_flux.tif")
flux.names5  <- list.files(fam.dir, full.names=TRUE, pattern= "r0005_28days_flux.tif")
flux.names05 <- list.files(fam.dir, full.names=TRUE, pattern="r00005_28days_flux.tif")
sed.names1  <- list.files(fam.dir, full.names=TRUE, pattern= "r0001_28days_sed.tif")
sed.names2  <- list.files(fam.dir, full.names=TRUE, pattern= "r0002_28days_sed.tif")
sed.names5  <- list.files(fam.dir, full.names=TRUE, pattern= "r0005_28days_sed.tif")
sed.names05 <- list.files(fam.dir, full.names=TRUE, pattern="r00005_28days_sed.tif")

for(i in 1:10){
  npp.ra.list[[i]] <- rast(npp.names[i])
  sed.list1[[i]]   <- rast(sed.names1[i])
  sed.list2[[i]]   <- rast(sed.names2[i])
  sed.list5[[i]]   <- rast(sed.names5[i])
  sed.list05[[i]]  <- rast(sed.names05[i])
  flux.list1[[i]]   <- rast(flux.names1[i])
  flux.list2[[i]]   <- rast(flux.names2[i])
  flux.list5[[i]]   <- rast(flux.names5[i])
  flux.list05[[i]]  <- rast(flux.names05[i])
  track3D.ra.list[[i]] <- rast(npp3D.names[i])
  s <- sds(sed.list5[[i]],sed.list2[[i]],sed.list1[[i]],sed.list05[[i]])
  sed.list.sum[[i]] <- app(s, fun="sum", na.rm=TRUE)
  s2 <- sds(track3D.ra.list[[i]],sed.list5[[i]],sed.list2[[i]],sed.list1[[i]],sed.list05[[i]])
  sed.list.sum2[[i]] <- app(s2, fun="sum", na.rm=TRUE)
  f <- sds(flux.list5[[i]],flux.list2[[i]],flux.list1[[i]],flux.list05[[i]])
  flux.list.sum[[i]] <- app(f, fun="sum", na.rm=TRUE)
}

#cols=rev(magma(99))
#cols=rev(viridis(99))
library(RColorBrewer)
cols.lower <- colorRampPalette(c("#F2F2F2","#FDE725FF"))
cols=c(cols.lower(10), rev(viridis(99)))
```

# NPP, sinking and settling per region

```{r, fig.height=10, fig.width=10}
for(i in 1:10){
par(mfrow=c(4,2))
plot(npp.ra.list[[i]], range=c(0,1500), main=paste0("Region ",i," surface NPP"), col=cols)
plot(track3D.ra.list[[i]], range=c(0,1500), main=paste0("Region ",i," sinking model output"), col=cols)
plot(sed.list5[[i]], range=c(0,1500), main=paste0("Region ",i," settling model 0005 output"), col=cols)
plot(sed.list2[[i]], range=c(0,1500), main=paste0("Region ",i," settling model 0002 output"), col=cols)
plot(sed.list1[[i]], range=c(0,1500), main=paste0("Region ",i," settling model 0001 output"), col=cols)
plot(sed.list05[[i]], range=c(0,1500), main=paste0("Region ",i," settling model 00005 output"), col=cols)
plot(sed.list.sum[[i]], range=c(0,6000), main=paste0("Region ",i," settling model summed output"), col=cols)
plot(sed.list.sum2[[i]], range=c(0,7500), main=paste0("Region ",i," settling model + sink summed output"), col=cols)
}
```

# Flux per region

```{r, fig.height=10, fig.width=10}
for(i in 1:10){
par(mfrow=c(1,1))
plot(flux.list05[[i]], main=paste0("Region ",i," flux model 00005 output"), col=cols)
plot(coast.proj, add=TRUE)
}
```

# Comparison with TOC

```{r}
## load TOC data
sed.dir <- "C:/Users/jjansen/OneDrive - University of Tasmania/science/data_biological/sediment_data/"
mertz.toc.dat <- read.csv(paste0(sed.dir,"Circumpolar_TOC_data_CEAMARC_sediment_isotopes.csv"))
mertz.toc.locs <- terra::project(as.matrix(mertz.toc.dat[,2:3]),from="+proj=longlat",to=ant.proj)
ross.toc.dat <- read.csv(paste0(sed.dir,"Circumpolar_TOC_data_villinski_2008.csv"))
ross.toc.locs <- terra::project(as.matrix(ross.toc.dat[,c(3,2)]),from="+proj=longlat",to=ant.proj)

## CHECK IF DATA IS MISSING (COORDS OR TOC IN FOLDER "biological_data/sediment data")
ap1 <- read.csv(paste0(sed.dir,"Circumpolar_TOC_data_AP_Isla_2018.csv"))
ap2 <- read.csv(paste0(sed.dir,"Circumpolar_TOC_data_AP_Isla_2021.csv"))
ap3 <- read.csv(paste0(sed.dir,"Circumpolar_TOC_data_AP_Taylor_2020.csv"))

ap.locs <- read.csv(paste0(sed.dir,"Circumpolar_TOC_coords_DeMaster_Taylor.csv"), skip=12)[-12,-1]
ap.locs$deg <- as.numeric(ap.locs$deg)
# Convert to decimal degrees
ap.locs$lat <- ap.locs$deg + ap.locs$min/60
ap.locs$lon <- -(ap.locs$deg.1 + ap.locs$min.1 / 60)  # Negative for West longitude
## values above don't match with other datasets, what's happening?

## other data (mostly peninsula)
oth.toc.dat.raw <- read.csv(paste0(sed.dir,"Circumpolar_TOC_data_Leventer.csv"), skip=4)[-c(90,95),] ## rows 90 and 95 have no coordinates
oth.toc.dat <- oth.toc.dat.raw[,8]
oth.toc.locs.raw <- read.csv(paste0(sed.dir,"Circumpolar_TOC_coords_Leventer.csv"))
oth.toc.locs <- terra::project(as.matrix(oth.toc.locs.raw[-c(90,95),c(5,4)]),from="+proj=longlat",to=ant.proj)

## COMBINE DATA FOR "ALL OF ANTARCTIC"
ant.toc.dat <- c(oth.toc.dat, ross.toc.dat[,5], mertz.toc.dat[,7])
ant.toc.locs <- rbind(oth.toc.locs, ross.toc.locs, mertz.toc.locs)
```

```{r}
## extract values at toc-locations
mertz.df <- data.frame("sink.vals"=rep(NA,nrow(mertz.toc.locs)))
mertz.df$sink.vals <- extract(track3D.ra.list[[9]],mertz.toc.locs)[,1]
mertz.df$sed5      <- extract(sed.list5[[9]],    mertz.toc.locs)[,1]
mertz.df$sed2      <- extract(sed.list2[[9]],    mertz.toc.locs)[,1]
mertz.df$sed1      <- extract(sed.list1[[9]],    mertz.toc.locs)[,1]
mertz.df$sed05     <- extract(sed.list05[[9]],   mertz.toc.locs)[,1]
mertz.df$sedsum    <- extract(sed.list.sum[[9]], mertz.toc.locs)[,1]
mertz.df$sedsum2   <- extract(sed.list.sum2[[9]],mertz.toc.locs)[,1]

ross.df <- data.frame("sink.vals"=rep(NA,nrow(ross.toc.locs)))
ross.df$sink.vals <- extract(track3D.ra.list[[1]],ross.toc.locs)[,1]
ross.df$sed5    <- extract(sed.list5[[1]],   ross.toc.locs)[,1]
ross.df$sed2    <- extract(sed.list2[[1]],   ross.toc.locs)[,1]
ross.df$sed1    <- extract(sed.list1[[1]],   ross.toc.locs)[,1]
ross.df$sed05   <- extract(sed.list05[[1]],  ross.toc.locs)[,1]
ross.df$sedsum  <- extract(sed.list.sum[[1]],ross.toc.locs)[,1]
ross.df$sedsum2<- extract(sed.list.sum2[[1]],ross.toc.locs)[,1]

## the below datasets need to loop through the 10 regions to fill the values
ant.df <- data.frame("sink.vals"=rep(NA,nrow(ant.toc.locs)))
ant.df$sink.vals <- extract(track3D.ra.list[[1]],ant.toc.locs)[,1]
ant.df$sed5    <- extract(sed.list5[[1]],   ant.toc.locs)[,1]
ant.df$sed2    <- extract(sed.list2[[1]],   ant.toc.locs)[,1]
ant.df$sed1    <- extract(sed.list1[[1]],   ant.toc.locs)[,1]
ant.df$sed05   <- extract(sed.list05[[1]],  ant.toc.locs)[,1]
ant.df$sedsum  <- extract(sed.list.sum[[1]],ant.toc.locs)[,1]
ant.df$sedsum2<- extract(sed.list.sum2[[1]],ant.toc.locs)[,1]

for(i in 2:10){
  #print(i)
  l <- extract(track3D.ra.list[[i]],ant.toc.locs)
  l5  <- extract(sed.list5[[i]], ant.toc.locs)
  l2  <- extract(sed.list2[[i]], ant.toc.locs)
  l1  <- extract(sed.list1[[i]], ant.toc.locs)
  l05 <- extract(sed.list05[[i]],ant.toc.locs)
  ls  <- extract(sed.list.sum[[i]],ant.toc.locs)
  ls2 <- extract(sed.list.sum2[[i]],ant.toc.locs)
  if(any(!is.na(l))){
    l.na.sel <- which(is.na(l))
    ant.df$sink.vals[-l.na.sel] <- l[-l.na.sel,]
  }  
  if(any(!is.na(l5))){
    l5.na.sel <- which(is.na(l5))
    ant.df$sed5[-l5.na.sel] <- l5[-l5.na.sel,]
  }
  if(any(!is.na(l2))){
    l2.na.sel <- which(is.na(l2))
    ant.df$sed2[-l2.na.sel] <- l2[-l2.na.sel,]
  }
  if(any(!is.na(l1))){
    l1.na.sel <- which(is.na(l1))
    ant.df$sed1[-l1.na.sel] <- l1[-l1.na.sel,]
  }
  if(any(!is.na(l05))){
    l05.na.sel <- which(is.na(l05))
    ant.df$sed05[-l05.na.sel] <- l05[-l05.na.sel,]
  }
  if(any(!is.na(ls))){
    ls.na.sel <- which(is.na(ls))
    ant.df$sedsum[-ls.na.sel] <- ls[-ls.na.sel,]
  }
  if(any(!is.na(ls2))){
    ls2.na.sel <- which(is.na(ls2))
    ant.df$sedsum2[-ls2.na.sel] <- ls2[-ls2.na.sel,]
  }
}

oth.df <- ant.df[1:nrow(oth.toc.locs),]
```

```{r, fig.height=10, fig.width=10}
main.names <- c("sinking", "settling 5", "settling 2", "settling 1", "settling 05", "settling sum", "settling+sink sum")

par(mfrow=c(3,3), mar=c(3,2,2,1))
x = ross.df
y = ross.toc.dat[,5]
for(i in 1:7){
  plot(x[,i], y, main= paste0("Ross ", main.names[i]))
  legend("topright",legend=paste0("cor= ",round(cor(x=x[,i], y=y, use="complete.obs"),2)), bty="n")
  legend("topleft", legend=paste0("R^2= ",round(summary(lm(y~x[,i]))$r.squared,2)), bty="n")
}

par(mfrow=c(3,3))
x = mertz.df
y = mertz.toc.dat[,7]
for(i in 1:7){
  plot(x[,i], y, main= paste0("Mertz ", main.names[i]))
  legend("topright", legend=paste0("cor= ",round(cor(x=x[,i], y=y, use="complete.obs"),2)), bty="n")
  legend("topleft", legend=paste0("R^2= ",round(summary(lm(y~x[,i]))$r.squared,2)), bty="n")
}

par(mfrow=c(3,3))
x = oth.df
y = oth.toc.dat
for(i in 1:7){
  plot(x[,i], y, main= paste0("Other ", main.names[i]))
  legend("topright", legend=paste0("cor= ",round(cor(x=x[,i], y=y, use="complete.obs"),2)), bty="n")
  legend("topleft", legend=paste0("R^2= ",round(summary(lm(y~x[,i]))$r.squared,2)), bty="n")
}

par(mfrow=c(3,3))
x = ant.df
y = ant.toc.dat
for(i in 1:7){
  plot(x[,i], y, main= paste0("All ", main.names[i]))
  legend("topright", legend=paste0("cor= ",round(cor(x=x[,i], y=y, use="complete.obs"),2)), bty="n")
  legend("topleft", legend=paste0("R^2= ",round(summary(lm(y~x[,i]))$r.squared,2)), bty="n")
}

par(mfrow=c(3,3))
x = log(ant.df)
y = log(ant.toc.dat)
for(i in 1:7){
  plot(x[,i], y, main= paste0("All log ", main.names[i]))
  legend("topright", legend=paste0("cor= ",round(cor(x=x[,i], y=y, use="complete.obs"),2)), bty="n")
  legend("topleft", legend=paste0("R^2= ",round(summary(lm(y~x[,i]))$r.squared,2)), bty="n")
}
```


## Region by region - Pred vs Observed on Map:

```{r, fig.height=10}
## observed values
cex_values <- sqrt(ant.toc.dat*10/pi)
## predicted-observed values (proportional to max value on each scale)
## positive means overprediction
pred.vals <- ant.df$sedsum/max(ant.df$sedsum, na.rm=TRUE)
obs.vals <- ant.toc.dat/max(ant.toc.dat)
res.vals <- pred.vals-obs.vals
neg.vals <- which(res.vals<0)
cex.vals.raw <- res.vals
cex.vals.raw[neg.vals] <- cex.vals.raw[neg.vals]*-1

cex.vals <- sqrt(cex.vals.raw*20/pi)

for(i in 1:10){
par(mfrow=c(1,1))
plot(sed.list.sum[[i]], range=c(0,6000), main=paste0("Region ",i," settling model summed output"), col=cols)
#points(ant.toc.locs, cex=cex.vals, col="red")
points(ant.toc.locs[neg.vals,], cex=cex.vals[neg.vals], col="red", lwd=3)
points(ant.toc.locs[-neg.vals,], cex=cex.vals[-neg.vals], col="purple", lwd=3)
add_legend("topright",col=c("red","purple"), legend=c("under-predicted", "over-predicted"), cex=1, pch=1)
}
```

## Region by region - Pred vs Observed vs depth and current speeds and tpi:

```{r, fig.height=10}
# 
#### load non-ROMS data
# ## load projected coastline for plotting
# load(paste0(env.derived,"Circumpolar_Coastline.Rdata"))
# ## load circumpolar Diatom data
# load(paste0(ARC_data_dir,"FAM/Circumpolar_Diatom_metadata.Rdata"))
# ## load Mertz Diatom data
# mertz.diatom.dat <- read.csv("C:/Users/jjansen/Desktop/science/PhD/1 bentho-pelagic-coupling/DiatomAbundance.csv")
# mertz.diatom.locs <- project(as.matrix(mertz.diatom.dat[,c(4,3)]),proj=ant.proj)


## depth and current speed
sample.depths <-        extract(r.d, ant.toc.locs)[,1]
sample.currents.mean <- extract(r.c, ant.toc.locs)[,1]
sample.currents.max <-  extract(r.stack$seafloorcurrents_max, ant.toc.locs)[,1]
sample.tpi <-           extract(r.t, ant.toc.locs)[,1]
par(mfrow=c(2,2))
plot(sample.depths,res.vals, main="depth vs proxy-residuals")
plot(sample.currents.mean,res.vals, main="currents mean vs proxy-residuals")
plot(sample.currents.max,res.vals, main="currents max vs proxy-residuals")
plot(sample.tpi,res.vals, main="tpi vs proxy-residuals")

#summary(lm(res.vals~sample.depths+sample.currents.max+sample.currents.mean))

```

```{r, fig.height=10, eval=FALSE}
cex=ross.toc.dat[,5]
plot(sed.list1[[1]], range=c(0,1500), main=paste0("Region 01 settling model output"))
points(ross.toc.locs, cex=cex)

plot(flux.list1[[1]], main=paste0("Region 01 particle flux"))
points(ross.toc.locs, cex=cex)

plot(log(flux.list1[[1]]), main=paste0("Region 01 particle flux"))
points(ross.toc.locs, cex=cex)
```

# Distances from sources to sinks, and from sinks to sources

Note that the Ross Sea raster is wrong, need to recreate the raster

```{r}
d.to.sink.avg.files <- list.files("C:/Users/jjansen/OneDrive - University of Tasmania/science/data_environmental/FAM_outputs/", pattern="disttosinkavg", full.names=TRUE)
d.to.sink.avg <- list()
d.sink.4k <- list()

d.to.source.avg.files <- list.files("C:/Users/jjansen/OneDrive - University of Tasmania/science/data_environmental/FAM_outputs/", pattern="disttosourceavg", full.names=TRUE)
d.to.source.avg <- list()
d.source.4k <- list()

for(i in 1:10){
  d.to.sink.avg[[i]] <- rast(d.to.sink.avg.files[i])
  d.to.source.avg[[i]] <- rast(d.to.source.avg.files[i])
  #plot(d.to.sink.avg[[i]])
  ## lower res:
  d.sink.4k[[i]] <- terra::aggregate(d.to.sink.avg[[i]], fact=2, na.rm=TRUE)
  d.source.4k[[i]] <- terra::aggregate(d.to.source.avg[[i]], fact=2, na.rm=TRUE)
  plot(d.sink.4k[[i]],   col=cols, main=paste0("distance to sink - Region ",i))
  plot(d.source.4k[[i]], col=cols, main=paste0("distance to source - Region ",i))
}
```






# 3D representations of particle flux and trajectories

#### setup data

```{r}
## load flux and sedimentation
RS.s <- sed.list.sum[[1]]
RS.s[RS.s>1500] <- 1500
RS.f <- flux.list.sum[[1]]
RS.f[RS.f>1000000] <- 1000000

## match raster dimensions
RS.d <- crop(r.d,RS.s)
RS.c <- crop(r.c,RS.s)
RS.t <- crop(r.t,RS.s)
RS.npp <- crop(r.npp,RS.s)
```

```{r}
#### load trajectories
#load(paste0(fam.dir,"/tracking2D_Region01_NPP9_200mday_21days_traj_r0001_28days_Region01.Rdata"))
load(paste0(fam.dir,"/tracking2D_Region01_NPP1_200mday_21days_traj_28days.Rdata"))
#track.2D$ptrack_x_list
#track.2D$ptrack_y_list

## x-coords for each particle (when looping, need to keep particle ids in mind)
x.crds <- matrix(track.2D$ptrack_x_list[[1]])
y.crds <- matrix(track.2D$ptrack_y_list[[1]])
for(i in 2:113){
  #print(i) 
  fill.here <- track.2D$id_list[[i]]
  x.crds <- cbind(x.crds, NA)
  x.crds[fill.here,i] <- track.2D$ptrack_x_list[[i]]
  y.crds <- cbind(y.crds, NA)
  y.crds[fill.here,i] <- track.2D$ptrack_y_list[[i]]
}

#### translate coordinates to lines

## Combine particle locations into a data frame
df <- data.frame(
  particle = rep(1:nrow(x.crds), each=ncol(x.crds)),
  time = rep(1:ncol(x.crds), times=nrow(x.crds)),
  x = as.vector(t(x.crds)),
  y = as.vector(t(y.crds))
)
df <- na.omit(df) ## some particles didn't make it to the end, for whatever reason I don't know

# ## Add lines to the existing plot, coloring by time (veeeeery slow)
# particles <- unique(df$particle)
# colors <- rainbow(max(df$time))
# plot(RS.d)
# for (particle in particles) {
#   particle_data <- df[df$particle == particle, ]
#   for (i in 1:(nrow(particle_data) - 1)) {
#     lines(particle_data$x[i:(i+1)], particle_data$y[i:(i+1)], col = colors[particle_data$time[i]])
#   }
# }

library(sf)
## Create a list of line geometries
lines_list <- lapply(split(df, df$particle), function(p) {
  if (nrow(p) > 1) {  ## Ensure there are at least two points
    st_linestring(as.matrix(p[, c("x", "y")]))
  } else {
    NULL
  }
})
## Remove NULL entries
valid_indices <- !sapply(lines_list, is.null)
lines_list <- lines_list[valid_indices]
## Create an sf object
lines_sf <- st_sfc(lines_list, crs = ant.proj)
## Create an sf object with particle IDs
lines_sf <- st_sf(particle = unique(df$particle)[valid_indices], geometry = lines_sf)

lines_sf_subset <- lines_sf[seq(1,nrow(lines_sf),by=100),]

## Plot
plot(RS.d)
plot(lines_sf_subset, add=TRUE, col="black")

# ## Plot the lines
# library(ggplot2)
# ggplot() +
#   geom_sf(data = lines_sf) +
#   theme_minimal() +
#   labs(title = "Particle Movement")
## colouring lines by time takes long because each segment of the lines needs to be coloured differently
# # Add lines to the existing plot
# for(i in seq(1, nrow(lines_sf), by = 100)) {
#   coords <- st_coordinates(lines_sf[i, ])
#   lines(coords[, 1], coords[, 2], col = rainbow(nrow(coords))[coords[, 3]])
# }
# # Plot the lines with color by time
# ggplot() +
#   geom_sf(data = lines_sf[1:100,], aes(color = time)) +
#   theme_minimal() +
#   labs(title = "Movement of First 10 Particles Over Time", color = "Time")
```


```{r}
## plot 3D
library(rayshader)
library(raster)
library(sf)
##### change data to matrix
topo.dat <- RS.d
topo_mat=raster_to_matrix(RS.d)
curr_mat=raster_to_matrix(RS.c)
flux_mat=raster_to_matrix(RS.f)
sed_mat=raster_to_matrix(RS.s)

topo_mat[1:10,1:10]

sf_coast <- st_as_sf(coast.proj)
##### sample site information
## from "Readin_Circumpolar_DownwardImage_Data.Rmd"
# load(paste0(ARC_Data.dir,"Circumpolar_DownwardImages_metadata.Rdata"))
# sf_points <- st_as_sf(polar.dat.PS81)


##### plot basemap
# topo_mat %>%
#   height_shade() %>%
#   plot_map()
# 
# topo_mat %>%
#   sphere_shade() %>%
#   plot_map()
# 
# topo_mat %>%
#   # sphere_shade() %>%
#   height_shade() %>%
#   #add_water(detect_water(topo_mat, min_area=length(topo_mat)/10000), color="white") %>%
#   add_shadow(ray_shade(topo_mat, zscale=33, sunaltitude=6, lambert=FALSE), max_darken=0.5) %>%
#   add_shadow(lamb_shade(topo_mat, zscale=33, sunaltitude=6), max_darken=0.5) %>%
#   add_shadow(ambient_shade(topo_mat), max_darken=0.1) %>%
#   plot_map()

base_map2 = topo_mat %>%
  height_shade() %>%
  add_overlay(sphere_shade(topo_mat, texture = "desert", colorintensity = 3), alphalayer=0.5) %>%
  add_shadow(lamb_shade(topo_mat, zscale=33, sunaltitude=6), max_darken=0.3) %>%
  add_shadow(ambient_shade(topo_mat)) %>%
  # add_shadow(ray_shade(topo_mat, zscale=33, sunaltitude=6, lambert=FALSE), max_darken=0.5) %>%
  add_shadow(texture_shade(topo_mat,detail=6/10,contrast=10,brightness = 11), 0.1)# %>%
  #add_water(detect_water(topo_mat, min_area=length(topo_mat)/10000), color="white")
```

```{r, fig.height=10}
plot_map(base_map2)
```

## variables on 3D bathymetry

```{r}
zscale <- 100 ## the higher the less exaggerated the height values are
phi <- 40    ## vertical angle of view
theta <- 300 ## horizontal angle of view
zoom <- 0.7  ## zoom between 0 (close) and 1 (far)
windowsize <- c(1000,1000)

## calculate sphereshades and lineoverlays
sphereshades = sphere_shade(topo_mat, texture="desert", colorintensity=3)
rayshades = ray_shade(topo_mat, sunaltitude=6, zscale=zscale, lambert=FALSE)
lambshades = lamb_shade(topo_mat, sunaltitude=6, zscale=zscale)
ambientshades = ambient_shade(topo_mat)
textureshades = texture_shade(topo_mat,detail=6/10,contrast=10,brightness = 11)

currentoverlay  = height_shade(curr_mat, texture=rev(heat.colors(256)))
fluxoverlay     = height_shade(flux_mat, texture=rev(heat.colors(256)))
sedimentoverlay = height_shade(sed_mat,  texture=rev(heat.colors(256)))

coastoverlay = generate_line_overlay(sf_coast, extent=ext(topo.dat), heightmap=topo_mat)
tracksoverlay = generate_line_overlay(lines_sf_subset, extent=ext(topo.dat), heightmap=topo_mat)
#sampleoverlay = generate_point_overlay(sf_points, extent=extent(topo.dat), heightmap=topo_mat+500, color="deepskyblue")
#sampleoverlay_shades = generate_point_overlay(sf_points, extent=extent(topo.dat), heightmap=topo_mat, color="black", offset=c(1500,-1500))
```

### current speed

```{r, fig.height=15}
topo_mat %>%
  sphere_shade(texture = "bw") %>%
  add_overlay(currentoverlay, alphalayer = 0.6) %>% 
  add_overlay(coastoverlay) %>% 
  add_overlay(tracksoverlay) %>% 
#  add_overlay(sampleoverlay_shades) %>%   
#  add_overlay(sampleoverlay) %>%
  plot_map()
```

```{r, fig.height=15}
topo_mat %>%
  sphere_shade(texture = "bw") %>%
  add_overlay(currentoverlay, alphalayer = 0.6) %>%
    add_overlay(coastoverlay) %>% 
    add_overlay(tracksoverlay) %>% 
#  add_overlay(sampleoverlay_shades) %>%   
#  add_overlay(sampleoverlay) %>%
    plot_3d(topo_mat, zscale=zscale, phi=phi, theta=theta, zoom=zoom, 
          windowsize=windowsize)
render_snapshot()
```

### flux

```{r, fig.height=15}
topo_mat %>%
  sphere_shade(texture = "bw") %>%
  add_overlay(fluxoverlay, alphalayer = 0.6) %>%
  add_overlay(coastoverlay) %>% 
  add_overlay(tracksoverlay) %>% 
#  add_overlay(sampleoverlay_shades) %>%   
#  add_overlay(sampleoverlay) %>%
  plot_map()
```

```{r, fig.height=15}
topo_mat %>%
  sphere_shade(texture = "bw") %>%
  add_overlay(fluxoverlay, alphalayer = 0.6) %>%
  add_overlay(coastoverlay) %>% 
  add_overlay(tracksoverlay) %>% 
#  add_overlay(sampleoverlay_shades) %>%   
#  add_overlay(sampleoverlay) %>%
    plot_3d(topo_mat, zscale=zscale, phi=phi, theta=theta, zoom=zoom, 
          windowsize=windowsize)
render_snapshot()
```

### sedimentation

```{r, fig.height=15}
topo_mat %>%
  sphere_shade(texture = "bw") %>%
  add_overlay(sedimentoverlay, alphalayer = 0.6) %>%
  add_overlay(coastoverlay) %>% 
  add_overlay(tracksoverlay) %>% 
#  add_overlay(sampleoverlay_shades) %>%   
#  add_overlay(sampleoverlay) %>%
  plot_map()
```

```{r, fig.height=15}
topo_mat %>%
  sphere_shade(texture = "bw") %>%
  add_overlay(sedimentoverlay, alphalayer = 0.6) %>%
  add_overlay(coastoverlay) %>% 
  add_overlay(tracksoverlay) %>% 
#  add_overlay(sampleoverlay_shades) %>%   
#  add_overlay(sampleoverlay) %>%
    plot_3d(topo_mat, zscale=zscale, phi=phi, theta=theta, zoom=zoom, 
          windowsize=windowsize)
render_snapshot()
```





## 3D Animation

```{r, eval=FALSE}
library(particles)
library(tidygraph)
library(ggraph)
#library(ggplot2)
library(gganimate)
library(dplyr)

## Assuming your dataframe is named df
## Sample 1% of the particles
set.seed(123) # For reproducibility
sampled_particles <- df %>%
  distinct(particle) %>%
  slice(seq(1, n(), by = 100)) #sample_frac(0.005)
## Filter the dataframe to include only the sampled particles
sampled_df <- df %>%
  filter(particle %in% sampled_particles$particle)

##### First effort using ggplot, and a 2D representation, this works:

## Convert raster to a data frame for ggplot
bathy_df <- as.data.frame(RS.d, xy = TRUE)
colnames(bathy_df) <- c("x", "y", "elevation")

base_plot <- ggplot() +
  geom_raster(data = bathy_df, aes(x = x, y = y, fill = elevation)) +
  scale_fill_viridis_c() +
  theme_minimal()

## Add particle movement to the plot (for option works, but has no traces)
# particle_plot <- base_plot +
#   geom_point(data = sampled_df, aes(x = x, y = y, color = as.factor(particle)), size = 2) +
#   labs(color = "Particle") +
#   theme(legend.position = "none")
## geom_path for the traces (color can be "time" or "particles", or black I guess)
particle_plot <- base_plot +
  geom_path(data = sampled_df, aes(x = x, y = y, group = particle, color = as.factor(time)), linewidth = 1, alpha = 0.5) +
  geom_point(data = sampled_df, aes(x = x, y = y, color = as.factor(time)), size = 1) +
  labs(color = "Particle") +
  theme(legend.position = "none")

## check layout
particle_plot

## Animate the particle movements (first option works without traces)
# animation <- particle_plot +
#   transition_time(time) +
#   labs(title = "Time: {frame_time}")
animation <- particle_plot +
  transition_reveal(time) +
  labs(title = "Time: {frame_time}")

## and animate (~8min for 2000 particles across 113 timesteps)
animate(animation, nframes = 100, fps = 10, width = 800, height = 600)
## Save the animation
anim_save("particle_movement.gif", animation)


##### Second effort using a 3D representation, this doesn't work yet for the traces, only for the particle positions:
### Convert the raster to a matrix for 3D plotting:
library(plotly)
library(rgl)
bathy <- RS.d
bathy_matrix <- as.matrix(bathy, wide=TRUE)
x <- seq(from = xmin(bathy), to = xmax(bathy), length.out = ncol(bathy))
y <- seq(from = ymax(bathy), to = ymin(bathy), length.out = nrow(bathy))
## add bathymetry to df
sampled_df$z <- extract(bathy, cbind(sampled_df$x, sampled_df$y))[,1]

### Create a 3D plot of the bathymetry:
p <- plot_ly(z = ~bathy_matrix, x = ~x, y = ~y, type = "surface") %>%
  layout(scene = list(zaxis = list(title = "Elevation"),
                      xaxis = list(title = "Longitude"),
                      yaxis = list(title = "Latitude")))

## showing the initial positions of the particles
initial_positions <- sampled_df[sampled_df$time == 1, ]
initial_positions <- initial_positions[!is.na(initial_positions$z), ]
# Print the initial positions to verify
print(initial_positions)
p <- plot_ly(z = ~bathy_matrix, x = ~x, y = ~y, type = "surface") %>%
   layout(scene = list(zaxis = list(title = "Depth (z)"),
                       xaxis = list(title = "Longitude (x)"),
                       yaxis = list(title = "Latitude (y)"))) %>%
   add_trace(x = ~initial_positions$x, y = ~initial_positions$y, z = ~initial_positions$z,
             type = 'scatter3d', mode = 'markers',marker = list(size = 3, color = 'red'))
p

######################################################
## works until just here. something below doesn't work with the loop I guess
######################################################

# Initialize the 3D plot with bathymetry
p <- plot_ly(z = ~bathy_matrix, x = ~x, y = ~y, type = "surface") %>%
  layout(scene = list(zaxis = list(title = "Depth (z)"),
                      xaxis = list(title = "Longitude (x)"),
                      yaxis = list(title = "Latitude (y)")))

# Add particle traces
for (particle_id in unique(sampled_df$particle)) {
  particle_data <- sampled_df[sampled_df$particle == particle_id, ]
  particle_data <- particle_data[!is.na(particle_data$z), ]
  p <- p %>%
    add_trace(x = ~particle_data$x, y = ~particle_data$y, z = ~particle_data$z,
              type = 'scatter3d', mode = 'lines+markers', line = list(width = 2), marker = list(size = 3))
}

# Render the plot in RStudio Viewer
p

# Save the plot to an HTML file
htmlwidgets::saveWidget(p, "particle_movements.html", selfcontained = TRUE)
```



